\chapter{External dependencies}

\section{Dynamically Linked Dependencies?} \todo{Are these actually dynamically linked?}
\subsection{Libuv}
As mentioned in chapter \ref{planningTheProject}, CTaps is reliant on several
external libraries, both for testing and for core functionality. The largest
and most essential of these are libuv \cite{Libuv2025}. Libuv provides the central
event loop which allows CTaps to be asynchronous. It also provides several quality
of life features, such as built in support for asynchronous interactions with TCP
and UDP, through the \texttt{uv\_tcp\_t} and \texttt{uv\_udp\_t} handles respectively.
When adding protocols which libuv does not have native support for, such as QUIC,
it offers a custom work handle \texttt{uv\_work\_t}, which can be used to define
a custom piece of work. This can be used to define both a long-lived QUIC-based
server, or a short interaction with an existing server. Which underlying handle
is actually used in libuv is naturally abstracted away in CTaps, and for the
end-user of the library the interaction is the same no matter what. \todo {very informal lol}

\subsection{Glib}
In order to increase the reliability of internal data structures, as well as
minimizing the amount of time spent reinventing the wheel, CTaps relies on the
Glib library. Glib is a low level core library which is used at the core of
larger libraries such as GTK and GNOME \todo{citation}. Glib provides several
key data structures used internally in CTaps. Among others, it provides the
Hash Map used to keep track of received Connections in a Listener, the tree
used to gather candidates in the Candidate Gathering algorithm, among others.
While adding an external library, especially in core functionaly such as networking,
poses a risk, the benefits were deemed to be greater than the costs. Primarily,
this consideration was made on the basis of Glib being a relatively high-profile
library used in several large projects already, in addition to the benefits of
increased development speed, since CTaps would not have to implement its own 
basic data structures. Using a well-tested external library has the added benefit
of avoiding bugs, as if CTaps rolled its own data structures, it would have
to worry not only about bugs in the library logic, but also in the
data structures themselves.

In order to minimize how much a library consumers would have to
adhere to Glib, it was decided that the external API should be
kept free of Glib data structures, and that all external functions
should only take parameters defined in standard C or in CTaps itself.

\section{Statically Linked Dependencies}
Minor software which does not provide a core functionality to CTaps
is statically linked and is therefore distributed along with the
compiled executable. \todo{Hvordan blir biblioteket egentlig formatert}
This is done in order to make installation of
CTaps as easy as possible. Since the libraries are small, they
do not bloat the CTaps binary to the same degree as a larger library
such as libuv would, if it were to be statically linked.

\subsection{log.c}
Logging
https://github.com/rxi/log.c
