\chapter{The CTaps API}
CTaps is not a complete implementation of the TAPS standard,
as that would be too large of a scope for a singular thesis.
\todo{Check in on this later, describe how large of a subset}
This chapter will describe which parts of the TAPS API that has
been implemented in CTaps, and what operations that are supported
in CTaps.

\section{Basic TAPS Objects}
The basic TAPS objects:
\texttt{Preonnection}, \texttt{Connection}, \texttt{Local Endpoint}, \texttt{Remote Endpoint}, 
\texttt{Listener} and \texttt{Message} are each represented as a type defined
struct.

\begin{figure}[H]
\begin{minted}{C}
typedef struct Connection {
  TransportProperties transport_properties;
  LocalEndpoint local_endpoint;
  RemoteEndpoint remote_endpoint;
  // TODO - decide on if this has to be a pointer
  ProtocolImplementation protocol;
  uv_handle_t* protocol_uv_handle;
  ConnectionOpenType open_type;
  struct SocketManager* socket_manager;
  // TODO this is shared state and should be locked
  // Queue for pending receive() calls that arrived before the data
  GQueue* received_callbacks;
  GQueue* received_messages;
} Connection;

\end{minted}
\caption{Connection typedef structure}
\end{figure} \todo{Update this when having finalized Connection struct}

Each of these objects has a set of associated functions, in order to encapsulate
the responsibililty of this structure as best as possible. 

\begin{figure}[H]
\begin{minted}{C}
int send_message(Connection* connection, Message* message);
int receive_message(
    Connection* connection,
    ReceiveMessageRequest receive_message_cb
);
void connection_build_from_listener(
    Connection* connection,
    struct Listener* listener,
    RemoteEndpoint* remote_endpoint
);
void connection_close(Connection* connection);
\end{minted}
\caption{Function declarations from connection.h}
\end{figure}

\section{Event Loop}

\section{Callbacks}
Most modern interactions with networking applications are asynchronous
and the TAPS API reflects this. CTaps is therefore callback based. Any
interactions with the network or other I/O-bound activities are loaded
performed by the central event loop, which is managed by libuv. Whenever
a user wants to initiate a Connection, send a message or receive a message,
they pass a structure containing the relevant function pointers for the given
functionality. \todo{This struct way is not the current implementation, but perhaps the desired one?}
This includes function pointers for when a relevant action succeeds or fails.
If the initiation is successful, then the callback passed to the initiation
function which is associated with a successful initiation is invoked. \todo{Kronglete lol}
When invoking this callback, the new Connection is passed as a parameter,
along with a void pointer consisting of data specified by the user.
This void pointer allows the user to pass any data desired to the callback.

If initiating a preconnection fails because the endpoint could not be
reached, then the function pointer associated with the failure of creating
a connection is invoked. \todo{This is not currently implemented, show code when it has been}
As parameters to this callback, CTaps passes pointer to the relevant (pre?) connection \todo {sort out uncertainity}
and a reason for the failure, letting the user know what caused the error,
so that they can respond accordingly.

When receiving messages

When sending messages....


\section{Selecting the desired protocol}
Talk about how the user can use selection properties to select the desired protocol 
