\chapter{Transport System API} \label{transportSystemApi}
CTaps is not a complete implementation of the TAPS standard,
as that would be too large of a scope for a singular thesis.
\todo{Check in on this later, describe how large of a subset}
This chapter will describe which parts of the TAPS API that has
been implemented in CTaps, and what operations that are supported
in CTaps.

\section{Basic TAPS Objects}
The basic TAPS objects:
\texttt{Preonnection}, \texttt{Connection}, \texttt{Local Endpoint}, \texttt{Remote Endpoint}, 
\texttt{Listener} and \texttt{Message} are each represented as a type defined
struct.

\begin{figure}[H]
\begin{minted}{C}
typedef struct Connection {
  TransportProperties transport_properties;
  LocalEndpoint local_endpoint;
  RemoteEndpoint remote_endpoint;
  // TODO - decide on if this has to be a pointer
  ProtocolImplementation protocol;
  uv_handle_t* protocol_uv_handle;
  ConnectionOpenType open_type;
  struct SocketManager* socket_manager;
  // TODO this is shared state and should be locked
  // Queue for pending receive() calls that arrived before the data
  GQueue* received_callbacks;
  GQueue* received_messages;
} Connection;

\end{minted}
\caption{Connection typedef structure}
\end{figure} \todo{Update this when having finalized Connection struct}

Each of these objects has a set of associated functions, in order to encapsulate
the responsibililty of this structure as best as possible. 

\begin{figure}[H]
\begin{minted}{C}
int send_message(Connection* connection, Message* message);
int receive_message(
    Connection* connection,
    ReceiveMessageRequest receive_message_cb
);
void connection_build_from_listener(
    Connection* connection,
    struct Listener* listener,
    RemoteEndpoint* remote_endpoint
);
void connection_close(Connection* connection);
\end{minted}
\caption{Function declarations from connection.h}
\end{figure}

\section{Framer}
In CTaps, the Framer interface consists of a structure of function pointers,
which are implemented by the application. These are then passed to a preconnection
and are invoked at appropriate times during the lifetime of the Connection.
Along with the Connection, Message and MessageContext, the framer is also
passed a callback, which has to be invoked when the Framer is done modifying
the message. Since the Connection takes a deep copy of the message, the
Message data itself is owned by the protocol stack. This means that the
framer can modify the message in place and does not have to allocate new
memory, unless it is extending the length of the message.

An example
implementation from the CTaps framing tests can be seen in \cref{fig:framingTestImpl}.

\begin{figure}[p]
\begin{minted}{C}
// Message Framer Implementation Interface
struct ct_framer_impl_s {
  // Encode outbound message
  // Implementation should call the callback when encoding is complete
  void (*encode_message)(struct ct_connection_s* connection,
                        ct_message_t* message,
                        ct_message_context_t* context,
                        ct_framer_done_encoding_callback callback);

  // Decode inbound data into messages
  // Implementation should call ct_connection_deliver_to_app() for each complete message
  void (*decode_data)(struct ct_connection_s* connection,
                     const void* data,
                     size_t len,
                     ct_framer_done_decoding_callback callback);
};
\end{minted}
\caption{Framer structure definition from ctaps.h}
\end{figure}

\begin{figure}[p]
  \begin{minted}{C}
static void length_prepend_encode(ct_connection_t* connection,
                                   ct_message_t* message,
                                   ct_message_context_t* context,
                                   ct_framer_done_encoding_callback callback) {
    // Prepend the message length as a single ascii character
    // First byte is the length
    message->content = (char*)realloc(message->content, message->length + 1);
    memmove(message->content + 1, message->content, message->length);
    // Get ascii representation of length (as long as it is <= 9 bytes)
    message->content[0] = '0' + (char)(message->length);
    message->length += 1;

    callback(connection, message, context);
}

static void passthrough_decode(struct ct_connection_s* connection,
                     const void* data,
                     size_t len,
                     ct_framer_done_decoding_callback callback) {
    // Just pass through - deliver as-is
    ct_message_t* msg = (ct_message_t*)malloc(sizeof(ct_message_t));
    msg->content = (char*)malloc(len);
    msg->length = len;
    memcpy(msg->content, data, len);

    callback(connection, msg, NULL);
}

static ct_framer_impl_t length_prepend_framer = {
    .encode_message = length_prepend_encode,
    .decode_data = passthrough_decode
};

  \end{minted}
\caption{Example Framer implementation structure from the CTaps framing tests}
\label{fig:framingTestImpl}
\end{figure}

\section{Connection Cloning}
Cloning a Connection in TAPS yields a Connection Group. A Connection Group
is represented as an array of connections. The client operates on each individual
Connection, with certain operations, such as aborting/closing all being able
to operate on all connections in a connection group. For CTaps,
all Connections
have an associated Connection Group, with each Connection Group
only containing a single Connection until the original Connection has been cloned.

Invoking \mono{connection_clone} can lead to either a \mono{Ready} event or
a \mono{CloneError} event. Connection Cloning is primarily used to create
multiple logical Connections, and how this maps internally depends on the
underlying protocol being used. For protocols which support multistreaming,
such as QUIC, the connection group comes to represent a single overarching
Connection, while the individual connections in the group represents single
streams. The overarching connection can either be closed by a single operation
on the connection group or by closing each Connection individually.

For underlying protocols which do not support multistreaming, each cloned
Connection becomes functionally independent, while still supporting the
operations on Connection Groups. For a TCP Connection an independent Connection
will be established from an ephemeral local port, while UDP connections will
start listening on an ephemeral local port, immediately generate a \mono{Ready}
event, being ready to send messages from this ephemeral port.

\section{Event Loop}
CTaps is built around a single single-threaded event loop. This event
loop is powered by libuv \cite{Libuv2025}. This event loop is intialized
using \mono{ctaps_initialize()} and can be started using \mono{ctaps_start_event_loop()}.
Following the default libuv behaviour, this event loop will run until
there are no more open handles performing work. Any set up of Connections
which should be used during the lifetime of a given event loop has to
be performed before starting the event loop. The event loop takes over
execution of the main thread, and will block until execution of the underlying
handles completes, as described above.

This means that if the application invokes \mono{ct_preconnection_initiate},
then CTaps will attempt to connect to the specified endpoint when the event
loop is started. If this attempt is successful then an underlying handle
listening on a Local Endpoint is active. This means that the event loop
will not stop execution until this connection is closed.

Any behaviour that has to happen after an event loop is created has to be
specified in the callbacks provided by CTaps. For example, if a Connection
has to immediately send a message after successfully connection to an endpoint,
then this behaviour has to be defined in the \mono{ready} callback, which is
passed to a Preconnection through the \mono{ct_connection_callbacks_s} struct.

\todo{Write more here if we move to per-state config}

\section{Asynchrony} \label{asynchrony}
\todo{Can I write some numbers here, how much is I/O bound as a percentage?}
Most modern interactions with networking applications are asynchronous
and the TAPS API reflects this. CTaps is therefore callback based. Any
interactions with the network or other I/O-bound activities are loaded
performed by the central event loop, which is managed by libuv. Whenever
a user wants to initiate a Connection, send a message or receive a message,
they pass a structure containing the relevant function pointers for the given
functionality. \todo{This struct way is not the current implementation, but perhaps the desired one?}
This includes function pointers for when a relevant action succeeds or fails.
If the initiation is successful, then the callback passed to the initiation
function which is associated with a successful initiation is invoked. \todo{Kronglete lol}
When invoking this callback, the new Connection is passed as a parameter,
along with a void pointer consisting of data specified by the user.
This void pointer allows the user to pass any data desired to the callback.

If initiating a preconnection fails because the endpoint could not be
reached, then the function pointer associated with the failure of creating
a connection is invoked. \todo{This is not currently implemented, show code when it has been}
As parameters to this callback, CTaps passes pointer to the relevant (pre?) connection \todo {sort out uncertainity}
and a reason for the failure, letting the user know what caused the error,
so that they can respond accordingly.

Due to the asynchronous nature of the library, users can request to
receive messages at any time, and the library can receive messages
over the network and cache them at any time. When a user requests
to receive a message they pass a callback to the application, if a
message is available, that is, the library has received a message
over the network and has cached it, the callback is invoked immediately,
passing the first cached message as a parameter to the callback.
If no message has been received by the library, then the callback
itself is cached. The callback is then invoked the next time a message
is received over the network. This way, the user does not have to
spin while waiting for messages and is perform other work while
waiting for incoming messages. Each invocation only receives
a single message, and will always receive the first message in
the message queue.

When sending messages the message is not cached, instead
immediately passed to the libuv event loop. The message
may queue here for a while, depending on the number of tasks
in the loop.

\todo{Overview of callbacks here}


\section{Selecting the desired protocol}
Talk about how the user can use selection properties to select the desired protocol 

\section{Control Flow}
Since the primary functionality of the library is handled
by the underlying libuv event loop, the control flow has
to be handed over to the event loop after the primary setup
of an application has been handled. Further control has
to be
handled through the callbacks invoked by the library.

\section{Events}
As mentioned in \cref{asynchrony}, Taps is an asynchronous library.
For CTaps, this means that any asynchronous functions take a struct
containing function pointers as a parameter. In RFC 9622, each
asynchronous routine finishing invokes an Event \cite{trammellAbstractApplicationProgramming2025}.
The notation used in TAPS to denote events is shown in \cref{fig:ctapsEventNotation}.
A full list of events and the objects that spawn them can
be seen in \cref{fig:tapsEvents}. For each Event, CTaps
features an associated callback function.

\begin{figure}[H]
\centering
\begin{BVerbatim}
Object -> Event<>
\end{BVerbatim}
\caption{TAPS Event notation of Object spawning an Event}
\label{fig:ctapsEventNotation}
\end{figure}

This associated callback is set at various times. Mostly,
they are set when initially building the object associated
with the Event. For exapmle, the Ready event is set
when initializing a preconnection. \todo{For listeners as well or?}
In this case, the callback structure is stored in the
corresponding object, so the callback can be invoked when
the given Event occurs.
However, some Events are passed when inititating an asynchronous action,
such as CloneError, where the function pointer is passed as a parameter
to the Connection.clone() function. \todo{Make sure that this is true, will we implement clone?}

In order to allow the user to pass data into the callback, the structure
containing callbacks also contains a void pointer. This void pointer
is then passed as a parameter to the callback. This allows the user
implementation of the callback to dereference the void pointer
and pass any data to the callback, to make it accessible in
the body of the callback itself. This is a standard 
way if passing data to callbacks in C and can be
seen for example in the libuv library, where the
\texttt{uv\_handle\_t} struct contains a void pointer which
can be dereferenced in any callbacks \cite{libuvUv_handle_tDocumentation2025}.

\begin{figure}[H]
\centering
\begin{minted}{C}
typedef struct ConnectionCallbacks {
  int (*connection_error)(struct Connection* connection, void* udata);
  int (*establishment_error)(struct Connection* connection, void* udata);
  int (*expired)(struct Connection* connection, void* udata);
  int (*path_change)(struct Connection* connection, void* udata);
  int (*ready)(struct Connection* connection, void* udata);
  int (*send_error)(struct Connection* connection, void* udata);
  int (*sent)(struct Connection* connection, void* udata);
  int (*soft_error)(struct Connection* connection, void* udata);
  void* user_data;
} ConnectionCallbacks;
\end{minted}
\label{fig:connectionCallbackStructure}
\caption{Event callbacks associated 1-1 with a Connection object}
\end{figure}

\begin{figure}[H]
\centering
\begin{minted}{C}
typedef struct ReceiveCallbacks {
  int (*receive_callback)(
    struct Connection* connection,
    Message** received_message,
    MessageContext* ctx,
    void* user_data
  );
  int (*receive_error)(
    struct Connection* connection,
    MessageContext* ctx,
    const char* reason,
    void* user_data
  );
  int (*receive_partial)(
    struct Connection* connection,
    Message** received_message,
    MessageContext* ctx,
    bool end_of_message, 
    void* user_data
  );
  void* user_data;
} ReceiveCallbacks;
\end{minted}
\label{fig:receiveCallbackStructure}
\caption{Event callbacks associated with each request to receive a message}
\end{figure}

\begin{figure}[H]
\centering
\begin{minted}{C}
typedef struct ListenerCallbacks {
  int (*connection_received)(
    struct Listener* listener,
    struct Connection* new_conn,
    void* udata
  );
  int (*establishment_error)(
    struct Listener* listener,
    const char* reason,
    void* udata
  );
  int (*stopped)(
    struct Listener* listener,
    void* udata
  );
  void* user_data;
} ListenerCallbacks;
\end{minted}
\label{fig:listenerCallbackStructure}
\caption{Event callbacks associated 1-1 with a Listener object}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{ ||l|c|c| } 
 \hline
 \textbf{Event}      & \textbf{Class}                    & \textbf{Description} \\ 
 \hline \hline
 CloneError          & Connection                        & cell9 \\ 
 \hline
 Closed              & Connection                        & cell9 \\ 
 \hline
 ConnectionError     & Connection                        & cell9 \\ 
 \hline
 ConnectionReceived  & Listener                          & cell9 \\ 
 \hline
 EstablishmentError  & Connection/Preconnection/Listener & cell9 \\ 
 \hline
 Expired             & Connection                        & cell9 \\ 
 \hline
 PathChange          & Connection                        & cell9 \\ 
 \hline
 Ready               & Connection                        & cell6 \\ 
 \hline
 ReceiveError        & Connection                        & cell9 \\ 
 \hline
 ReceivedPartial     & Connection                        & cell9 \\ 
 \hline
 RendezousDone       & Preconnection                     & cell9 \\ 
 \hline
 SendError           & Connection                        & cell9 \\ 
 \hline
 Sent                & Connection                        & cell9 \\ 
 \hline
 SoftError           & Connection                        & cell9 \\ 
 \hline
 Stopped             & Listener                          & cell9 \\ 
 \hline
\end{tabular}
\end{center}
\caption{Table of events and associated objects}
\label{fig:tapsEvents}
\end{figure}
