\chapter{Transport System Implementation}  \label{transportSystemImpl}

TODO - define transport system in TAPS overview

Maybe this is defined somewhere else, do something like "as mentioned in section..."

\section{Candidate Gathering} \label{sec:candidateGathering}
CTaps follows the recommended procedure for Candidate Gathering from RFC 9623
very closely. CTaps represents the candidates in a tree, using a GNode struct from
GLib to represent a N-ary tree. CTaps first builds the entire tree, resolving
all possible local interfaces, all protocols and all possible resolutions of
remote hostnames respectively. If the client has specified multiple remote endpoints,
then they are treated the same way as each invidvidual resolution from a hostname
endpoint. After the entire tree has been built, CTaps prunes the branches which
are not compatible with the required or prohibited selection properties. After pruning,
the tree is sorted according to the stated selection properties, as described in RFC 9623 
\cite{brunstromImplementingInterfacesTransport2025}.

\section{Candidate Racing} \label{sec:candidateRacing}

\section{Protocol Implementations}

\subsection{Protocol Interface}
In order to make the CTaps design as modular as possible and ease the
process of adding more protocols, CTaps uses an internal interface for a protocol,
which abstracts away the implementation details. This interface consists of
a struct containing the name of the protocols, and function pointers for
the necessary functions for the wider CTaps library to interact with the protocol.
Each protocol implementation is this way kept as small as possible, and is
isolated from the rest of the library.

Implementations of the interface also register
a list of the applicable selection properties, allowing the library
to later prune and sort the Candidate Tree protocols, when selecting
candidates.

\begin{figure}[H]
\begin{minted}{C}
typedef struct ProtocolImplementation {
  const char* name;
  SelectionProperties selection_properties;
  int (*init)(struct Connection* connection, InitDoneCb init_done_cb);
  int (*send)(struct Connection*, Message*);
  int (*receive)(struct Connection*, ReceiveMessageRequest receive_cb);
  int (*listen)(struct SocketManager* socket_manager);
  int (*stop_listen)(struct SocketManager*);
  int (*close)(const struct Connection*);
} ProtocolImplementation;
\end{minted}
\caption{Protocol Implementation interface}
\end{figure} \todo{Update this when having finalized the code}

Each protocol supported instantiates a single instance of the
protocol interface, and adds this to a global variable keeping
track of the supported protocols: 

\begin{figure}[H]
  \begin{minted}{C}
int ctaps_initialize(const char *cert_file_name, const char *key_file_name) {
  ctaps_event_loop = malloc(sizeof(uv_loop_t));
  uv_loop_init(ctaps_event_loop);

  register_protocol(&udp_protocol_interface);
  register_protocol(&tcp_protocol_interface);
  register_protocol(&quic_protocol_interface);

  ctaps_global_config.cert_file_name = cert_file_name ? strdup(cert_file_name) : NULL;
  ctaps_global_config.key_file_name = key_file_name ? strdup(key_file_name) : NULL;
  return 0;
}
  \end{minted}
  \caption{Example of adding protocol support}
\end{figure}

Adding the protocols is done at runtime in order to allow any
application using CTaps to add support for custom protocols if
desired. Because all Connections using the same protocol are associated
with the same protocol structure, the protocol structure itself
is naturally stateless, providing only read-only values and function
pointers where the necessary state has to be passed in as paramaters.
How a client can add custom protocols is discussed further in \cref{sec:addingNewProtocols}.


Each Connection is associated with a single protocol, which
is selected through the process of candidate gathering and candidate
racing as described in \cref{sec:candidateGathering} and \cref{sec:candidateRacing}.
This means, that whenever a Connection wants to interact with the network, it
will invoke a function pointer from its associated protocol:

\begin{figure}[H]
  \begin{minted}{C}
int send_message(Connection* connection, Message* message) {
  return connection->protocol.send(connection, message, NULL);
}
  \end{minted}
  \caption{Example usage of interface when sending Message with Connection}
\end{figure}




\subsection{Supported protocols}
Currently, CTaps natively supports three transportation layer protocols:
UDP, TCP and QUIC. TCP and UDP were added because of their ubiquity and
relative ease of implementation due to native support in libuv. QUIC was
added since having easier access to QUIC was one of the primary motivations
behind the thesis as a whole. Using CTaps, an application can get access to QUIC,
without having to implement custom fallback-logic or error handling. This
greatly eases the implementation complexity and could help QUIC adoption
on both the server and client side.

\subsubsection{UDP}
UDP is a connectionless, message-based unreliable protocol first designed
in 1980 \cite{postelUserDatagramProtocol1980}. It has since become one
of the two primary transport layer protocols used today and has native
support in libuv through the libuv \texttt{udp\_udp\_t} handle \cite{libuvUdpHandle}.
Due to its simplicity, there is fairly little logic required to integrate
UDP into CTaps, which can be seen by it being the protocol requiring the
fewest lines of code to implement, as seen in \cref{fig:protocolLoc}.

\begin{figure}[H]
\begin{minted}{C}
int udp_init(Connection* connection, InitDoneCb init_done_cb);
int udp_close(const Connection* connection);
int udp_send(Connection* connection, Message* message);
int udp_receive(Connection* connection, ReceiveMessageRequest receive_message_cb);
int udp_listen(struct SocketManager* socket_manager);
int udp_stop_listen(struct SocketManager* listener);

static ProtocolImplementation udp_protocol_interface = {
    .name = "UDP",
    .selection_properties = {
      .selection_property = {
        get_selection_property_list(create_property_initializer)
        [RELIABILITY] = {.value = {.simple_preference = PROHIBIT}},
      }
    },
    .send = udp_send,
    .init = udp_init,
    .receive = udp_receive,
    .close = udp_close,
    .listen = udp_listen,
    .stop_listen = udp_stop_listen
};
\end{minted}
\caption{UDP protocol implementation struct}
\end{figure} \todo{Update this when having finalized the code}

UDP does however not fit perfectly into the model assumed by
TAPS, which is that of a Connection-based system. There is therefore
several pieces of logic in the CTaps UDP implementation which adapt
the underlying UDP layer to this assumption. Opening a client is
fairly simple, as this simply creates a new UDP handle through
libuv, binds it to either the local endpoint required or an ephemeral one
before finally starting to receive data over that handle, and
invoking the \texttt{Ready}-callback for the associated Connection.
This does not actually transfer any data to the remote endpoint
before the clients sends its first message, and any data received
on the local handle has to be filtered based on remote endpoint,
before being passed to the associated Connection. \todo{Write
about how this filtering happens after impelmenting it}
A \texttt{Ready}-callback does therefore not necessarily
mean that the remote endpoint is actually reachable.
\todo{Or does it, what implications does candidate racing have here?}

Another piece of logic required to adapt UDP to the structure
of a TAPS library is that surrounding the Listener. A Listener
listens on a single \todo{correct?} port, creating a new 
Connection each time a new client connects. With UDP not
featuring a handshake, the Listener instead has to keep track
of which remote endpoints it has received messages from, creating
a new Connection whenever it receives a message from a remote endpoint
it has not seen before. This is similar to how TCP connections
are identified by a four-tuple of Source IP address, Source port,
Destination IP address and Destination port. Instead of the
demultiplexing to per-Connection happening as part of the
operating system stack as with TCP, it has to happen as
part of the CTaps custom UDP logic. After a message has
been received by a Listener, the Connection associated with
the sender is found, and any callback for received messages
for this connection is invoked, if registered by the client.
When sending data from a Connection created by a Listener, the
Connection itself contains the destination remote endpoint,
meaning that the Listener does not have to look up any state
in this case. \todo{Listeners should receive on seeveral protocols according to RFC9623}

For a TAPS system, Listeners are only responsible for receiving new
Connections, while the created Connections are independent
of the listener. This means that the created Connections should
remain active, even if the Listener is stopped. In order to ensure
the validity of the underlying UDP handles and other data structures
used, even if the Listener is stopped and freed by the client, CTaps
uses an underlying structure called a Socket Manager to actually
handle contain the libuv handle and to handle the demultiplexing of
messages. The Socket Manager contains everything necessary for keeping
track of the received Connections and for transferring data, while the
Listener contains all state necessary for spawning new Connections.
When a Listener is closed and freed, the Socket Manager remains active,
still receiving messages and passing them along to the associated Connection.
If a message from a new remote endpoint is received after closing a
Listener, the associated Socket Manager will simply discard it. The
Socket Manager needs to remain active as long as there is at least
one active associated Connection. It therefore has an internal counter,
keeping track of the associated Listener and the number of active Connections.
When a Connection is closed, it checks whether or not it was created from a Listener
and if it was, then it informs its associated Socket Manager,
which then internally checks whether there are any remaining associated
objects. If there are none, the socket manager closes the associated socket
through the associated protocol, and then finally frees itself. \todo{This might be more Socket Manager section fitting}

\todo{Any more UDP logic?}

\begin{figure}[H]
  \begin{minted}{C}
void socket_manager_decrement_ref(SocketManager* socket_manager) {
  socket_manager->ref_count--;
  log_debug("Decremented socket manager reference count, updated count: %d", socket_manager->ref_count);
  if (socket_manager->ref_count == 0) {
    int rc = socket_manager->protocol_impl.stop_listen(socket_manager);
    if (rc < 0) {
      log_error("Error stopping socket manager listen: %d", rc);
    }
    free(socket_manager);
  }
}
  \end{minted}
  \caption{Internal Socket Manager counter decrementation}
\end{figure}


 
\subsubsection{TCP}

\begin{figure}[H]
  \begin{minted}{C}
static ProtocolImplementation tcp_protocol_interface = {
    .name = "TCP",
    .selection_properties = {
      .selection_property = {
        get_selection_property_list(create_property_initializer)
        [RELIABILITY] = {.value = {.simple_preference = REQUIRE}},
        [MULTISTREAMING] = {.value = {.simple_preference = PROHIBIT}},
        [PRESERVE_ORDER] = {.value = {.simple_preference = REQUIRE}},
      }
    },
    .send = tcp_send,
    .init = tcp_init,
    .receive = tcp_receive,
    .close = tcp_close,
    .listen = tcp_listen,
    .stop_listen = tcp_stop_listen,
    .remote_endpoint_from_peer = tcp_remote_endpoint_from_peer
};
  \end{minted}
  \caption{TCP protocol implementation struct}
\end{figure}

\subsubsection{QUIC}
\begin{figure}[H]
\begin{minted}{C}
static ProtocolImplementation quic_protocol_interface = {
    .name = "QUIC",
    .selection_properties = {
      .selection_property = {
        get_selection_property_list(create_property_initializer)
        [RELIABILITY] = {.value = {.simple_preference = NO_PREFERENCE}},
        [PRESERVE_ORDER] = {.value = {.simple_preference = REQUIRE}},
        [PRESERVE_MSG_BOUNDARIES] = {.value = {.simple_preference = NO_PREFERENCE}},
        [MULTISTREAMING] = {.value = {.simple_preference = NO_PREFERENCE}},
        [ACTIVE_READ_BEFORE_SEND] = {.value = {.simple_preference = PROHIBIT}}, // Temporary - to make it easy to ban quic
      }
    },
    .send = quic_send,
    .init = quic_init,
    .receive = quic_receive,
    .close = quic_close,
    .listen = quic_listen,
    .stop_listen = quic_stop_listen,
    .remote_endpoint_from_peer = quic_remote_endpoint_from_peer
};
\end{minted}
\caption{QUIC protocol implementation struct}
\end{figure} \todo{Update on final code}


\subsection{Adding new protocols} \label{sec:addingNewProtocols}

When implementing a new protocol, all that is needed is an
implementation of the struct, filling out the function pointers
with custom behaviour for the given protocol. Libuv features
designated handles for TCP and UDP support, while generic
handles can be used for other protocols. 
With the abstraction of the protocol interface, adding new protocols
becomes fairly simple, depending on the complexity of the protocol
itself.

\begin{figure}[H]
\begin{center}
\begin{tabular}{ ||l|l|c| } 
 \hline
 \textbf{Protocol}   & \textbf{Source files}             & \textbf{Lines of Code} \\ 
 \hline \hline
 UDP                 & \texttt{udp.h \& udp.c}                    & 263 \\ 
 \hline
 TCP                 & \texttt{tcp.h \& tcp.c}                    & 344 \\ 
 \hline
 QUIC                & \texttt{quic.h \& quic.c}                  & 743 \\
 \hline
\end{tabular}
\end{center}
\caption{Overview of Lines of Code (LoC) on a per-protocol basis}
\label{fig:protocolLoc}
\end{figure} \todo{Update lines of code at end of development}
