\begin{abstract}
Network sockets have been the dominant way for applications to interact with the
network for several decades. The design of the BSD socket API forces applications to
bind to an underlying transport protocol at compile time, which, together
with middlebox interference, greatly hinders innovation at the transport layer.

The IETF Transport Services (TAPS) working group was established to develop a
protocol-agnostic interface to the network, to decouple applications from the underlying
transport protocols and ease innovation.
The working group has produced three core RFCs: 9621, 9622 and 9623, which cover the architecture,
API and implementation guidance of a TAPS System.
No open-source implementations
of TAPS support QUIC, making it unclear exactly how QUIC, as the most successful recent protocol innovation
at the transport layer, maps onto the TAPS API.

This thesis presents CTaps, a C implementation of TAPS using the libuv event loop library and the Picoquic
QUIC implementation.
CTaps supports QUIC, TCP and
UDP, along with several important QUIC features, such as multistreaming, connection
migration and session resumption. In addition, CTaps supports important internal TAPS
features such as candidate gathering and candidate racing.

A survey of 28 QUIC features found that 24 are
either fully or partially available through the TAPS API, with only 5 gaps.
We evaluate CTaps against native TCP, UDP and Picoquic QUIC clients in term sof handshake
latency, ping overhead, multistream transfer, candidate racing and connection migration.
We argue that CTaps effectively serves as a reference for future implementations of TAPS,
especially when supporting QUIC.
\end{abstract}
