\begin{preface}
  I would like to thank my supervisor, Michael Welzl, for his
  guidance, feedback and enthusiasm throughout the entire process. He
  has helped shape this thesis and has provided
  invaluable help during the process of writing it.

  I would like to thank all my friends, my family and my girlfriend
  for your support during this project. You have made working
  on this thesis a breeze. A special thanks to Henrik Heidenberg for taking the time
  to proofread this thesis.

  I would like to thank every member of Radio Nova for making
  my time as a student here in Oslo enjoyable. I stumbled in
  but was welcomed with open arms.

  Finally, thank you to all developers of every open source
  tool I have used throughout this thesis, especially
  Neovim, Zellij, Picoquic and libuv.
\begin{flushright}
  Ingebrigt Kristoffer Thomassen Hovind \\
  Oslo, May 2026
\end{flushright}
\end{preface}

\newpage
\section{Declaration of the use of generative AI}
In this scientific work, generative artificial intelligence (AI) has been used.
All data and personal information have been processed in accordance with the
University of Oslo's regulations, and I, as the author of the document,
take full responsibility for its content, claims, and references.
An overview of the use of generative AI is provided below.

\begin{itemize}
  \item \textbf{Figures:} Claude~\cite{claude}, developed by Anthropic has been used to generate initial scaffolding for all TikZ figures
      featured in the thesis. After initial scaffolding they have
      been heavily modified by the author to achieve their final design.
    \item \textbf{Source Code:} Claude has been used to generate various parts of the
      source code.
      All AI-generated code has been reviewed and modified
      when necessary.
      The behavior of these
      sections has been verified both through manual inspection and interaction with
      tests and other parts of CTaps.
    \item \textbf{Tooling:} Claude has been used for
      architecture discussions, bug fixes and other minor interventions in all
      parts of this project.
      Claude has also been used to debug \LaTeX{} code for this document itself
      and to discover new features of \LaTeX{} such as microtype.
    \item \textbf{Experiments:}
      Claude has been used for the initial design of
      the experiments and for discussing what results are valuable to display.
      Claude has also been used for generating the scaffolding around experimenting 
      and visualization. This scaffolding consists of
      the ping servers and the Python scripts \mono{run_benchmarks.py} and
      \mono{visualize.py}. These have been modified by the author after-the-fact
      and have been verified to work as expected through manual inspection.
    \item \textbf{Prose:} For the thesis text itself Claude has been
      used to proofread and for feedback on structure, but all
      actual writing has been done by the author.
    \item \textbf{Other AI models:} Other models such as Gemini \cite{gemini} developed by Google and GPT UiO, maintained
      by the UiO IT-department with support for various models, have been used in the
      same way as Claude, but to a much smaller degree.
\end{itemize}
