\chapter{Background}\label{chap:background}
This chapter will lay out background information necessary
for understanding TAPS and CTaps.
It will describe the necessary networking concepts,
transport layer protocols and the current barriers
for innovation on the transport layer.

\section{Network Sockets}
When it comes to network programming,
the network socket in its most basic definition is an abstraction
for a network connection through which an application can send
and receive data~\cite[p. 8]{donahooTCPIPSockets2009}. 
First mentioned in 1970 in relation to ARPANET \cite{rfc33}, the
simplicity of sockets and their ubiquity in all UNIX distributions
since version 4.2 of the Berkeley Software Distribution have
led to them becoming the de facto standard for network
programming~\cite[p.~502]{tanenbaumComputerNetworks2011}.

Modern sockets are standardized across UNIX-like systems through
the POSIX standard. Implementations of the C POSIX library define
the socket API primarily through the \texttt{sys/socket.h} header file~\cite{posix:2024:socket}.

When creating a socket on an UNIX-like system, the application
invokes the \texttt{socket} system call. This system call
\enquote{creates an endpoint for communication and returns a file 
descriptor that refers to that endpoint}~\cite{Socket2LinuxManual}.
This system call takes three parameters, the domain, the type
and the protocol which will be used for the socket.

The domain specifies the type of
addressing which will be used, for example \texttt{AF\_UNIX}
for local sockets, \texttt{AF\_INET} for IPv4 and \texttt{AF\_INET6}
for IPv6. The type specifies the protocol which will be used over the socket, for
example \texttt{SOCK\_DGRAM} for datagrams and \texttt{SOCK\_STREAM}
for a reliable stream of bytes. The type usually corresponds to a
single transport layer protocol, usually UDP or TCP.

The protocol parameter is used if there are multiple protocols
which support the requirements given by the domain and type.

\section{Barriers to Protocol Adoption}\label{sec:transport-protocol-barriers}
The fact that an application has to specify both the domain
and the type of the socket means that the
underlying addressing and the transport layer protocol are bound at compile time.
Updating
the underlying transport protocol will therefore require at least recompilation and 
likely a refactoring of the application. 
This naturally makes adopting new transport protocols difficult, even
if a new protocol is better suited to the application's requirements.

Another challenge with the adoption of new transport layer protocols
is network middleboxes. Network Middleboxes have been
optimized to support TCP and UDP and may sometimes
drop packets if the protocol field in the IP header
does not match TCP or UDP.

In an attempt to combat this aspect of network ossification, transport layer protocols are often built
using UDP as a substrate, allowing them to use the same IP
protocol number as UDP. Examples of this include QUIC \cite{rfc9000} and
SCTP over UDP \cite{rfc6951}.
While this does present some overhead,
it is fairly minimal with the UDP header only being \qty{8}{\Bytes}.
Since UDP presents no guarantees of reliability or ordered
delivery, protocols can choose to develop only the features they need.

\section{TCP}
TCP is by far the most common transport layer protocol used on the
internet today. It is a connection-oriented protocol and provides
a reliable, in-order transfer of a stream of bytes to the application layer.
It has also become common to use TCP for message-based applications,
using a framing protocol on top of TCP to divide the TCP byte stream into
discrete messages. A description of TCP was first published as an RFC~in 1974~\cite{rfc675},
and it has since grown to become ubiquitous across the internet.

In TCP, connections are uniquely identified by a four-tuple consisting
of the source IP address, the source port, the destination IP address
and the destination port.

\section{UDP}
UDP has, along with TCP, grown to become one of the two dominant
transport layer protocols on the internet. It is a connectionless
protocol and provides an unreliable, unordered transfer of datagrams
to the application layer. It is often used for applications which
prioritize low latency over reliability.

UDP is designed to be a lightweight protocol, and it uses
a simple 8-byte header. The header contains
only the source port, destination port, length and checksum.

\section{SCTP}
SCTP is a transport layer protocol which was first introduced in 2000~\cite{rfc2960},
and it features several innovations over TCP.
Notably, it was the first protocol to support both multistreaming and multihoming.
Multistreaming allows multiple independent logical streams to be multiplexed over
a single connection and is described in detail in \cref{sec:background-multistreaming}.
Multihoming allows a connection to span multiple network interfaces simultaneously,
a concept explored further in the context of QUIC in \cref{sec:connection-migration}.

Despite its innovations, SCTP has struggled with widespread adoption
for several reasons.
First of all, it uses a separate protocol number from TCP and UDP, which
makes it vulnerable to middlebox interference.
SCTP over UDP~\cite{rfc6951} was developed in an attempt to mitigate these issues, but
it too has struggled to achieve widespread use.

The second reason for why SCTP has struggled with adoption is that the standard
socket API has been ossified itself \cite{sctpstruggles}.
The ubiquity of TCP and UDP means that the socket API
is tuned to these two transports
and does not present the advantages of a new protocol
such as SCTP to applications. This means that the socket
API had to be extended in order to support SCTP.

QUIC addresses both of these barriers by using UDP as 
a substrate and implementing the protocol entirely in user space,
while still preserving multistreaming and multihoming.

\section{Multistreaming}\label{sec:background-multistreaming}
The simplest way of organizing network transfers is to open
a new TCP connection for each object which needs to be transferred.
This creates independent streams each responsible for transferring
a single object and was the basis of early HTTP-standards.

Utilizing TCP this way has the major drawback that each new
connection has to go through slow start. This means that
when transferring small objects, the full bandwidth of the
connection cannot be utilized.
In TCP, this issue is usually mitigated by re-using the same
TCP connection for several transfers. An example
here is HTTP/1.1,
which introduced the concept of persistent
TCP connections when fetching multiple URLs~\cite[Sec.~8.1]{rfc2616}.

This way allows for a more
efficient use of bandwidth, but also introduces the issue of
head-of-line blocking. Head-of-line blocking is a performance bottleneck where a queue of packets
is held up by an issue with earlier packet. Commonly caused
either by reordering or packet loss.

Multistreaming is an innovation which solves both of these issues.
It allows a single connection to be reused across multiple logical streams,
improving bandwidth efficiency while avoiding head-of-line blocking by
only enforcing ordering within individual streams.
The bandwidth gains from CWND reuse are illustrated in~\cref{fig:connection-clone-gains},
while \cref{fig:background-multi-streaming} shows how separate streams avoid
head-of-line blocking caused by reorderi.



\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\small, >=stealth]

%% ============ LEFT PANEL: TCP ============
\begin{scope}

  % Lost-capacity shading
  \fill[blue!10]
    (3.1, 3.0) -- (5.1, 3.0)
    .. controls (4.6, 3.0) and (3.6, 0.15) .. (3.1, 0.15) -- cycle;

  % Axes
  \draw[->, thick] (0, 0) -- (5.6, 0) node[right] {Time};
  \draw[->, thick] (0, 0) -- (0, 4.2) node[above] {CWND};

  % W_max reference
  \draw[dashed, gray!60] (0, 3.0) -- (5.3, 3.0);
  \node[right, gray!60, font=\scriptsize] at (5.5, 3.0) {$W_{\max}$};

  % Transfer 1: quick ramp then flat at W_max
  \draw[very thick, blue!70!black]
    (0, 0.15) .. controls (0.5, 0.15) and (1.5, 3.0) .. (2.0, 3.0)
    -- (3.0, 3.0);

  % Transfer boundary
  \draw[dashed, gray] (3.1, -0.3) -- (3.1, 3.9);
  \node[above, font=\scriptsize, gray] at (3.1, 3.9) {New Transfer};

  % Transfer 2: TCP slow starts again
  \draw[very thick, blue!70!black]
    (3.1, 0.15) .. controls (3.6, 0.15) and (4.6, 3.0) .. (5.1, 3.0);

  % Lost capacity annotation
  \node[font=\scriptsize, black] at (3.85, 2.75) {Lost};
  \node[font=\scriptsize, black] at (3.85, 2.35) {capacity};

  % X-axis region labels
  \node[below, font=\scriptsize] at (1.5, 0) {First Transfer};
  \node[below, font=\scriptsize] at (4.2, 0) {Second Transfer};

  % Panel title
  \node[font=\small\bfseries] at (3.0, 4.7) {TCP};

\end{scope}

%% ============ RIGHT PANEL: QUIC ============
\begin{scope}[xshift=6.7cm]

  % Axes
  \draw[->, thick] (0, 0) -- (5.6, 0) node[right] {Time};
  \draw[->, thick] (0, 0) -- (0, 4.2) node[above] {CWND};

  % W_max reference
  \draw[dashed, gray!60] (0, 3.0) -- (5.3, 3.0);
  \node[right, gray!60, font=\scriptsize] at (5.5, 3.0) {$W_{\max}$};

  % Transfer 1: quick ramp then flat at W_max
  \draw[very thick, green!50!black]
    (0, 0.15) .. controls (0.5, 0.15) and (1.5, 3.0) .. (2.0, 3.0)
    -- (3.0, 3.0);

  % Transfer boundary
  \draw[dashed, gray] (3.1, -0.3) -- (3.1, 3.9);
  \node[above, font=\scriptsize, gray] at (3.1, 3.9) {New Transfer};

  % Transfer 2: QUIC resumes at W_max
  \draw[very thick, green!50!black]
    (3.1, 3.0) -- (5.1, 3.0);

  % CWND reuse annotation
  \node[font=\scriptsize, green!40!black] at (4.1, 3.6) {Stream};
  \node[font=\scriptsize, green!40!black] at (4.1, 3.2) {multiplexing};

  \node[below, font=\scriptsize, green!40!black] at (1.8, 2) {First};
  \node[below, font=\scriptsize, green!40!black] at (1.8, 1.6) {Stream};

  \node[below, font=\scriptsize, green!40!black] at (4.2, 2) {Second};
  \node[below, font=\scriptsize, green!40!black] at (4.2, 1.6) {stream};

  % X-axis region labels
  \node[below, font=\scriptsize] at (1.5, 0) {First Transfer};
  \node[below, font=\scriptsize] at (4.2, 0) {Second Transfer};

  % Panel title
  \node[font=\small\bfseries] at (3.0, 4.7) {Multistreaming};

\end{scope}

\end{tikzpicture}
\caption[CWND re-use for multistreaming]
{
An example of CWND re-use with multistreaming. If TCP is organized into logical
streams then each subsequent stream will be a separate TCP connection
and each will have to go through slow start. Multistreaming allows multiple
streams to be multiplexed over a single connection and they can
therefore re-use the achieved CWND.
}
\label{fig:connection-clone-gains}
\end{figure}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  font=\small,
  >=latex,
  node distance=1.2cm and 1.6cm,
  pkt/.style={circle,draw,minimum size=6mm,inner sep=0pt},
  lost/.style={pkt,fill=red!50},
  ok/.style={pkt,fill=orange!40},
  waiting/.style={pkt,fill=yellow!40},
  conn/.style={
  draw,
  shape=cylinder,
  cylinder uses custom fill,
  cylinder body fill=blue!10,
  cylinder end fill=blue!20,
  minimum height=6cm,  % thickness
  minimum width=2cm    % length
},
  endpoint/.style={
    draw,rounded corners=8pt,
    minimum width=1.8cm,
    minimum height=2.1cm,
    align=center,
    fill=gray!10,
  }
]

%======================
% TOP: TCP / HTTP/2
%======================
\begin{scope}
\node[endpoint] (tcpclient) at (-5,2.5) {Client};
\node[endpoint] (tcpserver) at (5,2.5) {Server};


% TCP connection "pipe"
\node[conn] (tcpconn) at (0,2.5) {};

\draw[->,thick,gray!70]
  ($(tcpconn.east)+(-1.5,1.2cm)$) -- ($(tcpconn.west)+(1.5cm,1.2cm)$);

\draw[dashed] (tcpclient.east)++(0,0.5) -- (tcpconn.west);
\draw[dashed] (tcpclient.east)++(0,00) -- (tcpconn.west);
\draw[dashed] (tcpclient.east)++(0,-0.5) -- (tcpconn.west);

\draw[dashed] (tcpserver.west)++(0,0.5) -- (tcpconn.east);
\draw[dashed] (tcpserver.west)++(0,0) -- (tcpconn.east);
\draw[dashed] (tcpserver.west)++(0,-0.5) -- (tcpconn.east);
\draw[dashed] (tcpconn.west) -- (tcpconn.east);



% packets in order (1..8) inside pip
% place first packet at the left edge of tcpconn, slightly inside
\node[lost]    (p1) at ($(tcpconn.west)+(0.6cm,0)$)  {\scriptsize 5};
\node[waiting] (p2) at ($(p1)+(0.9cm,0)$)            {\scriptsize 1};
\node[ok]      (p3) at ($(p2)+(0.9cm,0)$)            {\scriptsize 3};
\node[ok]      (p4) at ($(p3)+(0.9cm,0)$)            {\scriptsize 4};
\node[waiting] (p5) at ($(p4)+(0.9cm,0)$)            {\scriptsize 2};
\node[lost]    (p6) at ($(p5)+(0.9cm,0)$)            {\scriptsize 6};

\node[font=\small\bfseries] at (0.0, 4.0) {TCP};
\end{scope}

%======================
% BOTTOM: QUIC
%======================

\begin{scope}[yshift=-3.4cm]
\node[endpoint] (quicclient) at (-5,2.5) {Client};
\node[endpoint] (quicserver) at (5,2.5) {Server};

% TCP connection "pipe"
\node[conn, cylinder body fill=green!10, cylinder end fill=green!20] (quicconn) at (0,2.5) {};

\draw[dashed]($(quicclient.east)+(0,-0.5)$) -- ($(quicconn.west)+(0,-0.5)$);
\draw[dashed]($(quicclient.east)+(0,0)$) -- ($(quicconn.west)+(0,0)$);
\draw[dashed]($(quicclient.east)+(0,+0.5)$) -- ($(quicconn.west)+(0,+0.5)$);

% \draw[dashed] (quicconn.west)+(0,-0.5)   -- (quicconn.east)+(0,-0.5);
% \draw[dashed] (quicconn.west)            -- (quicconn.east);
\draw[dashed]($(quicconn.west)+(0,-0.5)$) -- ($(quicconn.east)+(0,-0.5)$);
\draw[dashed]($(quicconn.west)+(0,0)$) -- ($(quicconn.east)+(0,0)$);
\draw[dashed]($(quicconn.west)+(0,+0.5)$) -- ($(quicconn.east)+(0,+0.5)$);

\draw[dashed]($(quicserver.west)+(0,-0.5)$) -- ($(quicconn.east)+(0,-0.5)$);
\draw[dashed]($(quicserver.west)+(0,0)$) -- ($(quicconn.east)+(0,0)$);
\draw[dashed]($(quicserver.west)+(0,+0.5)$) -- ($(quicconn.east)+(0,+0.5)$);

\draw[->,thick,gray!70]
  ($(quicconn.east)+(-1.5,1.2cm)$) -- ($(quicconn.west)+(1.5cm,1.2cm)$);


\node[lost]    (p1) at ($(quicconn.west)+(0.6cm,-0.5)$)  {\scriptsize 5};
\node[waiting] (p2) at ($(p1)+(0.9cm,1.0)$)            {\scriptsize 1};
\node[ok]      (p3) at ($(p2)+(0.9cm,-0.5)$)            {\scriptsize 3};
\node[ok]      (p4) at ($(p3)+(0.9cm,0)$)            {\scriptsize 4};
\node[waiting] (p5) at ($(p4)+(0.9cm,0.5)$)            {\scriptsize 2};
\node[lost]    (p6) at ($(p5)+(0.9cm,-1.0)$)            {\scriptsize 6};

\node[font=\small\bfseries] at (0.0, 4.0) {Multistreaming};
\end{scope}

\end{tikzpicture}
\caption[Avoiding head-of-line blocking with multistreaming]
{
An example of head-of-line blocking caused by reordering. If packets were sent in order from 1 to 6,
then TCP cannot hand the data in packet 5 to the application until packets 1--4 have
arrived. For multistreaming, order is not enforced between streams and this issue
is avoided.
}
\label{fig:background-multi-streaming}
\end{figure}

\section{Network Paths}\label{sec:back-multihoming}
Modern devices are often connected to multiple networks simultaneously.
For example, a laptop may be connected to both a Wi-Fi network and a wired Ethernet at the same time.
Each separate network is interacted with through a separate network interface
and each network interface is associated with distinct source IP addresses for outgoing data.
If a device can reach the same endpoint through several network interfaces,
then it can utilize multiple interfaces for a single network connection as long as this is supported
by the underlying transport protocol.

The combination of source IP, source port, destination IP and destination port
is referred to as a \term{network path} following the terminology used by QUIC in RFC~9000~\cite{rfc9000}.
Since each network interface is associated with distinct IP addresses,
each interface that can reach the same destination will represent a distinct network path.
Utilizing multiple network interfaces for a single connection can improve
both stability and the throughput of the connection.
Support for multiple network paths is a notable feature of QUIC
and is described in detail in~\cref{sec:quic}.


\section{TLS}
Transport layer security (TLS) is a cryptographic protocol which provides secure communication
over a computer network. It is commonly used on top of TCP and is
most visible as the basis for HTTPS, the secure version of HTTP.

TLS features its own handshake protocol when establishing a connection,
in order to exchange the cryptographic parameters needed.
This means that connections using TLS have to spend an additional round trip time (RTT)
during connection establishment, compared to a non-secure TCP connection.
To mitigate this, TLS 1.3 also supports 0-RTT handshakes
which allow a client to send data during the initial handshake when resuming
a previous connection to a server~\cite{rfc8446}.

TLS is a mandatory part of QUIC, which inherits both the security
and the 0-rtt handshakes described above.


\section{QUIC}\label{sec:quic}
QUIC is a protocol which was originally developed by Google and then formalized
through RFC~9000 in 2021~\cite{rfc9000}. It is developed in user-space over UDP
and is intended to be a replacement for TCP.
By 2020, over \qty{75}{\percent} of traffic to Meta servers was using
QUIC~\cite{joras2020quic}.

QUIC is a reliable, connection-oriented protocol and provides several
features which are not available
in TCP. TLS is a mandatory part of QUIC 
and most of the QUIC header is also encrypted, unlike TLS over TCP.
Encrypting the header hides as much as the wire image as possible
from middleboxes, combating future network ossification.

While QUIC does present an overarching connection, the
abstraction for each actual data transfer is called a stream.
A stream is a unidirectional or bidirectional ordered
continuous transfer of data, much like what an application sees when using
TCP. Streams can be created by both peers,
and multiple streams can exist over a single QUIC connection.
QUIC mandates support for in-order delivery within a single
stream, but not between streams. This avoids the issue
of head of line blocking, where the
transfer of a small object potentially has to wait for the
transfer of a previous large object to complete before being
handed to the application.

For QUIC, a single connection can exist over several paths and
congestion control is done on a per-path basis. Multiple streams
can be transmitted over a single connection and data from
multiple stream frames can be transmitted in a single packet.
If a packet is lost, only the streams with data contained
in that packet will be affected.

A given stream can exist in several different states.
These states differ based on whether the stream is sending or receiving
data. A bidirectional stream consists of both sending and receiving parts,
and each part has a separate state.
A flow chart describing the QUIC stream state machine can be
seen in~\cref{fig:stream_states}.

\begin{figure}[htbp]
    \centering
    \begin{BVerbatim}
   o
   | Create Stream (Sending)
   | Peer Creates Bidirectional Stream
   v
+-------+
| Ready | Send RESET_STREAM
|       |-----------------------.
+-------+                       |
   |                            |
   | Send STREAM /              |
   |      STREAM_DATA_BLOCKED   |
   v                            |
+-------+                       |
| Send  | Send RESET_STREAM     |
|       |---------------------->|
+-------+                       |
   |                            |
   | Send STREAM + FIN          |
   v                            v
+-------+                   +-------+
| Data  | Send RESET_STREAM | Reset |
| Sent  |------------------>| Sent  |
+-------+                   +-------+
   |                           |
   | Recv All ACKs             | Recv ACK
   v                           v
+-------+                   +-------+
| Data  |                   | Reset |
| Recvd |                   | Recvd |
+-------+                   +-------+
    \end{BVerbatim}
    \caption[QUIC stream state machine]{States of a QUIC stream sending data~\cite{rfc9000}.}
    \label{fig:stream_states}
\end{figure}

\subsection{Congestion Control}
Like TCP, QUIC also implements congestion control, but there are
some differences between the algorithms. A key concept
to QUIC's congestion control is the \term{packet number space} which
is defined as \enquote{the context in which a packet can be processed
and acknowledged}~\cite{rfc9000}. 

The packet number space reflects the level of encryption that has been reached, and there 
are three separate packet number spaces. \term{Initial space},
for the initial handshake packets, \term{handshake space}, for packets
exchanged during the cryptographic handshake, and \term{application
data space}, for packets exchanged after the cryptographic handshake
has been completed.

Separating these spaces ensures that acknowledgments of packets at one level do not cause
spurious retransmission at another encryption level.
Although the packet number spaces are handled
separately for acknowledgments, they are
unified when it comes to congestion control and round trip
time measurements~\cite[Sec.~4.1]{rfc9002}.

Another difference between how QUIC and TCP handle acknowledgments
is that QUIC separates the transmission order from the delivery order.
This is done by using stream offset as delivery order, and packet number as
transmission order. This removes ambiguity about whether the
original or the retransmitted packet is acknowledged in the case
of retransmissions, making RTT measurements in these cases more accurate~\cite[Sec.~4.2]{rfc9002}.

\subsection{Transport parameters}
Transport parameters in QUIC are settings which are transferred to the
peer during the QUIC handshake. These can be used to for example
limit the initial amount of streams the peer can open or the amount
of data it can send.

\subsection{Frames}
In QUIC, a frame is defined as \enquote{a unit of structured protocol information}~\cite{rfc9000}.
Different types of frames are used to carry different types of information
and data. Examples are \mono{STREAM} frames which are used to carry
data over a stream or \mono{NEW_CONNECTION_ID} which are used
to share new valid connection IDs with the peer.

The format of a QUIC \mono{STREAM} frame as described in RFC~9000 can
be seen in~\cref{lst:stream-frame-format}.

\begin{listing}[htbp]
    \centering
    \begin{BVerbatim}
STREAM Frame {
  Type (i) = 0x08..0x0f,
  Stream ID (i),
  [Offset (i)],
  [Length (i)],
  Stream Data (..),
}
\end{BVerbatim}
\caption[Format of QUIC stream frame]{Format of a QUIC stream frame~\cite{rfc9000}.}
    \label{lst:stream-frame-format}
\end{listing}

\subsection{Connection ID}\label{sec:quic-conn-id}
In order to uniquely identify a connection QUIC uses connection IDs
instead of a four-tuple like TCP. Each endpoint generates a list of
valid IDs which it shares with the peer. The peer can then use one
of these IDs as a destination when transferring data~\cite{rfc9000}.

Since Connection IDs are needed for routing packets they are
carried in the clear text part of both the short and the long QUIC header.
The long header carries a source connection ID in addition to the mandatory
destination ID.

In order to prevent observers from being able to link a connection
through its connection IDs, QUIC connections rotate the IDs used.
A connection can switch the destination connection ID it is using
to one of the approved IDs previously shared by the peer at any time.

Connection IDs make addressing in QUIC independent of addressing in lower layer
protocols~\cite{rfc9000} which allows QUIC to survive
changes in the underlying four-tuple and therefore enables
connection migration.

\subsection{Multistreaming in QUIC}
SCTP was the first transport layer
protocol to feature multistreaming, but it is also a major feature
of QUIC. This means that QUIC connections can benefit from
CWND re-use and avoid head-of-line blocking as described
in \cref{sec:background-multistreaming}. This is one of the key
reasons why QUIC is the transport protocol used in HTTP/3~\cite{rfc9114}.

\subsection{Connection Migration}\label{sec:connection-migration}
QUIC connections can survive changes to the underlying four-tuple
because they are identified by connection IDs. This enables connection
migration, where an established connection is moved to a new network path,
for example switching between Wi-Fi and cellular, without requiring a new handshake.
In the current version of QUIC, only clients are able to migrate~\cite[Sec.~1]{rfc9000}.
Servers can state a preferred local IP address and request that the client
migrate by using the \mono{preferred_address} transport parameter,
but this can only be done during the initial TLS handshake~\cite[Sec.~9.6.1]{rfc9000}.

If the server does not support migration, such as when it is behind a simple load balancer which multiplexes
by source IP instead of the QUIC connection ID, then the server can mandate
that the client does not change local address by setting the \mono{disable_active_migration}
transport parameter~\cite[Sec.~5.2.3]{rfc9000}.

In order to initiate path migration, an endpoint will send a non-probing packet 
over a new network path. Non-probing packets are any packets which contain any
frame that is not \mono{PATH_CHALLENGE}, \mono{PATH_RESPONSE}, \mono{NEW_CONNECTION_ID}, or
\mono{PADDING}. When an endpoint receives non-probing packets from a new address, it can assume 
that the peer has migrated to that address. Both peers are required to use a new connection
IDs when performing migration, in order to ensure that the connection cannot
be used to link packets from the same connection over different network paths
\cite[Sec.~9.5]{rfc9000}.

If the receiving endpoint permits connection migration,
then any new packets must be sent to the new address
and path validation must be initiated.
The endpoint initiating connection migration may defer path validation until after
the receiving endpoint has sent non-probing data to the initiating endpoint's new
address~\cite[Sec.~9.2]{rfc9000}.
Path validation is used to verify
that a peer is reachable on a given path and it is done by using \mono{PATH_CHALLENGE} and \mono{PATH_RESPONSE} 
frames. The challenge frame contains \qty{8}{\Bytes} of arbitrary data,
which the peer must echo back in the response frame. This ensures that the source address
is not spoofed and therefore avoids potential amplification attacks.

If path validation fails, which happens if no matching response
is received within the Probe Timeout (PTO)~\cite[Sec. 6.2]{rfc9002},
the endpoints can continue communicating using other known paths.
If no viable paths remain, then the connection fails. This can be signaled to the peer 
using a \mono{NO_VIABLE_PATH} connection error, although the only
situation where this can be communicated is
if a path exists, but it does not support the required
maximum transmission unit (MTU)~\cite[Sec.~8.2.4]{rfc9000}.

A figure comparing QUIC connection migration with TCP can be seen in~\cref{fig:connectionBeforeAfterMigration}.
Here, a local address becomes unavailable and the TCP connection is therefore broken
since the underlying four-tuple has changed. The
QUIC connection is able to survive since the connection
ID is recognized by the peer.
When performing the migration, a new connection ID is picked
from the pool of valid IDs generated by the peer to
avoid linkability.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  >=Stealth,
  field/.style={draw, rectangle, minimum width=2.6cm, minimum height=0.6cm,
                font=\small, fill=white},
  highlight/.style={field, fill=green!15},
  dead/.style={field, fill=red!10, text=red!60},
  lbl/.style={font=\footnotesize\itshape, gray},
  title/.style={font=\normalsize\bfseries},
  every node/.style={align=center}
]
 
% ── Column titles ─────────────────────────────────────────────────────────────
\node[title] (tcptitle) at (1.3, 0.5)  {TCP};
\node[title] (quictitle) at (8, 0.5) {QUIC};
 
\draw[gray!40] (4.2, 1.0) -- (4.2, -8.5);
 
% ═══════════════════════════════════════════════════════════════════════════════
% TCP side
% ═══════════════════════════════════════════════════════════════════════════════
\node[lbl] at (1.3, 0.0) {Connection = 4-tuple};
 
% Before
\node[lbl] at (1.3,-0.5) {before migration};
\node[field] (t1) at (1.3,-1.05) {src IP:\;192.0.2.0};
\node[field] (t2) at (1.3,-1.75) {src port:\;54321};
\node[field] (t3) at (1.3,-2.45) {dst IP:\;198.51.100.0};
\node[field] (t4) at (1.3,-3.15) {dst port:\;443};
 
% Arrow
\node at (1.3,-3.7) {$\downarrow$ IP change};
 
% After
\node[lbl] at (1.3,-4.15) {after migration};
\node[dead] at (1.3,-4.7)  {src IP:\;192.0.2.1\;{\normalfont\bfseries\textcolor{red!70}{changed}}};
\node[field] at (1.3,-5.4)  {src port:\;54321};
\node[field] at (1.3,-6.1)  {dst IP:\;198.51.100.0};
\node[field] at (1.3,-6.8)  {dst port:\;443};
 
\node[draw=red!60, fill=red!10, rounded corners, font=\small, text=red!70]
  at (1.3,-7.6) {Connection broken \xmark};
 
% ═══════════════════════════════════════════════════════════════════════════════
% QUIC side
% ═══════════════════════════════════════════════════════════════════════════════
\node[lbl] at (8, 0.0) {Connection = Connection ID};
 
% Before
\node[lbl] at (8,-0.5) {before migration};
\node[highlight] at (8,-1.05) {dest. ID\;0xABCD1234};
\node[field]     at (8,-1.75) {src IP:\;192.0.2.0};
\node[field]     at (8,-2.45) {src port:\;54321};
\node[field]     at (8,-3.15) {dst IP:\;198.51.100.0};
\node[field]     at (8,-3.85) {dst port\;443};
 
% Arrow
\node at (8,-4.4) {$\downarrow$ IP change};
 
% After
\node[lbl] at (8,-4.85) {after migration};
\node[highlight] at (8,-5.55)  {dest. ID\;0x9F3E7A21\;{\normalfont\bfseries\textcolor{green!50!black}{recognized by peer}}};
\node[dead]      at (8,-6.25)  {src IP:\;192.0.2.1\;{\normalfont\bfseries\textcolor{red!70}{changed}}};
\node[field]     at (8,-6.95)  {src port:\;54321};
\node[field]     at (8,-7.65)  {dst IP:\;198.51.100.0};
\node[field]     at (8,-8.35)  {dst port:\;443};
 
\node[draw=green!60!black, fill=green!15, rounded corners, font=\small,
      text=green!50!black] at (8,-9.05) {Connection survives \cmark};
 
\end{tikzpicture}
\caption[TCP four-tuple vs. QUIC connection ID after path event]
{
Illustration of how a TCP connection fails when a parameter in the underlying
four-tuple changes, but a QUIC connection can survive since it relies on Connection
IDs.
}
\label{fig:connectionBeforeAfterMigration}
\end{figure}

In addition to supporting multiple paths through connection migration,
support for using multiple paths simultaneously through a single QUIC 
connection is currently actively being
standardized through the QUIC multipath extension draft~\cite{ietf-quic-multipath-21}.
Multipath extension builds on the mechanisms of connection migration,
but introduces a new path ID in order to manage multiple paths.

The QUIC multipath extension draft does not specify how
alternative paths are discovered, and actual setup and
teardown of paths is handled by the application.

\subsection{Session Resumption}\label{sec:back-session-resumption}
Like TLS, QUIC supports session resumption and 0-rtt
initial data. This requires that the server and the client
have already exchanged cryptographic keys, that the
client remembers critical parameters such as the selected
application-layer protocol negotiation (ALPN)
value and that the client provides the server with a
session ticket for the server to recover the same
information~\cite[Sec.~4.6]{rfc9001}.

\section{Transport Services}
At its core, a protocol is an abstraction which provides a set of
services to an higher-layer protocol or application. In the context of
networking, these services encompass transporting data from one endpoint
to another, with the protocol itself deciding the which services are
provided. For example, TCP provides a guarantee of reliable
delivery while UDP does not.

As an example, if an application requires that the
transferred data is received as an in-order stream, then
any transport layer protocol offering this as a service to the application layer
could function as a transport protocol, and swapping the
protocols should in theory be simple. This is of course
not currently the case due to issues described in \cref{sec:motivation}.

Exactly which services are provided by which transport
protocols has been mapped in RFC~8095 by the Transport Services (TAPS) working group~\cite{rfc8095}.

\section{TAPS}
The TAPS WG was an IETF working group which aimed to solve the strong
coupling between application
level code and the underlying transport protocols through
focusing on the services provided by each transport layer protocol.
This is intended to allow faster adaptation of new protocols and to
abstract away network-layer details from the application layer.

The main RFCs produced by the TAPS WG are RFC~9621~\cite{rfc9621},
RFC~9622~\cite{rfc9622} and RFC~9623~\cite{rfc9623}. Together, these
describe a TAPS System, which is an implementation of of the TAPS
API and the TAPS Implementation. The TAPS API is the API of a
TAPS System which offers a generic way to interact with the network
independent of the underlying protocol. The TAPS Implementation 
is the internal workings of a TAPS System. This includes candidate
gathering, candidate racing, as well as the interaction
to the supported transport protocols.

Despite being implemented in user-space, a TAPS System conceptually resides
in the transport layer. The TAPS API provides convenience to the application
layer by expanding the responsibility of the transport layer.
This minimizes how much the application layer has to adapt to
the underlying transport layer, and reduces the coupling between the two layers.

The difference in API models for socket and for TAPS can be
seen in~\cref{fig:socket-programming-api-model} and
\cref{fig:taps-api-model}.

\begin{figure}[htbp]
    \centering
    \begin{BVerbatim}
+-----------------------------------------------------+
|                    Application                      |
+-----------------------------------------------------+
        |                 |                  |
  +------------+     +------------+    +--------------+
  |  DNS Stub  |     | Stream API |    | Datagram API |
  |  Resolver  |     +------------+    +--------------+
  +------------+          |                  |
                    +---------------------------------+
                    |    TCP                UDP       |
                    |    Kernel Networking Stack      |
                    +---------------------------------+
                                    |
+-----------------------------------------------------+
|               Network-Layer Interface               |
+-----------------------------------------------------+
    \end{BVerbatim}
    \caption[Socket programming API model]{API model for socket programming~\cite{rfc9622}.}
    \label{fig:socket-programming-api-model}
\end{figure}

\begin{figure}[htbp]
    \centering
    \begin{BVerbatim}
+-----------------------------------------------------+
|                    Application                      |
+-----------------------------------------------------+
                          |
+-----------------------------------------------------+
|              Transport Services API                 |
+-----------------------------------------------------+
                          |
+-----------------------------------------------------+
|          Transport Services Implementation          |
|  (Using DNS, UDP, TCP, SCTP, DCCP, TLS, QUIC, etc.) |
+-----------------------------------------------------+
                          |
+-----------------------------------------------------+
|               Network-Layer Interface               |
+-----------------------------------------------------+
    \end{BVerbatim}
    \caption[TAPS API model]{API model for TAPS~\cite{rfc9621}.}
    \label{fig:taps-api-model}
\end{figure}

\subsection{TAPS Architecture} \label{sec:tapsArch}
RFC~9621 describes the required architecture for a TAPS System,
and a diagram can be seen in~\cref{fig:tapsArch}.
RFC~9621 also describes overarching requirements of a TAPS System and the
key concepts for both the TAPS Implementation and the TAPS API. Examples
of these include the basic objects used to interact with TAPS such as
the Connection and Listener, as well as candidate racing which forms
the basis of how a TAPS Implementation initiates a Connection.

\begin{figure}[htbp]
    \centering
    \begin{BVerbatim}
 +-----------------------------------------------------+
 |                    Application                      |
 +-+----------------+------^-------+--------^----------+
   |                |      |       |        |
 pre-               |     data     |      events
 establishment      |   transfer   |        |
   |        establishment  |   termination  |
   |                |      |       |        |
   |             +--v------v-------v+       |
 +-v-------------+   Connection(s)  +-------+----------+
 |  Transport    +--------+---------+                  |
 |  Services              |                            |
 |  API                   |  +-------------+           |
 +------------------------+--+  Framer(s)  |-----------+
                          |  +-------------+
 +------------------------|----------------------------+
 |  Transport             |                            |
 |  Services              |        +-----------------+ |
 |  Implementation        |        |     Cached      | |
 |                        |        |      State      | |
 |  (Candidate Gathering) |        +-----------------+ |
 |                        |                            |
 |  (Candidate Racing)    |        +-----------------+ |
 |                        |        |     System      | |
 |                        |        |     Policy      | |
 |             +----------v-----+  +-----------------+ |
 |             |    Protocol    |                      |
 +-------------+    Stack(s)    +----------------------+
               +-------+--------+
                       V
+-----------------------------------------------------+
|               Network-Layer Interface               |
+-----------------------------------------------------+
    \end{BVerbatim}
    \caption[TAPS System architecture]{TAPS System architecture as described in RFC~9621~\cite{rfc9621}.}
    \label{fig:tapsArch}
\end{figure}

\subsection{TAPS API}
RFC~9622 describes the TAPS API. This is an abstract API designed to be
able to interact with the network while enabling dynamic selection of
transport protocols and network paths~\cite{rfc9622}.
The TAPS API is object-oriented and is primarily interacted
with through the \term{Preconnection}, \term{Connection},
\term{Message}, \term{Listener}, \term{Local Endpoint}, \term{Remote Endpoint},
\term{Transport Properties} and \term{Security Parameters} objects.

A Connection represents a \enquote{shared state of two or more endpoints that persists
across Messages that are transmitted between these endpoints}~\cite{rfc8303}.

When working with Connection objects, TAPS provides five high
level groups of actions that an application can perform:
\begin{itemize}
    \item Preestablishment: Pass properties to specify
        requirements and preferences for establishing a Connection.
    \item Establishment: Prepare a Connection object for data
        transfer.
    \item Data transfer: Sending and receiving Messages
        over the Connection.
    \item Event handling: Notifications that an application
        can receive from a Connection, and the handling of them.
    \item Termination: How transmission is stopped and
        a Connection is torn down.
\end{itemize}

In addition, the lifeline of a Connection is divided into three distinct
phases: preestablishment, the established state, and
termination of a Connection~\cite{rfc9621}.
Each of these phases are associated
with one or more of the groups of actions that an application can perform,
as described above. 

After the desired Transport Properties and Security Parameters have
been set during the preestablishment phase, the Connection can be established. 
After a Connection has been established it is in the established phase and can be used to
transfer data. 

A Connection can be established in one of three ways, \mono{initiate()}, \mono{listen()} or 
\mono{rendezvous()}~\cite{rfc9621}.

Invoking \mono{initiate()} establishes a Connection to a remote endpoint which
is assumed to be listening for incoming Connections. How this Connection
is established internally is dependent on the underlying transport
protocol. The Created Connection object is passed to the application through a callback.
Exactly how the underlying transport protocol is chosen is described in detail
in \cref{sec:candidate-gathering} and \cref{candidateRacing}.

Invoking \mono{listen()} is used to passively open a Connection. Invoking \mono{listen()}
creates a Listener object and wait for incoming Connections.
When a client connects to this Listener, a \term{Received}
event occurs. This event contains a Connection object for the 
established Connection.

Invoking \mono{rendezvous()} is used to establish a peer-to-peer
Connection between two endpoints simultaneously.

After a Connection has been established, it can either
be closed or aborted, with closing being associated with graceful teardown
and aborting being abnormal termination associated with errors.

\subsection{Cloning}
One or more TAPS Connections can be organized into a \term{Connection Group}.
The Connection Group is not an object exposed through the TAPS API, but
is maintained internally in a TAPS Implementation.
A Connection Group is used to group related Connections together, which have been created
by cloning. Connections in the same Connection Group can be
multiplexed together if this is supported by the underlying protocol.

\subsection{Framer}
In TAPS a \term{Framer} is an object responsible for bridging the TAPS concept
of a Message, with the underlying protocol concept of either a data stream
or a datagram. In its simplest form, a Framer is responsible for encoding
or decoding outgoing and incoming data, as well as interacting with setup and
teardown of a Connection if necessary~\cite[Sec.~9.1.2]{rfc9622}. Because Framers may need to send
data upon the creation of a Connection, they must be added to a Preconnection
during the preestablishment phase.
The application can define custom framers and pass them to a Preconnection
by implementing an interface provided by the TAPS System.

Because they may need to interact with the network, 
framers are also asynchronous and need
to notify their associated Connection when the processing of a given Message
has completed.

The concept of a framing protocol also exists outside of TAPS since most
application-level protocols are both message-based and run over TCP. Several
ways of converting TCP's byte-stream into usable messages has been implemented.
These largely fall into two categories, type-length-value (TLV) structures where
the length of the following data is described in a header-structure, and
delimiter-separated formats, where the beginning and end of a piece of
data is separated by delimiters~\cite[Sec.~6.1]{rfc9623}.
Examples of the two categories are HTTP/1.1 for a delimiter-separated format
and DHCP for TLV~\cite{rfc2131}.

Framers become useful in TAPS because they can be used to ensure Messages
are received in a consistent format no matter what underlying transport
protocol is selected. Framers can be used both when sending and receiving,
and can both split and join Messages.

Conceptually, a framer resides on a separate layer between a Connection and the
underlying Transport Protocol Stack as seen in~\cref{fig:tpsStack}.

\begin{figure}[htbp]
\centering
\begin{BVerbatim}
 Initiate()   Send()   Receive()   Close()
     |          |         ^          |
     |          |         |          |
+----v----------v---------+----------v-----+
|                Connection                |
+----+----------+---------^----------+-----+
     |          |         |          |
     |      +-----------------+      |
     |      |    Messages     |      |
     |      +-----------------+      |
     |          |         |          |
+----v----------v---------+----------v-----+
|                Framer(s)                 |
+----+----------+---------^----------+-----+
     |          |         |          |
     |      +-----------------+      |
     |      |   Byte-stream   |      |
     |      +-----------------+      |
     |          |         |          |
+----v----------v---------+----------v-----+
|         Transport Protocol Stack         |
+------------------------------------------+
  \end{BVerbatim}
  \caption[Protocol stack with framer]{Diagram of a protocol stack featuring a Framer~\cite{rfc9622}.}\label{fig:tpsStack}
\end{figure}

\subsection{Transport Properties}
Transport Properties are properties used to influence the underlying transport protocol.
They can be divided into two categories, \term{Selection Properties}
which influence which underlying transport protocol is selected and
\term{Connection Properties} which influence the behavior of the underlying
protocol.


\subsubsection{Selection Properties}
Selection properties are the subset of Transport Properties which are
primarily used to filter and sort candidate protocols
when performing candidate gathering as described
in~\cref{sec:candidate-gathering}. Certain Selection Properties
can also be used to influence the Local and Remote Endpoints used.

Typical Selection Properties are specified as
a \term{Preference}, with the possible values being \term{Require},
\term{Prefer}, \term{No Preference}, \term{Avoid} and \term{Prohibit}.
Require means that the Connection establishment will fail if the property cannot
be satisfied, while Prefer means that the TAPS Implementation will prioritize
protocols which can satisfy the given property when performing candidate racing.
This distinction is mirrored in the difference between Prohibit and Avoid.
Not all Selection Properties are a Preference,
and they can also be specified as a boolean, an integer, an enum or other types.

An example of a Selection Property is \mono{reliability}.
This property is used to specify whether or not the application
should use a protocol which ensures that all data is received in order,
without loss or duplication~\cite[Sec.~6.2.1]{rfc9621}. Setting \mono{reliability}
to Require will force TAPS to use an underlying protocol
which handles loss such as TCP or QUIC. Setting this as
Prohibit will force TAPS to use an unreliable protocol
such as UDP.

\subsubsection{Connection Properties}
Connection Properties are the subset of Transport Properties which
are used to control the behavior of the underlying protocol after
a Connection has been established.

An example of a Connection Property is \mono{connTimeout}.
This property is used to specify the how long to wait before deciding that an
active Connection has failed when trying to reliably deliver data to the
Remote Endpoint~\cite[Sec.~8.1.3]{rfc9622}

\subsection{Message Properties}
Message Properties are settings set on an
individual Message. These are used to set the behavior
of a single Message when transferring data, but do
not affect any other Messages transferred by the Connection.

In certain cases these can override the Transport Properties
of the Connection, for example by setting \mono{msgOrdered}
to indicate that this individual message does not have to be
ordered~\cite[Sec.~9.1.3.4]{rfc9622}.

\subsection{TAPS Implementation}
While the TAPS RFCs does not prescribe implementation specific details,
RFC~9623 serves as a guide in how to best develop a TAPS System.
This includes how to present actions which are present in the TAPS API
such as data transfer as well as provided services which are not
directly exposed such as candidate gathering or candidate racing.

\subsection{Candidate Gathering}\label{sec:candidate-gathering}
When the application invokes \mono{initiate()} on a Preconnection,
a TAPS System will gather a list of candidates for establishing a
Connection. This process is known as candidate gathering. The candidates are based on
the specified Local and Remote Endpoints of the Preconnection,
as well as the given Selection Properties which are available to use. 

RFC~9623 conceptualizes these combinations as a tree,
with the first level of child nodes representing the
possible Local Endpoints, which are resolved to local socket addresses
(each a specific IP address and port) when building the tree~\cite{rfc9623}.
The second level of the candidate tree represents the possible protocol
options which can be used. This includes not only the protocol
itself, but also any combination of options for this protocol which
influence Connection establishment.

The selected local interface can be restricted through 
either using specific local endpoints or through the
\mono{interface} Selection Property. The viable
remote socket addresses can be influenced by for
example the \mono{useTemporaryLocalAddress} Selection Property,
as this is only supported in IPv6 \cite[Sec.~6.2.13]{rfc9622}.

RFC~9623 does not mandate that implementations follow the same tree
structure when performing candidate gathering, but it does recommend to branch in the
same order as the RFC, in order to branch by the most important concepts
first~\cite[Sec.~4.1.2]{rfc9623}.

After branching, incomplete paths and leaf nodes
not compatible with the specified Selection Properties are
pruned. After pruning, each leaf node represents a single possible
Connection attempt, containing a combination of local socket address, 
protocol options and remote socket address. These results are
sorted and candidate racing can be initiated.
A diagram of the tree used in candidate gathering can be seen in
\cref{fig:candidate-tree}.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  grow=right,
  level 1/.style={sibling distance=48mm, level distance=22mm},
  level 2/.style={sibling distance=24mm, level distance=28mm},
  level 3/.style={sibling distance=12mm, level distance=34mm},
  every node/.style={font=\small},
  edge from parent/.style={draw, -latex},
  pruned/.style={gray, dashed},
  candidate/.style={draw, rounded corners, fill=green!15, inner sep=3pt},
  pruned node/.style={gray},
]

\node {Root}
  child { node {eth0}
    child { node {QUIC}
      child { node[candidate] {\texttt{1.2.3.4}} }
      child { node[candidate] {\texttt{5.6.7.8}} }
    }
    child { node[pruned node] {\sout{TCP}}
      child { node[pruned node] {\texttt{1.2.3.4}} edge from parent[pruned] }
      child { node[pruned node] {\texttt{5.6.7.8}} edge from parent[pruned] }
    }
  }
  child { node {wlan0}
    child { node {QUIC}
      child { node[candidate] {\texttt{1.2.3.4}} }
      child { node[candidate] {\texttt{5.6.7.8}} }
    }
    child { node[pruned node] {\sout{TCP}}
      child { node[pruned node] {\texttt{1.2.3.4}} edge from parent[pruned] }
      child { node[pruned node] {\texttt{5.6.7.8}} edge from parent[pruned] }
    }
  };

% Level labels above tree — tweak y-offset to clear the topmost node
% \node[anchor=south] at (0,    4.6) {\textit{Precon.}};
\node[anchor=south] at (2.2,  4.6) {\textit{Resolved Local}};
\node[anchor=south] at (5.0,  4.6) {\textit{Protocol}};
\node[anchor=south] at (8.4,  4.6) {\textit{Resolved Remote}};

\end{tikzpicture}
\caption[Candidate gathering tree]{
An example of a candidate gathering tree for two local interfaces, two protocols and two
remote addresses. Grey dashed branches are pruned based on the Selection Properties
set by the application.
}
\label{fig:candidate-tree}
\end{figure}



\subsection{Candidate Racing}\label{candidateRacing}
Candidate racing is the process of dynamically attempting
to establish Connections over multiple candidates, in order
to find out which one is best suited to the network conditions
and the requirements of the application.

RFC~9623 describes three primary ways of performing candidate
racing:

\begin{enumerate}
    \item Simultaneous
    \item Staggered
    \item Failover
\end{enumerate}

Simultaneous racing is the process of trying to establish
Connections with all candidates at the same time. RFC~9623
discourages this method, as it can lead to an unnecessarily large
network load.

Staggered racing is the process of trying to establish Connections
with several candidates at once, but with a delay between each attempt.
This delay is configurable, but should be based
on the expected handshake time~\cite[Sec.~4.3.2]{rfc9623}. 
The most desirable candidates are raced first, and all subsequent
attempts are canceled after the first success.
RFC~9623 recommends following the algorithm described in
RFC~8305~\cite{rfc8305}, known as Happy Eyeballs, when performing staggered racing of IP addresses.
Happy Eyeballs is a staggered racing algorithm designed to maximize IPv6 usage by
racing IPv4 and IPv6 addresses on dual-stack hosts.

RFC~8305 lists four steps for protocol racing,
which conceptually divide into candidate gathering and candidate racing 
in a TAPS Implementation: 

\begin{enumerate}
  \item Initiation of asynchronous DNS queries
  \item Sorting of resolved destination addresses
  \item Initiation of asynchronous connection attempts
  \item Establishment of one connection, which cancels all other attempts
\end{enumerate}

Step one and two map into the TAPS concept of candidate gathering
while step three and four map into the TAPS concept of candidate racing.

Failover racing is the process of racing a single candidate
at a time, only attempting the next candidate if the previous
fails. This can be used when an application has a strong preference
of one branch over another.

Connection establishment has been completed when a single candidate
has successfully established a Connection, or
when all candidates have failed. 

An example flow of candidate racing can be seen in~\cref{fig:racingSeqDiagram}.
Here, DNS resolution resolves two endpoints with delays of
RTT of \qty{300}{\milli\second} and \qty{100}{\milli\second} respectively. The slower endpoint is attempted first,
but the faster RTT leads to the second endpoint winning the race and being
presented to the application.
In the example the DNS resolution is set to take \qty{100}{\milli\second}
which is an arbitrary value with no influence on the following candidate racing.
The stagger delay is set to be \qty{250}{\milli\second}.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  >=Stealth,
  msglbl/.style={font=\scriptsize, fill=white, inner sep=1pt},
  head/.style={draw, rectangle, minimum width=1.8cm, minimum height=0.6cm,
               font=\small, fill=white}
]

% ── x positions ─────────────────────────────────────────────────────────────
\def\xO{0}     % Application
\def\xA{2.5}     % CTaps
\def\xB{4.5}   % DNS
\def\xC{6.5}
\def\xD{8.5}
\def\xE{10.5}  % Server

% ── y positions (y = -t/100 * 1.2) ─────────────────────────────────────────

%  t=0:    y=-0.4 (Query)
%  t=50:   y=-1.0
%  t=100:  y=-1.6 (DNS resolve)
%  t=200:  y=-2.0
%  t=300:  y=-2.8
%  t=350:  y=-3.4
%  t=400:  y=-4.0
%  t=500:  y=-5.2
\tikzmath{
  \yfiftyms = -0.5;
  \ytenms = \yfiftyms / 5;
  \ylifelinestart   = 0.1;
  \yzero   = -0.4 ;
  \yfifty = \yzero + \yfiftyms; 
  \yhundred = \yzero + \yfiftyms * 2; 
  \yhundredandfifty = \yzero + \yfiftyms * 3; 
  \ytwohundred = \yzero + \yfiftyms * 4; 
  \ytwohundredandfifty = \yzero + \yfiftyms * 5; 
  \ytreehundred = \yzero + \yfiftyms * 6; 
  \ytreehundredandfifty = \yzero + \yfiftyms * 7; 
  \yfourhundred = \yzero + \yfiftyms * 8; 
  \yfourhundredandfifty = \yzero + \yfiftyms * 9; 
  \yfivehundred = \yzero + \yfiftyms * 10; 
  \yfivehundredandfifty = \yzero + \yfiftyms * 11; 
  \ysixhundred = \yzero + \yfiftyms * 12; 
  \ydnsresolution = \yfifty - \ytenms;
  \yslowdelay = \yfiftyms * 3;
  \yfastdelay = \yfiftyms;
  \ylifelineend = \yfivehundredandfifty; 
  \yslowstart = \yzero + \yfiftyms;
  \yfaststart = \yslowstart + \yfiftyms * 5;
  \yslowend = \yfaststart + \yfastdelay * 4 + \ytenms * 2;
  \yone = \yslowstart + \yslowdelay * 3;
  \ytwo = \yslowstart + \yslowdelay * 4;
  \ythree = \yslowend;
  \xone = \xE;
  \xtwo = \xC;
  \slope = (\ytwo - \yone)/(\xtwo - \xone);
  \b = \yone;
  \xslowend = \xone + (\ythree - \b) / \slope;
} 



% ── Participant headers ──────────────────────────────────────────────────────
\node[head]                   at (\xO, 0.4) {Application};
\node[head]                   at (\xA, 0.4) {CTaps};
\node[head]                   at (\xB, 0.4) {DNS};
\node[head]                   at (\xC, 0.4) {Cand. 1};
\node[head, font=\small]      at (\xD, 0.4) {Cand. 2};
\node[head]                   at (\xE, 0.4) {Server};
% \node[font=\scriptsize, gray] at (\xA, -0.05) {(per-candidate)};
% \node[font=\scriptsize, gray] at (\xB, -0.05) {(per-candidate)};

% ── Lifelines ───────────────────────────────────────────────────────────────
\draw[dashed, gray!50] (\xO, \ylifelinestart) --  (\xO, \ylifelineend);
\draw[dashed, gray!50] (\xA, \ylifelinestart) --  (\xA, \ylifelineend);
\draw[dashed, gray!50] (\xB, \ylifelinestart) --  (\xB, \ylifelineend);
\draw[dashed, gray!50] (\xE, \ylifelinestart) --  (\xE, \ylifelineend);

\draw[dashed, gray!50] (\xC, \yslowstart) -- (\xC, \yslowend);
\draw[dashed, gray!50] (\xD, \yfaststart) -- (\xD, \ylifelineend);

%
% \draw[gray!20] (-0.2, -0.8) -- (12.2, -0.8);

% ── Time grid and labels ─────────────────────────────────────────────────────
\foreach \yval/\t in {
  \yzero/0,
  \yfifty/50,
  \yhundred/100,
  % \yhundredandfifty/150,
  \ytwohundred/200,
  % \ytwohundredandfifty/250,
  \ytreehundred/300,
  % \ytreehundredandfifty/350,
  \yfourhundred/400,
  % \yfourhundredandfifty/450,
  \yfivehundred/500
} {
  \node[font=\footnotesize, gray, left] at (-0.4, \yval) {\t\,ms};
  \draw[gray!20] (-0.2, \yval) -- (12.2, \yval);
}

% ── t≈0: Start Connection attempt ─────────────────────────────────────────────────────────
\draw[->] (\xO, \yzero) -- node[msglbl, above] {Connect to URL} (\xA, \yzero);

% ── t≈0: DNS queries ─────────────────────────────────────────────────────────
\draw[->] (\xA, \yzero) -- node[msglbl, above] {Query URL} (\xB, \yzero);

% ── DNS resolves A ────────────────────────────────────────────────────
\draw[->] (\xB, \ydnsresolution) -- node[msglbl, above] {Resolve {C1, C2}} (\xA, \ydnsresolution);

\draw[fill=black] (\xA,\yslowstart) circle (1.5pt);
\draw[->] (\xA, \yslowstart) -- node[msglbl, below] {Start attempt 1} (\xC, \yslowstart);

% ── QUIC handshake for candidate A ───────────────────────────────────────────
% Initial:        SM-A (t=100) → Server (t=175)
\draw[->] (\xC,\yslowstart)  -- node[msglbl, above, sloped] {}         (\xE,{\yslowstart + \yslowdelay});
% Handshake:      Server(t=175) → SM-A (t=250)
\draw[->] (\xE,{\yslowstart + \yslowdelay})  -- node[msglbl, above, sloped] {}       (\xC,{\yslowstart + \yslowdelay * 2});
\draw[->] (\xC,{\yslowstart + \yslowdelay * 2})  -- node[msglbl, above, sloped] {}       (\xE,{\yslowstart + \yslowdelay * 3});
% Finished:       SM-A (t=250) → Server (t=325)
\draw[-, red!90] (\xE,{\yslowstart + \yslowdelay * 3})
  -- node[msglbl, above, sloped] {}
     node[pos=1.0, cross out, draw, inner sep=2pt] {}
     (\xslowend,\yslowend);

% ── t=350: racing timer fires, create SM-B ───────────────────────────────────
\draw[fill=black] (\xA,\yfaststart) circle (1.5pt);
%\node[font=\scriptsize, right=3pt] at (\xA,-2.4) {250\,ms timer};
\draw[->] (\xA,\yfaststart) -- node[msglbl, above] {Start attempt 2} (\xD,\yfaststart);

% ── QUIC handshake for candidate A ───────────────────────────────────────────
\draw[->] (\xD,\yfaststart)  -- node[msglbl, above, sloped] {}         (\xE,{\yfaststart + \yfastdelay});
\draw[->] (\xE,{\yfaststart + \yfastdelay})  -- node[msglbl, above, sloped] {}       (\xD,{\yfaststart + \yfastdelay * 2});
\draw[->] (\xD,{\yfaststart + \yfastdelay * 2}) -- node[msglbl, above, sloped] {}         (\xE,{\yfaststart + \yfastdelay * 3});
\draw[->] (\xE,{\yfaststart + \yfastdelay * 3}) -- node[msglbl, above, sloped] {} (\xD,{\yfaststart + \yfastdelay * 4});
\draw[->] (\xD,{\yfaststart + \yfastdelay * 4}) -- node[msglbl, above, sloped] {}   (\xA,{\yfaststart + \yfastdelay * 4});

% ── t=400: SM-A wins ─────────────────────────────────────────────────────────
\node[draw, fill=green!15, font=\scriptsize, rounded corners, inner sep=2pt]
      at (\xD,{\ylifelineend + \ytenms * 2}) {Connected \cmark};

\node[draw, fill=red!15, font=\scriptsize, rounded corners, inner sep=2pt]
      at (\xC,{\ylifelineend + \ytenms * 2}) {Closed \xmark};

% ── t=400: abort SM-B ────────────────────────────────────────────────────────
\draw[->, red!70, thick] (\xA,\yslowend ) -- node[msglbl, below] {close()} (\xC,\yslowend);

\draw[->] (\xA, {\yslowend + \ytenms}) -- node[msglbl, above] {\mono{ready()} CB} (\xO, {\yslowend + \ytenms});

\end{tikzpicture}
\caption[Example candidate racing with slow and fast endpoint]
{Example of candidate racing functionality when the order of a DNS resolution causes a slow endpoint to be attempted first,
but the fast endpoint still wins.}
\label{fig:racingSeqDiagram}
\end{figure}

\section{ALPN}\label{sec:back-alpn}
ALPN is a feature in TLS \cite{rfc7301} where the application layer
protocol is negotiated between on
transport layer. TLS is a mandatory part of QUIC and QUIC mandates that endpoints
agree on application protocol through a 
mechanism such as ALPN~\cite[Sec. 7]{rfc9000}.

ALPN allows the client and server to negotiate a common protocol
without adding network round trips. The client puts 
a list of supported protocols in its \mono{clientHello} message
and the server chooses a single protocol and puts it in
its \mono{serverHello} message. If the server does not support any
protocols that the client supports, then it responds with an
error notifying the client of this.

\section{Asynchronous I/O and event loops} \label{sec:async-background}
The naive way to write a networking application is to treat
any interaction with the network like a normal CPU-bound task,
waiting for I/O bound operations to complete before progressing.
This is a simple, but very inefficient model and the CPU will
be spending much of its time waiting for I/O. Modern networking
applications therefore interact with the network in an asynchronous way, this
is reflected in the design of TAPS, which is asynchronous and
event-driven.

There are two primary ways of achieving asynchrony when talking to
the network. The first is multithreading,
where the application creates a thread per connection and threads yield
the processor while waiting for blocking I/O. The second is an event
loop. In its simplest form, an event loop is a single-threaded
program that executes tasks in order from a queue.
Functions that initiate asynchronous work return immediately and the
application is later notified of the result. 
Tasks are added to the queue either by the application itself or in
reaction to some asynchronous result being ready, such as incoming
data.

An illustration of an event loop with a simple queue can be seen in
\cref{fig:event-loop-queue}. Here, epoll is used to monitor non-blocking
sockets and the associated receive-callbacks are added to the queue.
This allows a single-threaded event loop to handle
several different connections at once.

\begin{figure}[htpb]
\centering
\begin{tikzpicture}[
    >=Stealth,
    thick,
    bigbox/.style={
        draw, rectangle,
        minimum width=2cm,
        minimum height=3.2cm,
        align=center,
        font=\small,
        rounded corners=2pt,
        fill=gray!10
    },
    task/.style={
        draw, rectangle,
        minimum width=3.7cm,
        minimum height=0.72cm,
        align=center,
        font=\footnotesize,
        fill=gray!10,
    },
    lbl/.style={font=\footnotesize, align=center},
    dimtask/.style={task, dashed, draw=gray!60, text=gray!70, fill=white},
]
 
%% === OS / Kernel box (left) ===
\node[bigbox] (os) at (0, 0) {
    \textbf{Epoll}
    % {\footnotesize TCP socket\\UDP socket\\Timers}
};
 
%% === Task queue (center) ===
% Top = head (next to execute), bottom = tail (enqueue here)
\node[dimtask] (p1) at (5,  0.72) {};
\node[dimtask](p4) at (5, -0.72) {};
\node[task, rotate=-15] (q1) at (5.15, 1.18) {UDP recv. cb};
\node[task] (q3) at (5,  0.00) {TCP recv. cb};
\node[task, rotate=-15] (q4) at (4.84,  -1.20) {QUIC receive cb};

 
%% === Event loop box (right) ===
\node[bigbox, text width=2cm] (el) at (10, 0.36) {
    \textbf{Event loop}\\[4pt]
};
 
\draw[->]
    (os.south)
    to[out=-90, in=-105]
    node[above=0.3, lbl, align=left, text width=3cm] {I/O ready\\ $\hookrightarrow$ enqueue callback}
    (q4.south);
 
\draw[->]
    (q1.north)
    to[out=75, in=90, looseness=0.9]
    node[below=0.4, lbl, align=left, text width=3cm] {Execute next item from queue}
    (el.north);
 
\end{tikzpicture}
\caption[Event loop with FIFO queue]{
An simple FIFO event loop driven by epoll. Epoll signals the event loop when
a monitored socket becomes ready and the corresponding receive callback is
added to the queue.
The event loop executes received callbacks in order,
allowing a single thread to service multiple connections.
} 
\label{fig:event-loop-queue}
\end{figure}


The largest advantage of a threaded model is that it allows code to be written sequentially
within each thread.
This is closer to the mental model of a normal CPU-bound program, but it requires synchronization
of shared data structures between threads. An event-loop model avoids the
synchronization issue by running in a single thread, but spreads control
flow across callbacks. Sequential flow in an event-loop driven program
can be achieved through the use of \term{coroutines}. Coroutines are common
in higher level languages, but in C they must either be implemented manually
or supported through external libraries.

The multithreaded and the event loop models also differ in how they scale. Multithreading makes use of
multiple CPU cores easily, but adds a non-negligible overhead when
creating a large number of threads.
Event loops avoid the overhead of multiple threads and scale well, but they are more difficult to use with
multiple cores. The two models are not entirely mutually exclusive
as event loop libraries can use threads internally to achieve an asynchronous flow
for synchronous operations, but from the view of an application
interacting with the network, they are fairly distinct.

Although real implementations naturally feature more advanced features than
illustrated in \cref{fig:event-loop-queue}, the event loop model is very common,
and is used in Node.js~\cite{nodejsEventLoop} and Python's asyncio~\cite{pythonAsyncio},
among others.

\section{Libuv}
Libuv~\cite{Libuv2025} is a multi-platform event loop library focused on asynchronous I/O.
It is primarily developed for Node.js, but is used in several other large projects as well.
Libuv is centered the \mono{uv_loop_t} structure which represents a single event loop.
Applications can register work to a loop through the use of handles and requests~\cite{libuv_design}.
Handles are long-lived objects for certain types of work, while requests are short-lived operations.
When initial handles and requests have been registered, the application must invoke \mono{uv_run()}
in order to execute the event loop. This
will take over the invoking thread and by default run until all registered handles
have been closed.

Handles are initiated and registered to a given event loop through
\mono{uv_*_init()}. Requests can then be added as discrete
pieces of work to the given handle. For example \mono{uv_write()} is
used to add a write request to a \mono{uv_tcp_t} handle. When registering
a request with the event loop, the application must also register an
associated callback. This callback will be invoked to notify the application
that execution has completed.
When the application is finished it must close the created handles for
the event loop to stop execution.

\section{Picoquic}\label{sec:picoquic}
Picoquic~\cite{christianhuitemaPicoquic} is a C library which aims to be a
minimalist implementation of the QUIC standard. It is a cross-platform
library compatible with Windows, Linux, macOS and BSD.
Though it features an optional event loop,
Picoquic is a sans-I/O \cite{cannon2016sansio} library
and does not perform I/O itself, instead relying on the
application to interact with the network on its behalf.
This decouples protocol logic from any
particular I/O model, allowing it to be integrated with existing
I/O solutions such as libuv.

Applications are required to
call into Picoquic at given intervals. These intervals
are specified by Picoquic and are updated after key
events. A timer
can then be set to wake at the specified interval
and perform any work necessary by Picoquic.

When this timer fires, the application has to query
Picoquic for any new packets to send to the peer.
Picoquic will inform the application of about the
expected source and destination socket addresses,
which allows it to support connection migration and
multipath in QUIC.

Since the application is responsible for interacting
with the network it is also responsible for passing any
received QUIC packets to Picoquic. When QUIC data is
received from the network it can be passed to Picoquic,
and Picoquic will perform any handling related
to QUIC itself, such as generating outgoing packets to respond
to an incoming handshake.

Picoquic handles all necessary TLS setup through Picotls~\cite{picotls},
only requiring that the application
provides the necessary parameters, such as
ALPN values and the various key files.

As an alternative to driving the polling API,
simple clients which do not want to run their own I/O
can run Picoquic's built-in event loop through
\mono{picoquic_packet_loop()}. This takes
control over the invoking thread and allows the application to
control the behavior of the loop through returning values
from a callback. When using this packet loop, Picoquic
manages its own sockets and interacts with the network
independently.

\section{Glib}
GLib~\cite{GLibGitLab} is a general purpose utility library which
is used at the core of several other larger libraries, such as GTK and GNOME.

Among other features, GLib provides implementations of several key data
structures not provided by the C standard library, such as dynamic arrays,
hash maps, trees and linked lists.
