\chapter{Introduction}\label{chap:intro}

\section{Motivation}
Network ossification has lead to innovation at the
In theory, applications should be able to quickly
and easily swap transport layer protocols, in order
to make of new protocols and the features provided
by them. This is however not reflected in reality
as network middleboxes prevent a challenge to using
any protocol that does not match TCP and UDP and
ossified TCP implementations lead applicatoins
not even being able to rely on TCP features such
as ECN.

Mitigations against this network ossification does exist,
primarily through using UDP as a substrate when developing
new transport layer protocol, but this comes at the cost of increased
overhead and does not address the actual root of the issue.

Another large issue with utilizing new innovations on the
transport layer is the network effect required in order to
start utilizing new protocols. Servers will not spend
time adding support for a new transport layer protocol
if no applications are using it, and no applications
will utilize a new protcol if no servers support it.
If a fallback to TCP is always needed, then all applications
can avoid a great deal of complexity by simply utilizing
TCP.

While this limits complexity it also limits access to
new transport layer protocols such as QUIC. This means
that applications cannot make use of genuine innovations
over TCP, such as connection migration and multistreaming.
Making it easier to innovate at the transport layer
would help make these features accessible to more
applications. This would then lead to more robust
connections through connection migration and a
more efficient use of bandwidth from multistreaming.

Transport Services (TAPS) is an IETF RFCs which aims
to tackle both the issue of network ossification,
the described network effect and to create a standardized
API which reflects how applications actually use the network.
By creating a common interface which is based on the application's
intent instead of the concrete action of the transport protocol,
Transport Services allows the application to be served by any
protocol which fulfills the requirements stated by the application.
With candidate gathering and candidate racing for endpoints and
protocols built-in, it handles both the compile-time binding to
specific protocols and minimizes the required network effect
for utilizing new protocols.

For a TAPS system the transport protocol fallback is built into the libary
itself and presents no extra complexity to the application. Because
TAPS is based on intentions as well, it allows the applications to state
their ideal intentions when interactin with the network, but still have
it be compatible no matter which underlying protocol is used. This means
that an application utilizing multistreaming and connection migration
when QUIC is available does not have to change or update its interaction
with TAPS in cases where QUIC is not available and the underlying protocol
falls back to TCP.

\section{Problem Statement}
But there is no implementations of the library supposed to make is suck less...
While there are open source implementations of TAPS, such as NEAT~\cite{neatCode}
and PyTAPS~\cite{frankePythonasynciotaps2025}, none of them are feature complete
and none of them were written after the TAPS RFCs were published in 2025~\cite{rfc9622}.
There are key features in TAPS which are not featured in any current open or closed source
implemtations of TAPS. The most important of these is candidate racing \todo{is this true?}.
In addition, no previous open source implementations of TAPS feature support for QUIC as one
of the underlying protocols. Since the TAPS RFCs were finalized before the publishing of the
QUIC RFCs there is currently no guidance in RFC~9623~\cite{rfc9623} on how to map QUIC to
the TAPS API, although an open draft exists~\cite{transportServicesMappingForQuic}.

This thesis therefore functions as not only an open source networking library that
can be developed further and be used by downstream applications, but also as an exploratory work
exploring exactly how well QUIC features map onto the TAPS API and how much overhead the TAPS
API presents as a layer between the application and the transport protocol. In addition
it creates a new up-to-date implementation of TAPS, being the first open source implementation
to be developed after the finalization of the TAPS RFCs.

\section{Resarch Questions}
For the development of this thesis, two major topics were extracted from the problem statement.
There is no open source implementations of TAPS developed after the finalization of the RFCs exists, and
there is no open source implementation of TAPS with support for QUIC. These were then formulated
into two research questions which this thesis will attempt to answer.

\newcommand{\rqone}{What design and architectural considerations have to be taken when implementing TAPS, in order to maintain flexibility?}
\newcommand{\rqtwo}{How well does the TAPS API capture QUIC semantics, both in coverage and preserving expected behavior?}

\begin{enumerate}
    \item \rqone
    \item \rqtwo
\end{enumerate}

The answer to these will be the unique insight gained from developing this thesis and
can be used to display the contributions made through this piece of work.

\section{Contributions}
The contributions made by this thesis to the TAPS ecosystem is CTaps.
CTaps is an open source C implementation of transport services with support
for QUIC, TCP and UDP. While CTaps is an exploratory work, primarily
focused on finding out how QUIC maps through a TAPS api and how
to maintain a flexible system when implementing TAPS, it features
several important features from both QUIC and TAPS itself.

The most important features of CTaps are multistreaming through
the use of Connection Cloning, Connection Migration through the
use of multiple Local Endpoints and candidate racing.
Connection Cloning as it is impelmented in CTaps allows
applications to make use of multiple logical streams for
a single logical connection, if supported by the underlying
protocol. All code perfoming connection cloning is functional
no matter what the underlying protocol is, allowing the
application to make use of it at any time, gaining large
benefits when QUIC is selected as the underlying protocol.

Connection migration in CTaps functions as a proof of concept,
showing that in a production-ready TAPS implementation then connection
migration could be made use of without additional complexity for the
application.

Candidate racing allows applications to attempt the most
desirable protocol first, for example attempting to connect
with QUIC if the application knows that it would benefit
from multistreaming, but it gracefully falls back to TCP
or other supported protocols if a connection cannot be
established using the most desirable protocol. Applications
can identify features and selection properties it wants to make
use of and make connection attempts based on that, instead of
having to manually maintain a fallback hierarchy for the connections
it wants to try.

In addition to CTaps as a software library this thesis
also contributes several reflections and insights which
will be useful for future implementaions of TAPS. These
include both exactly how QUIC features map into the
TAPS API, but also special architectural considerations
which need to be taken when implementing a single common
API which needs to support multiple underlying protocols.

\subsection{Scope}
With the nature of this thesis being an exploratory work
the scope of CTaps is naturally limited. CTaps will not
be a production-ready software library, and a large focus
is to explore as much of the TAPS API as possible, exposing
out-of-scope features as no-ops. While this will help
standardize the API presented in CTaps and to a better
degree show what a full TAPS library would look like,
it also means that applications using CTaps will face
an interface which is not entirely functional, even if
presented as such in the API.

With CTaps being exploratory and focusing on implementing
as supporting as many features of QUIC and the TAPS API as
possible, another natural limitation is error handling. The
largest chunk of development has been spent developing happy
paths in order to prove their viability. While this viability
is important, it is important to note that this focus on deveoloping
as many features as possible has naturally lead to the failure paths
not being a focus on development. This means that there are likely
several errors which are currently treated as fatal which likely
could be recovered from. There are also several error paths which
would cause memory leaks, as they lack testing and the valgrind
analysis which has been offered to the happy paths.

\subsection{Source Code}
CTaps is open source and is available under the MIT license
on Github \cite{cTaps} and the Doxygen documentation is available both
as a 
Github pages \cite{ctapsDocs} static web page and as an attachment to this thesis in~\cref{chap:doxygen-appendix}.

\section{Thesis Structure}
Background, related work, deisgn, implementation, evaluation and conclusion...

\textbf{\Cref{chap:background}} lays out the necessary background knowledge
to understand the basis of TAPS and CTaps. It explains the current
problems with innovating on the transport layer, the current
state of transport protocols and the advantages of QUIC. It also
describes exactly what TAPS is and the internal structures and the
advantages of TAPS.
Finally it explains the necessary background for understanding CTaps, with
the central concepts and dependencies used in CTaps.

\textbf{\Cref{chap:related-work}} lists other TAPS implementations. It
explains their features, backgrounds and current state of development.
Finally it compares them to the current state of CTaps, and reflects on
why CTaps is still needed with the current state of the ecosystem.

\textbf{\Cref{chap:design}} explains central decisions when desigining
CTaps. This includes all design decisions whether they the external API
of CTaps or the internal architecture. It explains exactly why CTaps
has the public interface that it does and reasons around the choice
of dependencies. It reflects on exactly how to expose as many QUIC features
as possible through a TAPS API and the tradeoffs which ahve to be made
by TAPS implementations here. It also reflects on decisions which
have been made which have caused a weakness in CTaps, before finally
discussing security in a TAPS system.


\textbf{\Cref{chap:impl}} explains how the goals and features of CTaps
actually were implemented internally in CTaps. It describes the various
features which are available thorugh the CTaps API as well as their biggest
advantages. In addition it describes workarounds which have been made due to
limitations in scope, and why the fact that the features are still available
can function as a proof of concept and is valuable for this thesis.

\textbf{\Cref{chap:eval}} contains a series of benchmarks performed to highlight
the various parts of CTaps. It compares CTaps with TCP and Picoquic clients, discussing
both the lack of overhead when using CTaps as an abstraction library, but also the
advantages in terms of featues and simplicity when utilizing CTaps over
direct protocol implementations.

\textbf{\Cref{chap:conc}} reflects on the final results of the thesis.
It summarizes the conclusions that this thesis has made and formulates
an answer to the research questions posed in this chapter. It lists
the weaknesses of CTaps as they exist currently and reflects on the
work which should be preformed if CTaps should be made into a
production-ready library. It discusses the internal architecture
of CTaps and how locked it is into its dependencies, before
finally reflecting on the broader significance of CTaps as a whole.
