\chapter{Background} \label{chap:background}
\section{What are network sockets}\todo{Headers should be removed after editing is finished}
Anyone programming any kind of networking application in a UNIX
environment are bound to come across sockets as the dominant
API. Sockets have been around, in more or less their current
form, since the 1980s
\cite[p.~xx]{stevensUNIXNetworkProgramming2013}. The initial
mention of sockets in relation to ARPANET comes from RFC 33,
published in 1970 \cite{crockerNewHostHostProtocol1970}.
Sockets have proven to be immensely popular, and have grown
to become the de facto standard for low-level network programming 
\cite[p. 502]{tanenbaumComputerNetworks2011}.

Network sockets have grown alongside with the growth of UNIX.
Network sockets have grown out of the Berkeley sockets which
first appeared in version 4.2 of the Berkeley Software
Distribution, originally used for inter-process communication.

The most basic definition of a socket is that it is an abstraction
for a network connection, through which an application can send
and receive data \cite[p. 8]{donahooTCPIPSockets2009}. When creating
a socket, the application needs to specify the type, domain and
protocol which will be used. The domain specifies the type of
addressing which will be used, for example \texttt{AF\_UNIX}
for local sockets, \texttt{AF\_INET} for IPv4 and \texttt{AF\_INET6}
for IPv6. The type specifies the protocol which will be used over the socket.
The two most frequent types here are \texttt{SOCK\_DGRAM} and \texttt{SOCK\_STREAM}.
Specifying \texttt{SOCK\_DGRAM} means that the socket will be using 
datagrams, meaning unreliable discrete messages.
Specifying \texttt{SOCK\_STREAM}
means that the socket will be providing a reliable stream of bytes.
The protocol parameter is used if there are multiple protocols
which support the requirements given by the domain and type.

The fact that the application has to specify both the domain
and the type of the socket at compile time means that both the
underlying addressing and the protocol are bound at compile time.
This means, that barring fairly advanced logic when setting up
the necessary sockets, the application will be bound to a specific
protocol and addressing scheme. This means that any update in
the underlying protocol will require a recompilation and 
redistribution of the application. 

This compile-time binding may not seem like a large problem, as
the type of protocol is usually deeply bound
to the application, and large redesigns likely have to take
place if an application were to change from using a
connection-oriented to a stream-oriented connection. However,
different parts of the application may have different
requirements, for example a connection to a video-streaming
website itself may be connection-oriented, while the
video-stream itself will likely be stream-oriented.

One of the key features of TAPS is to give the developer access
to more advanced features of protocols such as QUIC or SCTP,
while still being able to fall back to TCP if necessary. This
will allow TAPS to act in a message-oriented way, while still
being able to use a connection-oriented protocol such as TCP.

\section{Contemporary protocols}
The most common protocol today by far is TCP. This is a
well known protocol and will not be described in detail in
this thesis. In summary, it is a connection-oriented protocol,
which means that a connection has to be established before 
any data can be sent. This connection along with the underlying
packet loss detection allows it to
guarantee reliable delivery of packets. 
This additional
functionality comes at the cost of
increased complexity and overhead.

With modern socket programming, the most frequent protocols
are UDP and TCP. UDP is a connectionless protocol, and was
designed originally as a transport layer protocol on top of
IP, with as little overhead as possible \cite{postelUserDatagramProtocol1980}.
UDP is an unreliable protocol, menaing that it does not guarantee
delivery of sent packets. However, reliability algorithms can
be built into protocols built on top of UDP, which is done in
for example QUIC. 

When building new internet protocols, they are commonly built
on top of UDP. This is due to several reasons. Primarily it
is a result of network ossification. This is the process
of middleboxes and other network infrastructure becoming
reliant on the parts of certain parts of the protocol stack.
A middlebox may for example choose to drop an IP packet if
the transport layer protocol is not TCP or UDP. This makes
development of new protocols difficult, as they may in certain
scenarios become complete unusable due to underlying networking
hardware. Luckily, UDP is a simple protocol which contains
very little overhead. This allows future protocols to bypass
the issue of network ossification by using UDP as the underlying
protocol. This does however mean, that if the protocol needs
mechanisms present in TCP, such as reliability, then these have
to be implemented by the new protocol, as done in QUIC.

Building new protocols on top of UDP has become so common that
it has prompted an official RFC from the IETF, describing
the best practices \cite{eggertUnicastUDPUsage2008}. \todo{Forumulate
    this better}


\section{What is TAPS}
Transport Services (TAPS) is a \say{application Programming Interface 
(API) to the transport layer that enables the selection of 
transport protocols and network paths dynamically at runtime}
\cite{trammellAbstractApplicationProgramming2025}. TAPS is
designed to remove the compile-time binding to transport layer
protocols, enabling updates to underlying transport protocols
without having to rewrite application logic. TAPS does not
require that both endpoints of a connection use TAPS, and TAPS
can communicate with any endpoint which supports the underlying
transport protocol being used.

The following figure explains the API model when using
sockets: \todo{Should these be here? If yes, recreate in tikz}

\begin{figure}[H]
    \centering
    \begin{lstlisting}
        +-----------------------------------------------------+
        |                    Application                      |
        +-----------------------------------------------------+
                |                 |                  |
          +------------+     +------------+    +--------------+
          |  DNS Stub  |     | Stream API |    | Datagram API |
          |  Resolver  |     +------------+    +--------------+
          +------------+          |                  |
                            +---------------------------------+
                            |    TCP                UDP       |
                            |    Kernel Networking Stack      |
                            +---------------------------------+
                                            |
        +-----------------------------------------------------+
        |               Network-Layer Interface               |
        +-----------------------------------------------------+
    \end{lstlisting}
    \caption{API model for socket programming \cite{trammellAbstractApplicationProgramming2025}}
\end{figure}

The following figure explains the API model when using TAPS:

\begin{figure}[H]
    \centering
    \begin{lstlisting}
        +-----------------------------------------------------+
        |                    Application                      |
        +-----------------------------------------------------+
                                  |
        +-----------------------------------------------------+
        |              Transport Services API                 |
        +-----------------------------------------------------+
                                  |
        +-----------------------------------------------------+
        |          Transport Services Implementation          |
        |  (Using DNS, UDP, TCP, SCTP, DCCP, TLS, QUIC, etc.) |
        +-----------------------------------------------------+
                                  |
        +-----------------------------------------------------+
        |               Network-Layer Interface               |
        +-----------------------------------------------------+
    \end{lstlisting}
    \caption{API model for TAPS \cite{trammellAbstractApplicationProgramming2025}}
\end{figure}

Abstracting away the underlying transport protocols is done by
presenting a set of abstractions to the application. The most
central of these abstractions is the connection. A connection
represents one or more transport protocol instances that can
send or receive messages between local and remote endpoints 
\cite{trammellAbstractApplicationProgramming2025}.
When working with connection objects, TAPS provides five high
level groups of actions that an application can perform:
\begin{itemize}
    \item \texttt{Preestablishment}: Pass properties to specify
        requirements and preferences for establishing a connection.
    \item \texttt{Establishment}: Prepare connection object for data
        transfer.
    \item \texttt{Data transfer}: Sending and receiving messages
        over the connection.
    \item \texttt{Event handling}: Notifications that an application
        can receive from a connection, and the handling of them.
    \item \texttt{Termination}: How transmission is stopped and
        a connection is torn down.
\end{itemize}

In addition, the lifeline of a connection is divided into three distinct
phases, \texttt{preestablishment}, the \texttt{Established state}, and 
\texttt{termination of a Connection} \cite{paulyArchitectureRequirementsTransport2025}.
Each of these phases are associated
with one or more of the groups of actions that an application can perform
as described above. Establishment ends the preestablishment phase. When a
connection is in the established state, the connection can be used to
transfer data. When a connection has been established, it can either
be closed or aborted, with closing being associated with normal termination,
and aborting being abnormal termination associated with errors.

After the desired transport properties and cryptographic parameters have
been set, a connection can be established. There are three ways of
establishing a connection, \texttt{initiate}, \texttt{listen} and
\texttt{rendezvous} \cite{paulyArchitectureRequirementsTransport2025}.

Initiate is used to establish a connection to a remote endpoint which
is assumed to be listening for incoming connections. How this connection
is established under the hood, is dependent on the underlying transport
protocol selected based on the specified transport properties.

Listen is used to passively open a connection. This is done in order to
wait for connections from remote endpoints. Listen returns a listener
object. When a remote endpoint connects to this listener, a \texttt{ConnectionReceived}
event occurs. This event contains a Connection object for the 
established connection.

Rendezvous is used to establish a peer-to-peer connection between two endpoints
simultaneously.

4283


If a single application has multiple connections, these can be organized
into connection groups. Connection groups represent a set of connections
which shared connection properties and state generated by the underlying
transport protocols. Connections in the same connection group can be
multiplexed together, in order to more efficiently utilize the available
bandwith.

In addition to the connection, TAPS also provides several other objects
abstracting key elements. For example endpoint objects and message objects
among others.

When it comes to TAPS selecting the underlying transport protocol to
be used, this is done by
letting the application specify a set of requirements
and preferences for the network behavior. This set of requirements
and preferences are known as Transport Properties. These transport
properties can again be divided into three categories: selection
properties, connection properties and message properties. 

Selection
properties are properties used to select the path and protocol stack
to use for a given connection. Connection properties are used
during connection establishment and to define the behavior of the
established connection. Message properties are settings set on an
individual message. Typical transport properties are specified as
a \texttt{preference}, with the possible values being \texttt{require},
\texttt{prefer}, \texttt{no preference}, \texttt{avoid} and \texttt{prohibit}.
The difference between \texttt{require} and \texttt{prefer} is whether
or not the connection establishment will fail if the property cannot
be satisfied. This is mirrored in the difference between \texttt{avoid}
and \texttt{prohibit}.

An example of a selection property is \texttt{Reliable Data Transfer}.
This property is used to specify whether or not the application
should use a protocol which ensures that all data is received at 
a remote endpoint in order, without loss or duplication. Setting
this to \texttt{require} will force TAPS to use an underlying protocol
which guarantees this, such as TCP or QUIC. Setting this as
\texttt{prohibit} will force TAPS to use a protocol which does not
guarantee this, such as UDP.

An example of a connection property is \texttt{Timeout for Aborting Connection}.
This property is used to specify the how long to wait before deciding that an
active connection has failed when trying to reliably deliver data to the Remote Endpoint.
This naturally can only take effect if the underlying protocol supports reliability.

An example of a message property is \texttt{priority}. This property specifies
the prority of a given message relative to other messages over the same connection.
Lower priority message will yield in favor of higher priority messages.

TAPS is designed to provide an interface which more accurately
reflects how a typical application uses the network. Socket programming
is for example blocking by default, while most modern applications
interact with the network in an asynchronous way 
\cite{trammellAbstractApplicationProgramming2025}. TAPS
is therefore designed to be asynchronous. 

An additional
example of how TAPS is designed to reflect how applications
use the network is that TAPS is message-oriented. TCP
is a stream-oriented protocol and presents data to application
as a continious stream of bytes. However, most applications
use the network in a message-oriented way, and therefore
have to include delimiters and other logic to separate the
data into discrete messages. For example, several protocols
such as HTTP contain length fields, which are used to delimit
transferred data. In order to reflect
how the applications actually use the network, TAPS is therefore
message-oriented, while still supporitng applications reliant on 
streaming through the use of \texttt{framers}.

A framer is a data translation layer, which is used to define
how encapsulate or encode outbound message, and how to decapsulate
or decode inbound messages. This for example allows applications
to receive a byte stream as input if needed.

TAPS is backwards compatible with existing protocols as it does not
define how these underlying protocols should transfer data over
the network. TAPS is also designed to coexist with existing
socket programming APIs, allowing a gradual transition from
socket programming to TAPS.

Another key feature difference between TAPS and sockets is that socket
APIs are commonly limited to communicating with a single address over 
a single interface. TAPS is deisgned to handle multiple endpoints,
protocols and paths.

\section{Stream oriented vs message oriented (trash)}
Being stream-oriented or message-oriented has to do with how the
transferred data is represented. In a stream oriented protocol
such as TCP, the data is represented as a stream of bytes. From
the application's point of view, the data arrives continuously
and not in discrete chunks. This means that the appliaction
cannot easily differentiate between different messages. As
everything is ran over IP, which is a message-oriented protocol,
TCP, which is a transport layer protocol, has to be able to
handle receiving discrete messages and gluing them together into
a stream of bytes. This also includes handling messages arriving
out of order, which has to be handled before they can be
presented as a stream in order. This is called in-order delivery
and is a key feature guaranteed to the client by TCP.

With message-oriented protocols, such as UDP, the data is
represented as discrete messages. This means that the
application can easily differentiate between them. Because the
data is represented in discrete messages, this allows the
protocol to not have do deliver the messages in order. Firstly,
this allows the application to access the data quicker, as it
does not have to wait for any delayed packets to arrive, before
the other data can he handed over. This becomes very important
when multiplexing multiple streams of data over a single
connection. This is done in order to more efficiently utilize
bandwith, but faces problems if data from one stream is delayed
if we are using a protocol which guarantees in-order delivery.
Even if only one application's data is delayed, the second
application's data will have to wait, as they are together
treated as a single stream and have to be delivered in order.
This is called head-of-line blocking. When using a
message-oriented protocol, which makes it possible, but not
mandatory, to not guarantee in-order delivery, one application
does not have to wait for the other application's data.

\section{What is SCTP}

\section{What is quic}

QUIC is a internet protocol developed by Google, released to the
public in 2013, described in RFC 9000
\cite{iyengarQUICUDPBasedMultiplexed2021}. QUIC is a transport
layer protocol, which is designed to be a replacement for TCP.
QUIC is a connection-oriented protocol which is built on top of
UDP. QUIC enables 0-RTT connection establishment and integrates
encryption as a mandatory feature of the protocol.

The abstraction presented to an application for communication
via QUIC is called a stream. A stream can be either
unidirectional or biderectional. Streams can be created and
closed by either endpoint, and multple streams can exist over a
single QUIC connection. This is an example of the concept of
multiplexed connections. Individual pieces of data sent over a
stream are encapsulated as stream frames. The stream frame boundaries
are not expected to be preserved, and are therefore not visible
to the application. \todo{Is this true?}

By using the stream ID and the byte offset of the stream frame,
the receiving endpoint reassembles the data into the correct
order. All implementations must support in-order delivery of
data, they may choose to allow out-of-order delivery of data as
well, but this is not a requirement of the protocol This only
happens on a per-stream basis, so when multiplexing multiple
streams over a single connection, the streams are independent of
each other and we avoid head-of-line blocking.

A given stream frame has the following format, as describted in
RFC 9000:

\begin{figure}[H]
	\centering
	\begin{lstlisting}
        STREAM Frame {
          Type (i) = 0x08..0x0f,
          Stream ID (i),
          [Offset (i)],
          [Length (i)],
          Stream Data (..),
        }
    \end{lstlisting}
	\caption{Format of a QUIC stream frame \cite{iyengarQUICUDPBasedMultiplexed2021}}
	\label{fig:stream_frame_format}
\end{figure}

A given stream can exist in several different states.
These states differ whether the stream is sending or receiving
data. For a stream sending data, the states are:
\texttt{Ready}, \texttt{Send}, \texttt{Data Sent}, \texttt{Reset Sent},
\texttt{Data Recvd}, and \texttt{Reset Recvd}.
A receiving streams states maps the perceived states of the
sending stream. This means that there are some states
which are not represented in the receiving stream, as they
cannot be inferred from the data received. The states of a
receiving stream are: 
\texttt{Recv}, \texttt{Size Known}, \texttt{Data Recvd}, \texttt{Reset Recvd},
\texttt{Data Read}, and \texttt{Reset Read}.
A bidirectional stream consists of both sending and receiving parts,
and can therefore exist in states belonging to both sending and
receiving streams as described previously.

\begin{figure}[H]
    \centering
    \begin{lstlisting}
       o
       | Create Stream (Sending)
       | Peer Creates Bidirectional Stream
       v
   +-------+
   | Ready | Send RESET_STREAM
   |       |-----------------------.
   +-------+                       |
       |                           |
       | Send STREAM /             |
       |      STREAM_DATA_BLOCKED  |
       v                           |
   +-------+                       |
   | Send  | Send RESET_STREAM     |
   |       |---------------------->|
   +-------+                       |
       |                           |
       | Send STREAM + FIN         |
       v                           v
   +-------+                   +-------+
   | Data  | Send RESET_STREAM | Reset |
   | Sent  |------------------>| Sent  |
   +-------+                   +-------+
       |                           |
       | Recv All ACKs             | Recv ACK
       v                           v
   +-------+                   +-------+
   | Data  |                   | Reset |
   | Recvd |                   | Recvd |
   +-------+                   +-------+
    \end{lstlisting}
    \caption{States of a QUIC stream sending data \cite{iyengarQUICUDPBasedMultiplexed2021}}
    \label{fig:stream_states}
\end{figure}

Another important abstraction presented by QUIC is the connection.
A QUIC connection is a shared state between a client and a server 
\cite{iyengarQUICUDPBasedMultiplexed2021}.
A connection is established through a handshake, with support
for 0-RTT handshakes as mentioned previously. The connection
is identified by a connection ID, which is selected by each
endpoint's peer. Packets sent to this peer are identified as
belonging to a given connection, by a destination connection ID
field. Connection IDs make addressing in QUIC independent of addressing in lower layer
protocols \cite{iyengarQUICUDPBasedMultiplexed2021}. Each endpoint
has a unique connection ID, making sure that external observers cannot
identify that the two endpoints are communicating with each
other by looking at the connection ID.

Using a connection ID allows the connectoin to survive a change
in the addressing of the underlying protocol \cite{iyengarQUICUDPBasedMultiplexed2021}.
This functionality is enabled by default, but can be disbled by
one of the peers. When an address is changed, each endpoint has to
independently verify that they can communicate using the new address.
Path validation happens using a special type of stream frame
known as \texttt{PATH\_CHALLENGE} and \texttt{PATH\_RESPONSE}
frames. The \texttt{PATH\_CHALLENGE} frame contains unpredictable
data which must be mirrored back in the \texttt{PATH\_RESPONSE} frame.
If the endpoint validation fails, which happens if no matching response
is received within a given time period, the endpoints can continue
communicating using other known addresses. If there are no valid
addresses, the affected endpoint can signal this to the other
peer by using a \texttt{NO\_VIABLE\_PATH} error.

Much like TCP, QUIC also implements congestion control, there are
however some key differences between the algorithms. A key concept
to QUIC's congestion control is the 'packet number space'. Packet
number spaces are the context in which a packet can be processed
and acknowledged \cite{iyengarQUICUDPBasedMultiplexed2021}. 
\todo{Er det greit med direkte sitat på denne måten?}
This context reflects the level of encryption and there 
are therefore three separate packet number spaces. Initial space,
for the initial handshake packets, handshake space, for packets
exchanged during the cryptographic handshake, and application
data space, for packets exchanged after the cryptographic handshake
has been completed. Separating these spaces allows QUIC to
avoid spurious retransmission from one space affecting another,
and allows QUIC to handle retransmissions more efficiently, for
example by using a different expected acknowledgement delay for packets sent
in the initial space, as these are not expected to be cached
by the peer. Although the packet number spaces are handled
separately when it comes to acknowledgements, they are still
unified when it comes to congestion control and round trip
time measurements.

Another difference between how QUIC and TCP handle acknowledgements
is that QUIC separates the transmission order from the delivery order.
This is done by using stream offset as delivery order, and packet number as
transmission order. This removes ambiguity about whether the
original or the retransmitted packet is acknowledged in the case
of retransmissions, making RTT measurements in these cases more accurate
\cite{iyengarQUICUDPBasedMultiplexed2021}. Another key difference
is that instead of TCP algorithms such as retransmission timeout (RTO)
and tail-loss probe (TLP), QUIC uses probe timeout (PTO).
PTO is an algorithm which is used to detect lost packets. Every
time a packet requiring an acknowledgement is sent, a timer is started. If
the timer expires, the packet is resent and the PTO period is
set to twice its previous value. When an acknowledgement is received,
the PTO period is reset to its previous value, as long as the
client is certain that the server has finished validating the 
client's address.

For QUIC a single connection can exist over several paths and
congestion control is done on a per-path basis. Multiple streams
can be transmitted over a single connection and data from
multiple stream frames can be transmitted in a single packet.
If a packet is lost, only applications using the streams contained
in that packet will be affected. This is a key difference from TCP,
where a single lost packet will block all other streams of data
over the same connection. This is called head-of-line blocking.
This problem is avoided by QUIC, since QUIC only guarantees
in-order delivery of data on a per-stream basis.

Future network ossification related to QUIC is avoided
since QUIC mandates TLS 1.3 as part of the protocol.
By encrypting as much of the packet as possible, QUIC is able to
obscure details about the protocol from middleboxes, and
they cannot grow to depend on the contents of the packet. This
will avoid future problems with middleboxes expecting a certain
format like we see with them expecting either TCP or UDP.

The key advantages of QUIC over TCP is that it is designed to be
more efficient thatn TCP, and be able to handle more streams of
data in a better way due to the lack of head-of-line blocking.

As mentioned previously, one of the key selling points of TCP is
the ubiquity of sockets, which are implemented in kernel space.
There is no equivalent API for QUIC, and with a relatively
advanced implementation, users have to settle for one of the
existing libraries.

Currently, few applications can be built exclusively on top of
QUIC, as it requires support from both the client and the
server. QUIC is a transport protocol that is built on top of
UDP, and is designed to be a replacement for TCP. QUIC uses
multiplexed connections to for more efficient use of the
network. Multiplexed connections are also used in http/2, but
there it suffers from head-of-line blocking, which is caused by
TCP's requirement for all packets to be delivered in order. This
problem is exacerbated when using multiple streams of data over
a single TCP connection, as a single lost packet will block all
other streams.

With QUIC, the multiplexed connections can still be sent over a
single connection, but the streams are independent of each
other. This is made possible because QUIC is built on top of
UDP; which does not guarantee packet ordering for separate
streams.

\section{TAPS configurations explained}

TAPS exposes several configurations which can be used to specify
a preference of requirement, which then sets a preference or a
requirement for which underlying transport protocol to use.
These configurations come in the form of transport properties.
These properties are broken down into three categories:
Selection properites, Connection properties, and message
properties. Selection properties are requirements for the
underlying transport protocol, which TAPS uses to select which
protocol to use for the connection. Connection properties are
properties which are set on a per-connection basis, for example
the timeout for aborting a connection or the connection
priority. Message properties are properties which are set on a
per-message basis, for example the priority or the lifetime of
the given message. A full list of properites can be found in the
Architecture and Requirements for Transport Services RFC
\cite{paulyArchitectureRequirementsTransport2025}.

Selection properties can, as mentioned above, be specified to
varying degrees of strictness. With five different values
\texttt{Require}, \texttt{Prefer}, \texttt{No Preference},
\texttt{Avoid}, and \texttt{Prohibit}. Here, \texttt{Require}
means that establishing the connection will fail if the property
cannot be satisfied with the available transport protocols.
\texttt{Prefer} means that the connection can fall back to a
protocol not supporting the property, if no protocol satisfying
the property is available. \texttt{Avoid} and \texttt{Prohibit}
function in the same way as \texttt{Require} and
\texttt{Prefer}, but selecting for protocols which do not have
the selected property. A simple example of a selection property
is selecting for reliable data transfer. If this was set as
required, TAPS would be forced to use for example TCP or QUIC as
the underlying protocol. If UDP was the only available protocol,
establishing the connection would fail. If the user selected
\texttt{prefer}, then UDP could be selected if nothing else was
available, even though it does not fulfill the property.

There are also more advanced selection properties available. Out
of these, the least intuitive are \texttt{Multistream
	Connections in Group}, \texttt{Full Checksum Coverage on
	Sending} and \texttt{Full Checksum Coverage on Receiving},
\texttt{Provisioning Domain Instance or Type}, \texttt{Use
	Temporary Local Address}, and \texttt{Notification of ICMP soft
	error message arrival}.

The \texttt{Multistream Connections in Group} selection property
allows the user \todo{TODO - select term for "user" of TAPS api
	(developer of networking application)} to select whether the
application would prefer multiple Connections within a
Connection Group to be provided by streams of a single
underlying transport connection where possible
\cite{paulyArchitectureRequirementsTransport2025}.
Multistreaming in this context is the act of providing multiple
streams of data over a single underlying connection. This is
advantageous as it allows for more efficient use of the network,
since the connection will better saturate the available
bandwidth. This avoids long convergence times for the connection
and allows fresh connection to utilize the large congestion
window already achieved, without having to wait for slow start
to ramp up. This is especially useful for applications with many
small streams of data, such as web browsers.

The \texttt{Full Checksum Coverage on Sending} and \texttt{Full
	Checksum Coverage on Re\-ceiving} selection properties specify
the application's need for protection against corruption for all
data transmitted or received on a given connection. This is used
to specify how many bytes of the message which needs to be
protected by a checksum. This checksum is used to protect the
data from corruption caused by lower-layer protocols. The
setting \texttt{Full coverage} means that the entire message is
protected, while \texttt{0} specifies no checksum is required.
If any other value than \texttt{Full coverage} is specified, the
connection may still fall back to \texttt{Full coverage}.

The \texttt{Provisioning Domain Instance or Type} selection
property allows the applicatio \todo{common application term}n
to control path selection by selecting for specific Provisioning
Domain or categories of domains
\cite{paulyArchitectureRequirementsTransport2025}. Provisioning
Domains define consistent sets of network properties that might
be more specific than network interfaces
\cite{paulyArchitectureRequirementsTransport2025}.

The \texttt{Use Temporary Local Address} selection property
allows the application \todo{common application term} to specify
whether or not to use temporary addresses which are a feature in
IPv6, used for greater privacy
\cite{gontTemporaryAddressExtensions2021},\cite{paulyArchitectureRequirementsTransport2025}.

The \texttt{Notification of ICMP soft error message arrival}
selection property lets an application \todo{common application
	term} specify whether or not it should be notified when an ICMP
error, which does not force termination of the connection,
arrives \cite{paulyArchitectureRequirementsTransport2025}.

In addition to specifying which underlying protocol should be
used through selection properties, the application \todo{common
	application term} using TAPS can manage the connections
themselves through transport properties. These can be specific
for a given transport protocol such as the selection properties,
or generic and be available for all transport protocols.
\todo[]{Should this be separated from all the above parts? are
	they all part of the same 'category'}

The \texttt{Connection Group Transmission Scheduler} transport
property decides which scheduler is used within a given
connection group, to divide the available capacity between the
connections. \cite{paulyArchitectureRequirementsTransport2025}.

The \texttt{Capacity Profile} transport property is used to
select a capacity profile, which will be a tradeoff between
delay, delay variation, and efficient use of the available
capacity based on the capacity profile specified
\cite{paulyArchitectureRequirementsTransport2025}.

The \texttt{Isolate Session} transport property set in order to
have new connections use as little cacheed information as
possible from previous connections not in the same connection
group \cite{paulyArchitectureRequirementsTransport2025}.
