\chapter{Past work}

\section{NEAT}
One of the first open source implementations of TAPS is 
NEAT (A New, Evolutive API and Transport-Layer Architecture for the Internet) 
\cite{khademiNEATPlatformProtocolIndependent2017}. NEAT and its associated
library, NEAT Library were funded alongside the TAPS project from 2015 to 2018.
Developed alongside TAPS, the development of NEAT affected the TAPS API and
the TAPS API natrually affected the development of NEAT. NEAT abstracts
away transport layer communication through the use of flows. Much like
a connection in TAPS, a flow can be first created without specifying
which transport protocol to use. Much like TAPS, NEAT is also asynchronous,
using an event loop and allowing for users to add their own callbacks.
An example of where a user may add their own callback is the
\texttt{on\_connect} callback, which allows the user to specify
what to do when a conneciton is first established.

Much like TAPS, NEAT uses properties, which consists of key-value pairs,
in order to specify requirements for the underlying transport protocol.
Unlike TAPS, properties in NEAT are not given as transport property
objects, but rather as JSON objects. The biggest difference is however
that NEAT forces the application to directly specify one or more protocols
to be used, instead of selecting the characteristics desired for the communication.
According to the documentation, \todo{Reference the documentation}
NEAT supports two properties, \texttt{transport}
and \texttt{security}. The transport property specifies which underlying
transport protocol to use, and supports SCTP, SCTP/UDP (SCTP tunneled over UDP),
TCP, UDP and UDP-Lite. \todo{Is this actually true?, check source code
to double check}

Instead of making the application directly select the transport protocol,
TAPS focuses on the desired characterisitics of the connection, as described
in the chapter \ref{chap:background}.

Since the conclusion of the NEAT project, TAPS has continued to evolve
so today the APIs differ to some extent. NEAT also does not feature
key selling points of TAPS, such as the ability to use QUIC.

\section{NEATPy}
NEATPy \cite{gundersenNEATPyStandardsConformantProtocolIndependent2020}
is a Python implementation and further development of the NEAT project.
It was undertaken as a master thesis at the department of informatics
at the University of Oslo. It was undertaken as an attempt to create a
Python layer to bridge the growing gap between NEAT and TAPS.

Because NEATPy is a Python shim on top of NEAT, certain parts of it is
limited by NEAT itself. For example,
using NEATPy you cannot specify the \texttt{Interface Instance or Type}
or the \texttt{Use Temporary Local Address} selection properties 
\cite{NEATPyDocumentationNEATPy}. In addition, NEATPy does not
allow support QUIC, as this is not supported by NEAT. 

Another
limitation of NEATPy is that it was developed before the finalization of
the TAPS standard. Developed in 2020, NEATPy is based mostly on
version 4 and 5 of the TAPS interface draft \cite[p. 109]{gundersenNEATPyStandardsConformantProtocolIndependent2020}.
The TAPS API has since evolved, and was 
published as an proposed standard in January 2025
\cite{trammellAbstractApplicationProgramming2025}.

Since the finalization of NEATPy, several new features have been added
to TAPS.
For example, the support for giving more than one local and remote
endpoint when creating a PreConnection. This is done when the application
wants to indicate that several endpoints are eligible for the connection
or provide the same service. Another key change is the support for
multicast.

In addition, there are severeal new selection properties that have been added
to TAPS. These are:
\begin{itemize}
    \item Keep-Alive Packets
    \item Use Temporary Local Address
    \item Multipath Transport
    \item Advertisement of Alternative Addresses
    \item Initiating Side Is Not the First to Write
\end{itemize}

Some selection properties have also been removed from the TAPS API:
\begin{itemize}
    \item Parallel Use of Multiple Paths
    \item Notification of excessive retransmissions
\end{itemize}

In addition to the outward differences to the API, there are also several
changes to the required internal workings of TAPS since the development of NEATPy.
For example, in the final RFC TAPS has loosened the required
order of operations when working with selection properties.
Previously, there was a strict precedence in the order of
which preference values were to be applied. This has since been
relaxed, and the prioritzation has been moved from being
related to the preference level, to being ordered by which
selection properties will affect the path selection.

The combination of the limitations of NEAT and the changes to TAPS
since the development of NEATPy, NEATPy is an implementation with
limited usability for real-world applications wanting to use TAPS.

\section{PyTAPS}
PyTAPS \cite{frankePythonasynciotaps2025} is a Python library aiming
to be fully compliant with TAPS. Unlike NEATPy, PyTAPS is not built
on top of NEAT, and therefore does not suffer from the samme limitations
that NEATPy does. However, PyTAPS is based on the fourth version of the TAPS
interface draft \cite{frankePythonasynciotaps2025} and has not had active
development since 2020. As such, it naturally suffers from the same
issues as NEATPy when it comes to the changes in the TAPS API.

\section{Dynamic TAPS}
dynamic-taps is a C implementation of TAPS, developed by the
American technology company F5 \cite{dukeDynamictaps2025}. It
is in the very early stages of development and has not had
active development since 2021.


