\chapter{Implementation}  \label{chap:impl}

\section{Architecture}
Internally, CTaps follows the TAPS architecture as described in
RFC 9621 \cite{paulyArchitectureRequirementsTransport2025}.
This divides a TAPS system into two distinct sections, a public
Transport Services API described in \cref{cTapsApiDesign} and the
Transport System Implementation which is described in this chapter.

The architecture of CTaps has been designed to fit the restrictions
placed on socket lifetimes by the underlying protocols of TCP, UDP
and QUIC. Because the application can close and free connections
at any point, CTaps has to ensure that all required memory remains
valid until the final external resource using it has been freed.

CTaps therefore makes use of an internal structure called a
Socket Manager which is described in detail in \cref{sec:socketManager}.

Because UDP mandates that multiple unrelated Connections must
be able to share a socket, TCP mandates that Connections in the
same Connection Group must be able to have different sockets Socket
Managers are referenced on the Connection level. The application
interacts only with the connection
which in turn interacts with the Socket Manager. The Socket Manager
owns the protocol implementation and invokes one or more of the interface
function pointers, according to the will of the application.

Since several functions within the various protocol implementations
are asynchronous, communication between the protocol implementation
and the socket manager has to be asynchronous as well. The Socket
Manager has a structure with internal callbacks which are invoked by the protocol
implementation at appropriate times, for example to notify
the Socket Manager that a Connection has been closed when teardown
has completed.

This closed-layer model only applies to memory-critical paths, in order
to mitigate
potential memory leaks. Some remaining flows, such as messages being passed
from Connection to protocol do not currently flow through a dedicted
function in the Socket Manager. 

An example of the internal flow in CTaps can be seen in
\cref{fig:async-close-flow}, showing the asynchronous flow of
graceful connection teardown for a Connection using a shared
Socket Manager. Here, a single Connection in a Connection Group
is closed.

This Connection uses a Socket Manager which is shared
with not only its Connection Goup, but also with a Connection
in a completely separate Connection Group. This can for example
happen when listening for Connections over UDP, as all received
Connections will share a socket but belong to different Connection
Groups. When the Connection is closed, the Socket Manager invokes
the associated function pointer for the underlying protocol. This
asynchronously performs any necessary connection teardown over the
wire, and then notifies the Socket Manager when necessary.

Crucially,
it does not close the socket until notified by the Socket Manager,
since the socket needs to stay open until all Connections using
it have been closed. This is not the case in the example flow,
so the Socket Manager simply updates the state of the Connection
and notifies the application of the Connection closing through the
\mono{closed()} callback.

\definecolor{ConnectionGroupTextColor}{RGB}{139, 194, 110}
\begin{figure}[ph]
\centering
\begin{tikzpicture}[font=\sffamily,
    block/.style={rectangle, draw, text centered, rounded corners, fill=white, font=\bfseries},
    client/.style={rectangle, draw, text centered, fill=white, font=\small},
    arrow/.style={-Stealth, thick},
    asyncArrow/.style={arrow, dashed, red!70!black},
    label/.style={font=\scriptsize\itshape}
    ]

% Place connections first
\node[client] (Conn1) at (0,0) {Connection 1};
\node[client, below=0.3cm of Conn1] (Conn2) {Connection 2};
\node[client, below=1.2cm of Conn2] (ConnN) {Connection N};

% SM centred vertically on the connections group
\node[block, align=center] (SM) at
    ($(Conn1.north east)!0.5!(ConnN.south east) + (4,0)$)
    {Socket Manager};

\node[block, right=of SM] (Prot) {Protocol Impl};

\draw[arrow] (Conn1.north east) -- node[above, sloped, label, yshift=3pt] {close()} (SM.north west);
\draw[arrow] (Conn2.east) -- (SM.west);
\draw[arrow] (ConnN.east) -- (SM.south west);

\node[font=\Huge, text=gray] at ($(Conn2.south)!0.4!(ConnN.north)$) {$\vdots$};

\begin{scope}[on background layer]
    \draw[rounded corners, fill=blue!5!white, draw=blue!20]
        ($(Conn1.north west)+(-0.3, 0.4)$) rectangle ($(ConnN.south east)+(0.3,-0.3)$);
    \node[font=\normalsize\bfseries, above=0.4cm of Conn1, blue!70!black, align=center]
        {Connections};
    \node[fit=(Conn1)(Conn2), rounded corners, fill=green!5!white,
          draw=green!20, inner sep=0.13cm] (GreenBox1) {};
    \node[fit=(ConnN), rounded corners, fill=green!5!white,
          draw=green!20, inner sep=0.13cm] (GreenBoxN) {};
\end{scope}
\node[font=\tiny\bfseries, text=ConnectionGroupTextColor, above=-1.2mm of GreenBox1] {Connection Group 1};
\node[font=\tiny\bfseries, text=ConnectionGroupTextColor, above=-1.2mm of GreenBoxN] {Connection Group N};

% Request Flow
\draw[arrow]
    ([yshift=-0.5mm]SM.north east)
    -- node[above, sloped, label, align=center] {Close connection}
    ([yshift=-0.5mm]Prot.north west);
\draw[asyncArrow] 
    ([yshift=0.5mm]Prot.south west)
    -- node[below, label, text=red!70!black, align=center] {Socket Manager cb}
    ([yshift=0.5mm]SM.south east);
\draw[asyncArrow]
    ([yshift=4pt]SM.west)
    -- node[below, sloped, label, text=red!70!black, align=center] {User cb}
    ([yshift=3pt]Conn1.east);
\end{tikzpicture}
\caption{Flow of asynchronous close() through CTaps}
\label{fig:async-close-flow}
\end{figure}

\section{Candidate Gathering} \label{sec:candidateGathering}
Before a Connection can be established, a list of possible Local and Remote
need to be generated based on the data passed to the associated preconnection.
In CTaps this candidate gathering follows the procedure described in RFC 9623
very closely. CTaps represents the candidates in a tree, using a GNode struct from
GLib to represent a N-ary tree.

CTaps first builds the entire tree, resolving
all possible local interfaces, all protocols, specified ALPNs and all possible resolutions of
remote hostnames respectively. For each combination if these criteria,
CTaps generated a single candidate node.

After the entire tree has been built, CTaps prunes the branches which
are not compatible with the required or prohibited selection properties. After pruning,
the tree is sorted according to the stated selection properties, as described in RFC 9623 
\cite{brunstromImplementingInterfacesTransport2025}.

\section{Candidate Racing} \label{sec:candidateRacing}
After a list of candidates has been gathered when initiating a preconnection, CTaps
will race them according to the specification from RFC 9623
\cite{brunstromImplementingInterfacesTransport2025}. The combination
of protocol stack, protocol options, local and remote endpoints are referred to as
a candidate.

The goal of candidate racing is to successfully negotiate a protocol stack, local
and remote endpoints based on the user's specification. Candidate racing in a TAPS
system follows the same procedure as specified in Happy Eyeballs V2 \cite{schinaziHappyEyeballsVersion2017},
which was originally designed to increase IPv6 adoption, see \cref{happyEyeballs}.
TAPS takes racing a step further by not only racing IP-address families, but also
racing potential protocol configurations and endpoints.

CTaps follows a staggered
approach, first sorting the gathered candidates by preference and performing a connection
attempt for each, with a delay between each candidate. All outstanding racing attempts
are cancelled upon the first success.

Since CTaps has no internal state and no support for RTT estimation, the delay
between each candidate is set to a fixed 250 ms which is the recommended value
in Happy Eyeballs V2 \cite{schinaziHappyEyeballsVersion2017}.

An example flow of candidate racing can be seen in \cref{fig:racingSeqDiagram}.
Here, DNS resolution resolves two endpoints, with the endpoints having an
RTT of 300 and 100 ms respectively. The slower endpoint gets a head start,
but the faster RTT leads to the faster endpoint winning the race, and being
presented to the application.


\begin{figure}[ph]
\centering
\begin{tikzpicture}[
  >=Stealth,
  msglbl/.style={font=\scriptsize, fill=white, inner sep=1pt},
  head/.style={draw, rectangle, minimum width=1.8cm, minimum height=0.6cm,
               font=\small, fill=white}
]

% ── x positions ─────────────────────────────────────────────────────────────
\def\xO{0}     % Application
\def\xA{2.5}     % CTaps
\def\xB{4.5}   % DNS
\def\xC{6.5}
\def\xD{8.5}
\def\xE{10.5}  % Server

% ── y positions (y = -t/100 * 1.2) ─────────────────────────────────────────

%  t=0:    y=-0.4 (Query)
%  t=100:  y=-1.2 (DNS start)
%  t=200:  y=-2.0
%  t=300:  y=-2.8
%  t=400:  y=-4.0
%  t=500:  y=-5.2

% ── Participant headers ──────────────────────────────────────────────────────
\node[head]                   at (\xO, 0.4) {Application};
\node[head]                   at (\xA, 0.4) {CTaps};
\node[head]                   at (\xB, 0.4) {DNS};
\node[head]                   at (\xC, 0.4) {Cand. 1};
\node[head, font=\small]      at (\xD, 0.4) {Cand. 2};
\node[head]                   at (\xE, 0.4) {Server};
% \node[font=\scriptsize, gray] at (\xA, -0.05) {(per-candidate)};
% \node[font=\scriptsize, gray] at (\xB, -0.05) {(per-candidate)};

% ── Lifelines ───────────────────────────────────────────────────────────────
\draw[dashed, gray!50] (\xO, 0.1) --  (\xO, -4.9);
\draw[dashed, gray!50] (\xA, 0.1) --  (\xA, -4.9);
\draw[dashed, gray!50] (\xB, 0.1) --  (\xB, -4.9);
\draw[dashed, gray!50] (\xE, 0.1) --  (\xE, -4.9);

\draw[dashed, gray!50] (\xC, -1.2) -- (\xC, -4.1);
\draw[dashed, gray!50] (\xD, -2.4) -- (\xD, -4.9);

%
% \draw[gray!20] (-0.2, -0.8) -- (12.2, -0.8);

% ── Time grid and labels ─────────────────────────────────────────────────────
\foreach \yval/\t in {-0.4/0, -1.2/100, -2.0/200, -2.4/250, -2.8/300, -4.0/500} {
  \node[font=\footnotesize, gray, left] at (-0.4, \yval) {\t\,ms};
  \draw[gray!20] (-0.2, \yval) -- (12.2, \yval);
}

% ── t≈0: Start Connection attempt ─────────────────────────────────────────────────────────
\draw[->] (\xO, -0.4) -- node[msglbl, above] {Connect to URL} (\xA, -0.4);

% ── t≈0: DNS queries ─────────────────────────────────────────────────────────
\draw[->] (\xA, -0.4) -- node[msglbl, above] {Query URL} (\xB, -0.4);

% ── t=100: DNS resolves A ────────────────────────────────────────────────────
\draw[->] (\xB, -0.8) -- node[msglbl, above] {Resolve {C1, C2}} (\xA, -0.8);

\draw[fill=black] (\xA,-1.2) circle (1.5pt);
\draw[->] (\xA, -1.2) -- node[msglbl, above] {Start attempt 1} (\xC, -1.2);

% ── QUIC handshake for candidate A ───────────────────────────────────────────
% Initial:        SM-A (t=100) → Server (t=175)
\draw[->] (\xC,-1.2)  -- node[msglbl, above, sloped] {}         (\xE,-2.4);
% Handshake:      Server(t=175) → SM-A (t=250)
\draw[->] (\xE,-2.4)  -- node[msglbl, above, sloped] {}       (\xC,-3.6);
% Finished:       SM-A (t=250) → Server (t=325)
\draw[-, red!90] (\xC,-3.6)
  -- node[msglbl, above, sloped] {}
     node[pos=1.0, cross out, draw, inner sep=2pt] {}
  (\xC + 2.08,-4.1);

% ── t=350: racing timer fires, create SM-B ───────────────────────────────────
\draw[fill=black] (\xA,-2.4) circle (1.5pt);
%\node[font=\scriptsize, right=3pt] at (\xA,-2.4) {250\,ms timer};
\draw[->] (\xA,-2.4) -- node[msglbl, above] {Start attempt 2} (\xD,-2.4);

% ── QUIC handshake for candidate A ───────────────────────────────────────────
\draw[->] (\xD,-2.4)  -- node[msglbl, above, sloped] {}         (\xE,-2.8);
\draw[->] (\xE,-2.8)  -- node[msglbl, above, sloped] {}       (\xD,-3.2);
\draw[->] (\xD,-3.2) -- node[msglbl, above, sloped] {}         (\xE,-3.6);
\draw[->] (\xE,-3.6) -- node[msglbl, above, sloped] {} (\xD,-4.0);
\draw[->] (\xD,-4.0) -- node[msglbl, above, sloped] {}   (\xA,-4.0);

% ── t=400: SM-A wins ─────────────────────────────────────────────────────────
\node[draw, fill=green!15, font=\scriptsize, rounded corners, inner sep=2pt]
      at (\xD,-5.2) {Connected \cmark};

\node[draw, fill=red!15, font=\scriptsize, rounded corners, inner sep=2pt]
      at (\xC,-5.2) {Closed \xmark};

% ── t=400: abort SM-B ────────────────────────────────────────────────────────
\draw[->, red!70, thick] (\xA,-4.1) -- node[msglbl, below] {Close} (\xC,-4.1);

\draw[->] (\xA, -4.2) -- node[msglbl, above] {\mono{ready()} CB} (\xO, -4.2);

\end{tikzpicture}
\caption{Example flow of Candidate Racing with a fast and a slow endpoint.}
\label{fig:racingSeqDiagram}
\end{figure}

\section{Connection Cloning}
CTaps support connection cloning through the \mono{ct_connection_clone()}
function, which allows applications to create a new Connection to the same
endpoint, either through normal connection establishment or through
multistreaming if supported by the underlying protocol.

After cloning a Connection, the resulting clone is passed to the application
through the \mono{ready()} callback registered with the source Connection.
The clone will be organised in the same Connection Group as the source Connection,
and will therefore share most transport properties with the source Connection,
and will be affected by Connection Group operations such as \mono{ct_connection_close_group()}. 

The transport properties that are not shared is individual Connection state, the Connection
priority and the Connection's \mono{can_send} and \mono{can_receive} flags. The
internal Connection state track whether the Connection is initiating, open or closed.
The Connection priority is used to prioritize data transfer between the Connections
in the same Connection Group, while the \mono{can_send} and \mono{can_receive}
are used to keep track of whether the Connection is unidirectional.

Connection cloning is naturally the most useful when the underlying protocol can make
use of multistreaming, since this allows the clone to reuse the same underlying connection.
However, since CTaps strives to be as protocol agnostic as possible Connection cloning
can be performed no matter what the underlying protocol is. This is an advantage since
it allows code using CTaps to be protocol-agnostic, attempting to make use of benefits
like multistreaming when available, but being functional even if the underlying protocol
does not support it.

For simplicities sake, all connections in CTaps belong to a Connection Group. This
means that it is possible to handle Connections the same way, regardless of whether
the Connection has been cloned or not.

The implications of the application having to make an effort to write
protocol-agnostic code when using Connection cloning in CTaps is discussed
in \cref{sec:quicUnderTaps}.

A theoretical example displaying the benefits of Connection cloning and multistreaming
in general can be seen in \cref{fig:connection-clone-gains}. Here, two
transfers are performed sequentially. After the first transfer completes successfully,
the application makes use of connection cloning in CTaps and starts a second transfer.
If the underlying protocol is TCP, CTaps will initiate a new connection and transfer
successfully, but if the underlying protocol is QUIC then it will simply open a new
stream and therefore benefit from stream multiplexing. The second transfer will therefore
have significantly higher throughput over QUIC than TCP, since TCP has to go through
slow start again for the second transfer. In CTaps a single client can use both
TCP and QUIC this way, supporting connecting to a TCP endpoint, but making use
of QUIC benefits when available.

A benchmark showing the measured benefits of utilizing Connection cloning over
QUIC vs. TCP can be seen in \cref{sec:smallfiletransfer}.

\begin{figure}[ph]
\centering
\begin{tikzpicture}[font=\small, >=stealth]

%% ============ LEFT PANEL: TCP ============
\begin{scope}

  % Lost-capacity shading
  \fill[blue!12]
    (3.1, 3.0) -- (5.1, 3.0)
    .. controls (4.6, 3.0) and (3.6, 0.15) .. (3.1, 0.15) -- cycle;

  % Axes
  \draw[->, thick] (0, 0) -- (5.6, 0) node[right] {Time};
  \draw[->, thick] (0, 0) -- (0, 4.2) node[above] {CWND};

  % W_max reference
  \draw[dashed, gray!60] (0, 3.0) -- (5.3, 3.0);
  \node[right, gray!60, font=\scriptsize] at (5.5, 3.0) {$W_{\max}$};

  % Transfer 1: quick ramp then flat at W_max
  \draw[very thick, blue!70!black]
    (0, 0.15) .. controls (0.5, 0.15) and (1.5, 3.0) .. (2.0, 3.0)
    -- (3.0, 3.0);

  % Transfer boundary
  \draw[dashed, gray] (3.1, -0.3) -- (3.1, 3.9);
  \node[above, font=\scriptsize, gray] at (3.1, 3.9) {Clone Connection};

  % Transfer 2: TCP slow starts again
  \draw[very thick, blue!70!black]
    (3.1, 0.15) .. controls (3.6, 0.15) and (4.6, 3.0) .. (5.1, 3.0);

  % Lost capacity annotation
  \node[font=\scriptsize, black] at (3.85, 2.75) {Lost};
  \node[font=\scriptsize, black] at (3.85, 2.35) {capacity};

  % X-axis region labels
  \node[below, font=\scriptsize] at (1.5, 0) {First Transfer};
  \node[below, font=\scriptsize] at (4.2, 0) {Second Transfer};

  % Panel title
  \node[font=\small\bfseries] at (3.0, 4.7) {TCP};

\end{scope}

%% ============ RIGHT PANEL: QUIC ============
\begin{scope}[xshift=7.0cm]

  % Axes
  \draw[->, thick] (0, 0) -- (5.6, 0) node[right] {Time};
  \draw[->, thick] (0, 0) -- (0, 4.2) node[above] {CWND};

  % W_max reference
  \draw[dashed, gray!60] (0, 3.0) -- (5.3, 3.0);
  \node[right, gray!60, font=\scriptsize] at (5.5, 3.0) {$W_{\max}$};

  % Transfer 1: quick ramp then flat at W_max
  \draw[very thick, green!50!black]
    (0, 0.15) .. controls (0.5, 0.15) and (1.5, 3.0) .. (2.0, 3.0)
    -- (3.0, 3.0);

  % Transfer boundary
  \draw[dashed, gray] (3.1, -0.3) -- (3.1, 3.9);
  \node[above, font=\scriptsize, gray] at (3.1, 3.9) {Clone Connection};

  % Transfer 2: QUIC resumes at W_max
  \draw[very thick, green!50!black]
    (3.1, 3.0) -- (5.1, 3.0);

  % CWND reuse annotation
  \node[font=\scriptsize, green!40!black] at (4.1, 3.6) {Stream};
  \node[font=\scriptsize, green!40!black] at (4.1, 3.2) {multiplexing};

  \node[below, font=\scriptsize, green!40!black] at (1.8, 2) {First};
  \node[below, font=\scriptsize, green!40!black] at (1.8, 1.6) {Stream};

  \node[below, font=\scriptsize, green!40!black] at (4.2, 2) {Second};
  \node[below, font=\scriptsize, green!40!black] at (4.2, 1.6) {stream};

  % X-axis region labels
  \node[below, font=\scriptsize] at (1.5, 0) {First Transfer};
  \node[below, font=\scriptsize] at (4.2, 0) {Second Transfer};

  % Panel title
  \node[font=\small\bfseries] at (3.0, 4.7) {QUIC};

\end{scope}

\end{tikzpicture}
\caption{Lost capacity when Cloning TCP vs. QUIC Connections.}
\label{fig:connection-clone-gains}
\end{figure}

\section{Protocol Implementations}

\subsection{Protocol Interface}
In order to make the CTaps design as modular as possible,
CTaps uses an internal interface to interact with the underlying protocol.
This interface abstracts away the implementation details of the individual
protocols and allows the upper layers to interact with the network in an
asynchronous way. This interface consists of the protocol name and certain
protocol properties, a list of selection properties which is used
when filtering candidates during Candidate Gathering, as well as
a list of function pointers invoked when interacting with the protocol.
The full defintion of the protocol interface struct can be seen in
\cref{lst:protocol-impl-struct}.

A single protocol is implemented as a single implementation of the interface.
Together with providing implementations of the interface protocols, it also
has to register a list of settings for each selection protocol. These settings
indicate which properties the given protocol can fulfill, which is used 
when filtering candidates during candidate gathering.

\begin{listing}[tbp]
\begin{minted}{C}
typedef struct ct_protocol_impl_s {
    const char* name;                               ///< Protocol name (e.g., "TCP", "UDP", "QUIC")
    ct_protocol_enum_t protocol_enum;               ///< Protocol enumeration value
    bool supports_alpn;                             ///< True if protocol supports ALPN negotiation
    ct_selection_properties_t selection_properties; ///< Properties supported by this protocol
    /** @brief Initialize a new connection using this protocol. */
    int (*init)(ct_connection_t* connection);
    /** @brief Initialize a new connection using this protocol and attempt 0-rtt. */
    int (*init_with_send)(ct_connection_t*, ct_message_t*, ct_message_context_t*);
    /** @brief Send a message over the protocol. */
    int (*send)(ct_connection_t*, ct_message_t*, ct_message_context_t*);

    /** @brief Start listening for incoming connections. */
    int (*listen)(struct ct_socket_manager_s* socket_manager);

    /** @brief Stop listening for incoming connections. */
    void (*close_listener)(struct ct_socket_manager_s*);

    /** @brief Close a connection. */
    int (*close_connection)(ct_connection_t*);

    /** @brief Close the underlying socket for a socket manager. */
    void (*close_socket)(struct ct_socket_manager_s*);

    /** @brief Forcefully abort a connection without graceful shutdown. */
    void (*abort)(ct_connection_t* connection);

    /** @brief Clone a connection's protocol specific state. */
    int (*clone_connection)(const ct_connection_t* source_connection,
                            ct_connection_t* target_connection);

    /** @brief Free protocol-specific state in a connection. */
    void (*free_connection_state)(ct_connection_t* connection);

    /** @brief Free socket-specific state in a connection. */
    void (*free_socket_state)(struct ct_socket_manager_s* socket_manager);

    void (*close_connection_group)(ct_connection_group_t* connection_group);

    int (*set_connection_priority)(ct_connection_t* connection, uint8_t priority);

    /** @brief Free protocol-specific shared state in a connection group, useful for multiplexing */
    void (*free_connection_group_state)(ct_connection_group_t* connection_group);

} ct_protocol_impl_t;
\end{minted}
\caption{Protocol Implementation interface}
\label{lst:protocol-impl-struct}
\end{listing}

\subsection{Socket Manager} \label{sec:socketManager}
In order to manage sockets, CTaps uses a structure it refers to as a Socket Manager
internally. This is done to maintain the lifetime of the socket and to
own its internal state. Since a single socket can feature many connections
depending on the associated protocol, the socket may need to remain valid
for the entire lifetime of its most long-lived Connection or Listener.
In order to achieve this the Socket Manager contains a counter for the
number of objects still using it. This counter is decremented when the
memory associated with one of the objects using the Socket Manager is
freed. If this counter reaches 0, then the entire Socket Manager structure
is freed, as can be seen in \cref{lst:socketManagerRefCount}.

Socket Managers
are also responsible for closing sockets when the final associated Connection
or Listener has been closed. This is done through a separete functionality from
the reference counting. Instead, a Socket Manager maintains a list of Connections
and Listeners using its socket, only closing the socket when the last of these
Connections or the last Listener is closed. Freeing the internal state and closing
of the socket are implemented as two separate operations so that the application
can still query for internal state, such as the underlying protocol, after
a Connection or Listener has been closed.

\begin{listing}[tbp]
  \begin{minted}{C}
void socket_manager_decrement_ref(SocketManager* socket_manager) {
  socket_manager->ref_count--;
  log_debug("Decremented socket manager reference count, updated count: %d", socket_manager->ref_count);
  if (socket_manager->ref_count == 0) {
    int rc = socket_manager->protocol_impl.stop_listen(socket_manager);
    if (rc < 0) {
      log_error("Error stopping socket manager listen: %d", rc);
    }
    free(socket_manager);
  }
}
  \end{minted}
  \caption{Internal Socket Manager counter decrementation}
  \label{lst:socketManagerRefCount}
\end{listing}


\subsection{Supported protocols}
Currently, CTaps natively supports three transportation layer protocols:
UDP, TCP and QUIC. TCP and UDP were added because of their ubiquity and
relative ease of implementation due to native support in libuv. QUIC was
added since having easier access to QUIC was one of the primary motivations
behind the thesis as a whole. Using CTaps, an application can get access to QUIC,
without having to implement custom fallback-logic or error handling. This
greatly eases the implementation complexity and could help QUIC adoption
on both the server and client side.

\subsubsection{UDP} \label{sec:udpImpl}
UDP is a connectionless, message-based unreliable protocol first designed
in 1980 \cite{postelUserDatagramProtocol1980}. It has since become one
of the two primary transport layer protocols used today and has native
support in libuv through the libuv \texttt{udp\_udp\_t} handle \cite{libuvLibuvUDPHandle}.
Due to its simplicity, there is fairly little logic required to integrate
UDP into CTaps. The relative protocol complexity is discussed further in
in \cref{sec:protocolComplexity}.

\begin{listing}[tbp]
\begin{minted}{C}
const ct_protocol_impl_t
    udp_protocol_interface =
        {
            .name = "UDP",
            .protocol_enum = CT_PROTOCOL_UDP,
            .supports_alpn = false,
            .selection_properties =
                {.list =
                     {
                         [RELIABILITY] = {.value = {.simple_preference = PROHIBIT}},
                         [PRESERVE_MSG_BOUNDARIES] = {.value = {.simple_preference = REQUIRE}},
                         [PER_MSG_RELIABILITY] = {.value = {.simple_preference = PROHIBIT}},
                         [PRESERVE_ORDER] = {.value = {.simple_preference = PROHIBIT}},
                         [ZERO_RTT_MSG] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [MULTISTREAMING] = {.value = {.simple_preference = PROHIBIT}},
                         [FULL_CHECKSUM_SEND] = {.value = {.simple_preference = REQUIRE}},
                         [FULL_CHECKSUM_RECV] = {.value = {.simple_preference = REQUIRE}},
                         [CONGESTION_CONTROL] = {.value = {.simple_preference = PROHIBIT}},
                         [KEEP_ALIVE] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [INTERFACE] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [PVD] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [USE_TEMPORARY_LOCAL_ADDRESS] = {.value = {.simple_preference =
                                                                        NO_PREFERENCE}},
                         [MULTIPATH] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [ADVERTISES_ALT_ADDRESS] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [DIRECTION] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [SOFT_ERROR_NOTIFY] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [ACTIVE_READ_BEFORE_SEND] = {.value = {.simple_preference =
                                                                    NO_PREFERENCE}},
                     }},
            .init = udp_init,
            .init_with_send = udp_init_with_send,
            .send = udp_send,
            .listen = udp_listen,
            .close_listener = udp_close_listener,
            .close_connection = udp_close,
            .close_socket = udp_close_socket,
            .abort = udp_abort,
            .clone_connection = udp_clone_connection,
            .free_connection_state = udp_free_state,
            .free_socket_state = udp_free_socket_state,
            .close_connection_group = udp_close_connection_group,
            .free_connection_group_state = udp_free_connection_group_state,
};
\end{minted}
\caption{UDP protocol implementation struct}
\label{lst:udp-protocol-impl-struct}
\end{listing}

UDP does however not fit perfectly into the model assumed by
TAPS, which is that of a Connection-based system. There is therefore
several pieces of logic in the CTaps UDP implementation which adapt
the underlying UDP to fit into a TAPS model. Opening a client is
fairly simple, as this simply creates a new UDP handle through
libuv, binds it to either the local endpoint required or an ephemeral one
before finally starting to receive data over that handle, and
invoking the \texttt{ready}-callback for the associated Connection.
This does not actually transfer any data to the remote endpoint
and CTaps can therefore not actually guarantee that the remote endpoint
is reachable after invoking \mono{ready()}.

Another piece of logic required to adapt UDP to the structure
of a TAPS library is that surrounding the Listener. A Listener
listens on a single Local Endpoint, creating a new 
Connection each time a new client connects. With UDP not
featuring a handshake, the Listener instead has to keep track
of which remote endpoints it has received messages from, creating
a new Connection whenever it receives a message from a remote endpoint
it has not seen before. This is similar to how TCP connections
are identified by a four-tuple of Source IP address, Source port,
Destination IP address and Destination port. Instead of the
demultiplexing to per-Connection happening as part of the
operating system stack as with TCP, it has to happen as
part of the CTaps custom UDP logic. After a message has
been received by a Listener, the Connection associated with
the sender is found, and any callback for received messages
for this connection is invoked, if registered by the client.
When sending data from a Connection created by a Listener, the
Connection itself contains the destination remote endpoint,
meaning that the Listener does not have to look up any state
in this case.

The implementation of the UDP protocol implementation struct can be seen in
\cref{lst:udp-protocol-impl-struct}.

\subsubsection{TCP}
TCP is one of the three potocols currently supported by CTaps, and has native
support in libuv through the
\mono{uv_tcp_t} handle \cite{libuvLibuvTCPHandle}.

Overall, TCP fits relatively well into the external model and objects
provided by a TAPS API. TCP is conenection-oriented by default, and
therefore maps well to the Connection objects provided by a TAPS.
TAPS does however provide a message-based API, while TCP is stream-based.
For CTaps this means that every time libuv notifies CTaps of having
received new data over a TCP socket, this data is wrapped into a
single message and passed to the application. While the application
can be using a framer to wrap or unwrap received messages, this currently
functions as a proof of concept and does not have an internal buffer
to support combining multiple TCP packets into a single Message
using framers.

Since TCP is also connection-based, listening and creating new sockets
for incoming connections is handled by the TCP stack itself and there is
therefore far less CTaps-specific logic for listening on TCP than there is
on listening over UDP. For example, libuv
has the built-in function \mono{uv_listen}, which invokes a callback whenever
a new connection is received. Since the whole protocol is connection-oriented,
it also means that the logic of only closing the receiving socket when
the last connection closes is built-in, and does not have to be implemented
by CTaps. CTaps still uses a Socket Manager to keep track of the open TCP
handle. For TCP this only tracks the single listener, and can be
closed immediately when the listener is closed. The new TCP handles which
have been spawned from this Listener will have received their own separate
Socket Managers, since they are connected to separate sockets.

The implementation of the TCP protocol implementation struct can be seen in
\cref{lst:tcp-protocol-impl-struct}.

\begin{listing}[tbp]
  \begin{minted}{C}
const ct_protocol_impl_t
    tcp_protocol_interface =
        {
            .name = "TCP",
            .protocol_enum = CT_PROTOCOL_TCP,
            .supports_alpn = false,
            .selection_properties =
                {.list =
                     {
                         [RELIABILITY] = {.value = {.simple_preference = REQUIRE}},
                         [PRESERVE_MSG_BOUNDARIES] = {.value = {.simple_preference = PROHIBIT}},
                         [PER_MSG_RELIABILITY] = {.value = {.simple_preference = PROHIBIT}},
                         [PRESERVE_ORDER] = {.value = {.simple_preference = REQUIRE}},
                         [ZERO_RTT_MSG] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [MULTISTREAMING] = {.value = {.simple_preference = PROHIBIT}},
                         [FULL_CHECKSUM_SEND] = {.value = {.simple_preference = REQUIRE}},
                         [FULL_CHECKSUM_RECV] = {.value = {.simple_preference = REQUIRE}},
                         [CONGESTION_CONTROL] = {.value = {.simple_preference = REQUIRE}},
                         [KEEP_ALIVE] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [INTERFACE] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [PVD] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [USE_TEMPORARY_LOCAL_ADDRESS] = {.value = {.simple_preference =
                                                                        NO_PREFERENCE}},
                         [MULTIPATH] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [ADVERTISES_ALT_ADDRESS] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [DIRECTION] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [SOFT_ERROR_NOTIFY] = {.value = {.simple_preference = NO_PREFERENCE}},
                         [ACTIVE_READ_BEFORE_SEND] = {.value = {.simple_preference =
                                                                    NO_PREFERENCE}},
                     }},
            .init = tcp_init,
            .init_with_send = tcp_init_with_send,
            .send = tcp_send,
            .listen = tcp_listen,
            .close_listener = tcp_close_listener,
            .close_connection = tcp_close,
            .close_socket = tcp_close_socket,
            .abort = tcp_abort,
            .clone_connection = tcp_clone_connection,
            .free_connection_state = tcp_free_state,
            .free_socket_state = tcp_free_socket_state,
            .close_connection_group = tcp_close_connection_group,
            .free_connection_group_state = tcp_free_connection_group_state,
};
  \end{minted}
  \caption{TCP protocol implementation struct}
  \label{lst:tcp-protocol-impl-struct}
\end{listing}

\subsubsection{QUIC} \label{sec:quicImpl}
While providing access to QUIC is one of the most important selling points
of CTaps, QUIC is also by far the most complex of the supported protocols
which can be seen through the required number of lines of code in 
\cref{sec:protocolComplexity}. QUIC also features several features which
do not map cleanly into TAPS, which is discussed further in \cref{sec:quicUnderTaps}.

As QUIC and TAPS was developed relatively concurrently, the TAPS
RFCs feature no protocol considerations or implementations hints like is featured
for several other protocols, despite access to QUIC being one of the
major value propositions of a TAPS system \cite{brunstromImplementingInterfacesTransport2025}. 
There exists a draft for mapping Transport Services to QUIC, but this has
not made it into the official RFC \cite{transportServicesMappingForQuic}.

With the standardization of QUIC being fairly recent,
none of the current implementations of TAPS found during research for this
thesis, such as NEAT or PyTaps, feature QUIC as one of the underlying protocols
\cite{khademiNEATPlatformProtocolIndependent2017}\cite{frankePythonasynciotaps2025}.
As mentioned previously, this is one of the key motivations behind devloping CTaps,
and is a key value proposition.

QUIC is a large protocol and is very configurable. In the development of CTaps,
effort has been made to make this configurability as available as possible to
a client using CTaps, however, since CTaps aims to be a implementation of the
TAPS API, there are ome QUIC-specific configurations which are not available.
These features are discussed further in \cref{sec:quicUnderTaps}.

In the cases where the TAPS API do not map perfectly to QUIC configuration options,
CTaps has tried to follow the principle of least surprise, since the ideal
is that the application should not have to consider which underlying protocol
CTaps has chosen.

Most QUIC configurations are however available through the TAPS API in one of
several ways. 
For example, Application-Layer Protocol Negitiation (ALPN), which is mandatory for QUIC,
is available through a dedicated ALPN Security Parameter.

Since libuv does not feature native QUIC support like it does with TCP and UDP,
CTaps uses Picoquic as an external library, working as a "QUIC engine",
while performing the actual transfer of data over a raw network socket managed
by libuv. A raw socket is necessary since CTaps has to set the from and two
address in the IP packet, to match the IP addresses requested by picoquic.

CTaps is responsible for handling reading and writing data to the network, while picoquic
formats outgoing QUIC packets and handles incoming data.
Picoquic specifies when the application needs to query it next through the
\mono{picoquic_get_next_wake_delay} function. This wakeup time is then passed
to a libuv timer, which queries picoquic on wakeup.
A model of the data flow between CTaps, picoquic and the network can be
seen in \cref{fig:ctaps-quic-loop}.
More details on why exactly picoquic was chosen over other QUIC implementations
can be seen in \cref{sec:picoquic}.

\begin{figure}[ph]
\centering
\begin{tikzpicture}[
    box/.style={draw, thick, rounded corners, minimum width=2.8cm, minimum height=1cm, align=center},
    arr/.style={->, >=stealth, thick},
    lbl/.style={font=\footnotesize, align=center},
]

\node[box] (app)      at ( 0,  4) {Application};
\node[box] (ctaps)    at ( 0,  0) {CTaps};
\node[box] (picoquic) at ( 5,  0) {picoquic};
\node[box] (libuv)    at ( 0, -4) {libuv};

% Application <-> CTaps
\draw[arr] ([xshift=-0.3cm]app.south)   -- ([xshift=-0.3cm]ctaps.north)
    node[lbl, left, midway] {Application data};
\draw[arr] ([xshift= 0.3cm]ctaps.north) -- ([xshift= 0.3cm]app.south)
    node[lbl, right, midway] {received data};

% CTaps <-> picoquic
\draw[arr] ([yshift= 0.3cm]ctaps.east)    -- ([yshift= 0.3cm]picoquic.west)
    node[lbl, above, midway, yshift=0.2cm] {Application data,\\data from socket};
\draw[arr] ([yshift=-0.3cm]picoquic.west) -- ([yshift=-0.3cm]ctaps.east)
    node[lbl, below, midway,  yshift=-0.0cm] {QUIC packets,\\received data};

% CTaps <-> libuv
\draw[arr] ([xshift=-0.3cm]ctaps.south)   -- ([xshift=-0.3cm]libuv.north)
    node[lbl, left, midway] {Send QUIC packet\\via socket};
\draw[arr] ([xshift= 0.3cm]libuv.north) -- ([xshift= 0.3cm]ctaps.south)
    node[lbl, right, midway] {Data received from socket};

\end{tikzpicture}
\caption{Flow of data between CTaps, libuv and picoquic.}
\label{fig:ctaps-quic-loop}
\end{figure}


Picoquic features two primary types of state, the picoquic
context contained in a \mono{picoquic_quic_t} and the per QUIC connection
context contained within a \mono{picoquic_cnx_t}  \cite{huitemachristianPicoquicInternalStructures2020}.
In CTaps, the picoquic context
maps to a single Socket Manager, and can be shared among several QUIC Connections
if they are created from the same Listener. The most important state contained
in the picoquic context is the key files used for the connection and the available
ALPN values. The picoquic connection context is per QUIC connection and therefore
maps to a single Connection Group in CTaps.

When initiating a CTaps Connection over QUIC, CTaps only needs to open a socket on
the required local endpoint. This socket has to be a raw socket, since the libuv
UDP abstraction does not give access to the from and to fields when sending and
receiving data as required by picoquic. CTaps then informs picoquic that it wants
to inititate a new QUIC connection. Picoquic then starts to generate handshake
packets which it passes to CTaps to send over the network. When the QUIC connection
is ready picoquic informs CTaps through a callback, and this callback is then translated
to the appropriate application callback by CTaps.

When opening a Listener, CTaps opens a socket and whenever data is received it is passed
to picoquic. If the received data contains an attempt at creating a QUIC connection
then picoquic will handle the setup and pass the resulting QUIC connection to CTaps.
CTaps can then use this QUIC connection to create an inernal Connection, and invoke
the \mono{ConnectionReady} callback from the Listener.
When closing a Connection, picoquic is informed
through \mono{picoquic_close} and it will wind down the connection, invoking a
callback to CTaps when teardown has completed.

As mentioned previously ALPN is mandatory for QUIC and is supplied to picoquic
by CTaps through a configurable Security Parameter. ALPN is a negotiation
between a QUIC server and client, therefore, both a Listener using QUIC and
a client using QUIC has to take this parameter into account.

The implementation of the QUIC protocol implementation struct can be seen in
\cref{lst:quic-protocol-impl-struct}.


\begin{listing}[tbp]
\begin{minted}{C}
const ct_protocol_impl_t
    quic_protocol_interface =
        {.name = "QUIC",
         .protocol_enum = CT_PROTOCOL_QUIC,
         .supports_alpn = true,
         .selection_properties =
             {.list =
                  {
                      [RELIABILITY] = {.value = {.simple_preference = REQUIRE}},
                      [PRESERVE_MSG_BOUNDARIES] = {.value = {.simple_preference = REQUIRE}},
                      [PER_MSG_RELIABILITY] = {.value = {.simple_preference = PREFER}},
                      [PRESERVE_ORDER] = {.value = {.simple_preference = REQUIRE}},
                      [ZERO_RTT_MSG] = {.value = {.simple_preference = NO_PREFERENCE}},
                      [MULTISTREAMING] = {.value = {.simple_preference = NO_PREFERENCE}},
                      [FULL_CHECKSUM_SEND] = {.value = {.simple_preference = REQUIRE}},
                      [FULL_CHECKSUM_RECV] = {.value = {.simple_preference = REQUIRE}},
                      [CONGESTION_CONTROL] = {.value = {.simple_preference = REQUIRE}},
                      [KEEP_ALIVE] =
                          {.value = {.simple_preference = NO_PREFERENCE}},
                      [INTERFACE] = {.value = {.simple_preference = NO_PREFERENCE}},
                      [PVD] = {.value = {.simple_preference = NO_PREFERENCE}},
                      [USE_TEMPORARY_LOCAL_ADDRESS] = {.value = {.simple_preference =
                                                                     NO_PREFERENCE}},
                      [MULTIPATH] = {.value = {.simple_preference = NO_PREFERENCE}},
                      [ADVERTISES_ALT_ADDRESS] = {.value = {.simple_preference = NO_PREFERENCE}},
                      [DIRECTION] = {.value = {.simple_preference = NO_PREFERENCE}},
                      [SOFT_ERROR_NOTIFY] = {.value = {.simple_preference = NO_PREFERENCE}},
                  }},
         .init = quic_init,
         .send = quic_send,
         .init_with_send = quic_init_with_send,
         .listen = quic_listen,
         .close_listener = quic_close_listener,
         .close_connection = quic_close_connection,
         .close_socket = quic_close_socket,
         .abort = quic_abort,
         .clone_connection = quic_clone_connection,
         .free_socket_state = quic_free_socket_state,
         .close_connection_group = quic_close_connection_group,
         .set_connection_priority = quic_set_connection_priority,
         .free_connection_state = quic_free_state,
         .free_connection_group_state = ct_free_quic_connection_group_state
};
\end{minted}
\caption{QUIC protocol implementation struct}
\label{lst:quic-protocol-impl-struct}
\end{listing}


\subsection{Protocol complexity} \label{sec:protocolComplexity}
As mentioned previously, the protocol implementations in CTaps differ
greatly in complexity. This can be easily seen in the number of lines
of code required to implement the protocol interface for each protocol,
as seen in \cref{tab:protocolLoc}. In addition to the interface implementation
of QUIC being \textasciitilde2.6 larger than that of TCP, QUIC support also requires a fairly
complex external dependency to handle incoming and outgoing data instead of
relying on built-in libuv support.

This increased complexity reflects primarily two things, lower degrees of toolchain
support and an increase in the number of features provided.
For the first point, the complexity to integrate QUIC into CTaps
could likely be greatly reduced if if libuv integrated QUIC support into a custom handle
like it has with UDP and TCP, but with the relative recent development of QUIC,
this has not happened yet. For the second point it naturally follows that since
QUIC offers several features not available through UDP or TCP such as multistreaming
and connection migration, CTaps has to implement logic to make these features available
to the application.

\begin{table}[tbp]
\begin{center}
\begin{tabular}{ ||l|l|c| } 
 \hline
 \textbf{Protocol}   & \textbf{Source files}             & \textbf{Lines of Code} \\ 
 \hline \hline
 UDP                 & \texttt{udp.h \& udp.c}                    & 478 \\ 
 \hline
 TCP                 & \texttt{tcp.h \& tcp.c}                    & 739 \\ 
 \hline
 QUIC                & \texttt{quic.h \& quic.c}                  & 1904 \\
 \hline
\end{tabular}
\end{center}
\caption{Overview of Lines of Code (LoC) on a per-protocol basis}
\label{tab:protocolLoc}
\end{table}

\section{Testing} \label{testing}
A core tenet since the very start of development was that the library should be
as robust as possible. This would to a very large degree achieved primarily through
the use of automated testing, and continious integration.

By writing automated tests
which run on every commit to the repository it is easy to see exactly which change
has lead to a bug, making it far easier to fix. A well-tested library is also an absolute
necessity when trying to attract external dependents, or to integrate it into a larger software
library. This is especially important for large libraries such as CTaps, which abstract away
large amounts of logic away from the application, making it harder to detect bugs
and know exactly where unexpected behaviour stems from.

Tests are first separated into type and then into responsibility. When separating
by type, the main types are integration tests and unit tests. Integration tests perform a larger check on the
whole library and often only mock the system calls, or interact with with a real
server over localhost.

Unit tests perform a smaller check on individual functions, attempting to verify
all logic and that differing branches are hit when expected, based on the input data.
When dividing by responsibility, the testing stack largerly follows the same model
as the main library, with test files separated by responsibility of the original
file or struct that it is performing tests on.

CTaps uses Google Test to make writing tests as easy as possible \cite{GoogleTest2025}.
This is a C++ library which can easily be integrated with a C codebase using 
\mono{extern "C"} linkage. Google Test makes a rich set of features such as
test fixtures for standardizing behaviour for groups of tests. This is especially
useful for CTaps integration tests where a single test fixture can be used to
standardize how to start and teardown servers on localhost, which are
required for several tests.

A central concept when writing good tests is to isolate the component being tested from
external libraries. Both to limit the scope of our testing, but also to potentially
simulate a failure to 
check that CTaps handles an external library failure properly.
All of these concerns are best met by
mocking.

Mocking is the concept of defining a custom behaviour for an external library
or dependency when testing, instead of invoking the real library. By using this we can
simulate the full range of behaviours from the external library and verify that our application
reacts correctly, while also avoiding time-consuming API calls, making unit tests
significantly faster.

For mocking, CTaps uses FFF, which is a small open-source
framework for creating fake C functions when testing \cite{longFakeFunctionFramework2025}.
FFF can either mock the external library functions, such as \texttt{uv\_getaddrinfo},
which is the built in support for DNS lookup in libuv, or mock a call to a different
function, which either contains a call to an external library, or which we are simply
not interested in testing.

Mocking with FFF is especially useful when combined with wrapping
symbols in the linker \cite{gnuLinkerManual}. This can be used to redirect all invocations of a funciton
to a mock at link time, greatly simplifying the building of tests since symbol
ambiguity no longer is an issue.

An example of function mocking using FFF can be seen in \cref{lst:example-fff-mocking}.

\begin{listing}[tbp]
\begin{minted}{Cpp}
DEFINE_FFF_GLOBALS;
FAKE_VALUE_FUNC(int, uv_udp_bind, uv_udp_t*, const struct sockaddr*, unsigned int)
FAKE_VALUE_FUNC(int, uv_udp_recv_start, uv_udp_t*, uv_alloc_cb, uv_udp_recv_cb)

TEST(InitiationTests, respectsLocalEndpoint) {
    uv_udp_bind_fake.return_val = 0;
    uv_udp_recv_start_fake.return_val = 0;
    
    ... Rest of code which calls uv_udp_bind and uv_udp_recv_start ...
    
    // Assert that the mocked functions were invoked with the expected arguments    
    EXPECT_EQ(1, uv_udp_bind_fake.call_count);
    EXPECT_EQ(1234, htons(((struct sockaddr_in*)uv_udp_bind_fake.arg1_val)->sin_port));
    ... Rest of the assertions ...
}

\end{minted}
\caption{Example test mocking udp functions, using FFF}
\label{lst:example-fff-mocking}
\end{listing}

\section{Build System} \label{sec:build-system}
CTaps is built using CMake and follows conventions around build
and structure used by libuv
in order to make the structure as familiar as possible. This includes
having the API consist of a single header file, \mono{ctaps.h} which
is included to expose all functionality intended to be used by an application.

CTaps is also compiled with visibility of all symbols set to hidden by default,
with only the symbols marked as \mono{CT_EXTERN} being available in the final
shared library file. This is done to hide internal state from applications,
making it clear which parts of the compiled object file are part of the
external API and not.

By default CTaps is built as a shared library. This means that the library is compiled
into a single shared object file. When compiling an Application using CTaps,
this shared object file is not included in the compiled binary, instead the
compiled binary looks for a shared object file at a given location at
run time. This means that any Application depending on CTaps will by default
not include CTaps in its own binary, leading to a smaller file size, although this 
is naturally dependent on the build configuration of the application. 

The largest downside of a a dynamically
linked Application is that it requires its users to have CTaps installed on their
systems, in order for the Application find the correct symbols at run time.

In addition to the smaller binary size, the largest upside of a dynamically
linked application is that it is far easier to update its dependencies.
If a new version of CTaps was released then an application could update
its dependency by swapping the local shared object file without having
to recompile the entire application.

If the Application was statically linked
then it would have to update its dependency and recompile. This would also mean
that the entire application would have to be redistributed instead of simply
having users update their CTaps installation whenever needed.

This is especially relevant for networking applications where security
is a large concern. If a criticial security error was found in CTaps
and it was statically linked
then every user of downstream application would have to update every application
using CTaps. If it was dynamically linked, then each user would only have to update
a single shared object file.

As a limitation of scope in the original thesis description, CTaps is only
targeted towards Linux. However, since I/O and other system calls are handled
by libuv which does support windows, CTaps is likely very close to being
Windows-compatible, but this has not been tested \cite{Libuv2025}.

The library is distributed through GitHub \cite{cTaps}. No binaries
are compiled by default, but the project can be downloaded directly
in CMake using FetchContent, as exemplified in the CTaps example project
\cite{ctaps-example-project}.

\section{Dependencies}
As with most libraries, CTaps has several external dependencies
used to simplify the logic required to be implemented within CTaps itself.
These are again divided into statically and dynamically linked dependencies.

A description of the difference between statically linked and dynamically linked
dependencies can be found in \cref{sec:build-system}. CTaps statically links against
picoquic since it is a library very specific to CTaps, while it uses dynamic linking for
more general dependencies which it expects application developers to have present
on their machine. This includes GLib and libuv.

Care has been taken to keep dependencies away from the external API. This means
that when compiling CTaps, dependencies are required to be present on the compiling
system, while downstream applications only require the runtime libraries  to be present.
For dynamically linked dependencies such as
libuv and GLib, this means the shared libraries must be installed on any system
running an application built against CTaps.

\subsection{GLib}
As mentioned in \cref{sec:glib-design-reason}, CTaps uses GLib 
for certain internal data structures to avoid having to re-implement them
internally.
This includes the candidate tree when performing candidate gathering
and hash tables keeping track of Connections associated with a Connection Group.
