\part{Introduction}
\chapter{Background}
Anyone programming any kind of networking application in a UNIX
environment are bound to come across sockets as the dominant API.
Sockets have been around, in more or less their current form, since
the 1980s \cite[p.~xx]{stevensUNIXNetworkProgramming2013}. The
initial mention of sockets in relation to ARPANET comes from 
RFC 33, published in 1970 \cite{crockerNewHostHostProtocol1970}.

Network sockets have grown alongside with the growth of UNIX. 
Network sockets have grown out of the Berkeley sockets
which first appeared in version 4.2 of the Berkeley Software
Distribution, originally used for inter-process communication.
With their
early appearance for inter-process communication in the Berkeley 
Software Distribution version 4.2
in 1983, sockets have grown to become the dominant way low-level
network programming is done today. [citation]
As mentioned previously, the original Berkeley sockets were designed
for inter-process communication. As a continuation of this, the
implementation of network sockets can be either connection
oriented or stream-oriented, a setting which is decided at compile
time. This has propagated into network sockets, for which the type of
underlying connection, either TCP or UDP, has to be decided at
compile time. 

Being stream-oriented or message-oriented has to do with
how the transferred data is represented. In a stream oriented
protocol such as TCP, the data is represented as a stream of bytes.
From the application's point of view, the data arrives continuously
and not in discrete chunks. This means that the appliaction cannot
easily differentiate between different messages. As everything is
ran over IP, which is a message-oriented protocol, TCP, which
is a transport layer protocol, has to be able to handle receiving
discrete messages and gluing them together into a stream of bytes.
This also includes handling messages arriving out of order, which
has to be handled before they can be presented as a stream in order.
This is called in-order delivery and is a key feature guaranteed
to the client by TCP.

With message-oriented protocols, such as UDP, the data is represented
as discrete messages. This means that the application can
easily differentiate between them. Because the data is represented
in discrete messages, this allows the protocol to not have do deliver
the messages in order. Firstly, this allows the application to
access the data quicker, as it does not have to wait for any
delayed packets to arrive, before the other data can he handed over.
This becomes very important when multiplexing multiple streams of data
over a single connection. This is done in order to more efficiently utilize
bandwith, but faces problems if data from one stream is delayed if
we are using a protocol which guarantees in-order delivery. Even if
only one application's data is delayed, the second application's data
will have to wait, as they are together treated as a single stream
and have to be delivered in order. This is called head-of-line blocking.
When using a message-oriented protocol, which makes it possible,
but not mandatory, to not guarantee in-order delivery, one application
does not have to wait for the other application's data.


This compile-time binding may not seem like a large problem,
as it reflects the fact that a connection cannot swap protocols
midway through. And the type of protocol is usually deeply bound
to the application, and large redesigns likely have to take place if
an application were to change from using a connection-oriented
to a stream-oriented connection. However, different parts of
the application may have different requirements, for example
a connection to a video-streaming website itself may be 
connection-oriented, while the video-stream itself will likely
be stream-oriented.

One of the key features of TAPS is to give the developer access to more
advanced features of protocols such as QUIC or SCTP, while still being
able to fall back to TCP if necessary. This will allow TAPS to act in
a message-oriented way, while still being able to use a connection-oriented
protocol such as TCP.

One of the key selling points for TAPS is the ability to use different
network protocols with the same API. This will for example ease the
development of applications that for example use QUIC. 
QUIC is a internet protocol developed by Google, released to
the public in 2013, described in RFC 9000 
\cite{iyengarQUICUDPBasedMultiplexed2021}.
The key advantages of QUIC over TCP is that it is designed to be
more efficient thatn TCP, and be able to handle more streams of
data in a better way due to the lack of head-of-line blocking.
QUIC is implemented in user space on top of UDP, but acts as a
transport-level protocol. It is designed on top of UDP, in order
to avoid issues with middleware dropping packets which are not
TCP or UDP.

As mentioned previously, one of the key selling points of
TCP is the ubiquity of sockets, which are implemented in kernel space.
There is no equivalent API for QUIC, and with a relatively advanced
implementation, users have to settle for one of the existing
libraries. 

Currently, few applications can be built
exclusively on top of QUIC, as it requires support from both the client
and the server. QUIC is a transport protocol that is built on top of
UDP, and is designed to be a replacement for TCP. QUIC uses
multiplexed connections to for more efficient use of the network.
Multiplexed connections are also used in http/2, but there it
suffers from head-of-line blocking, which is caused by TCP's 
requirement for all packets to be delivered in order. This problem
is exacerbated when using multiple streams of data over a single
TCP connection, as a single lost packet will block all other
streams.

With QUIC, the multiplexed connections can still be sent
over a single connection, but the streams are independent of each other.
This is made possible because QUIC is built on top of UDP; which
does not guarantee packet ordering for separate streams.

TAPS exposes several configurations which can be used to specify
a preference of requirement, which then sets a preference or a
requirement for which underlying transport protocol to use.
These configurations come in the form of transport properties.
These properties are broken down into three categories:
Selection properites, Connection properties, and message
properties. Selection properties are requirements for the
underlying transport protocol, which TAPS uses to select
which protocol to use for the connection. Connection properties
are properties which are set on a per-connection basis, for
example the timeout for aborting a connection or the connection
priority. Message properties are properties which are set
on a per-message basis, for example the priority or the lifetime
of the given message. A full list of properites can be found
in the Architecture and Requirements for Transport Services RFC
\cite{paulyArchitectureRequirementsTransport2025}.

Selection properties can, as mentioned above, be specified
to varying degrees of strictness. With five different values
\texttt{Require}, \texttt{Prefer}, \texttt{No Preference}, \texttt{Avoid},
and \texttt{Prohibit}. Here, \texttt{Require} means that establishing
the connection will fail if the property cannot be satisfied with the
available transport protocols. \texttt{Prefer} means that the connection
can fall back to a protocol not supporting the property, if no protocol
satisfying the property is available. \texttt{Avoid} and \texttt{Prohibit}
function in the same way as \texttt{Require} and \texttt{Prefer},
but selecting for protocols which do not have the selected property.
A simple example of a selection property is selecting for reliable
data transfer. If this was set as required, TAPS would be forced
to use for example TCP or QUIC as the underlying protocol. If UDP
was the only available protocol, establishing the connection would fail.
If the user selected \texttt{prefer}, then UDP could be selected if
nothing else was available, even though it does not fulfill the property.

There are also more advanced selection properties available. Out
of these, the least intuitive are \texttt{Multistream Connections in Group},
\texttt{Full Checksum Coverage on Sending} and \texttt{Full Checksum Coverage on Receiving},
\texttt{Provisioning Domain Instance or Type}, \texttt{Use Temporary Local Address}, and
\texttt{Notification of ICMP soft error message arrival}.

The \texttt{Multistream Connections in Group} selection property allows
the user \todo{TODO - select term for "user" of TAPS api (developer of networking application)}
to select whether the application would prefer multiple Connections within a Connection Group 
to be provided by streams of a single underlying transport connection where possible
\cite{paulyArchitectureRequirementsTransport2025}. Multistreaming in this context is
the act of providing multiple streams of data over a single underlying connection. This is
advantageous as it allows for more efficient use of the network, since the connection
will better saturate the available bandwidth. This avoids long convergence times
for the connection and allows fresh connection to utilize the large congestion window
already achieved, without having to wait for slow start to ramp up. This is especially
useful for applications with many small streams of data, such as web browsers.

The \texttt{Full Checksum Coverage on Sending} and \texttt{Full Checksum Coverage on Re\-ceiving}
selection properties
specify the application's need for protection against corruption for all data transmitted or
received on a given connection. This is used to specify how many bytes of the message
which needs to be protected by a checksum. This checksum is used to protect the data
from corruption caused by lower-layer protocols. The setting \texttt{Full coverage} means
that the entire message is protected, while \texttt{0} specifies no checksum is required.
If any other value than \texttt{Full coverage} is specified, the connection may still
fall back to \texttt{Full coverage}.

The \texttt{Provisioning Domain Instance or Type} selection property 
allows the applicatio \todo{common application term}n to
control path selection by selecting for specific Provisioning Domain or
categories of domains \cite{paulyArchitectureRequirementsTransport2025}.
Provisioning Domains define consistent sets of network properties that 
might be more specific than network interfaces 
\cite{paulyArchitectureRequirementsTransport2025}.

The \texttt{Use Temporary Local Address} selection property
allows the application \todo{common application term} to
specify whether or not to use temporary addresses which are a
feature in IPv6, used for greater privacy 
\cite{gontTemporaryAddressExtensions2021},\cite{paulyArchitectureRequirementsTransport2025}.

The \texttt{Notification of ICMP soft error message arrival}
selection property lets an application \todo{common application term} specify whether or
not it should be notified when an ICMP error, which does not
force termination of the connection, arrives 
\cite{paulyArchitectureRequirementsTransport2025}.

In addition to specifying which underlying protocol should
be used through selection properties, the application \todo{common application term}
using TAPS can manage the connections themselves through transport
properties. These can be specific for a given transport protocol
such as the selection properties, or generic and be available for
all transport protocols. \todo[]{Should this be separated from all the above
parts? are they all part of the same 'category'}

The \texttt{Connection Group Transmission Scheduler} transport
property decides which scheduler is used within a given connection
group, to divide the available capacity between the connections.
\cite{paulyArchitectureRequirementsTransport2025}.

The \texttt{Capacity Profile} transport property is used
to select a capacity profile, which will be a tradeoff between 
delay, delay variation, and efficient use of the available capacity 
based on the capacity profile specified
\cite{paulyArchitectureRequirementsTransport2025}.

The \texttt{Isolate Session} transport property set in order
to have new connections use as little cacheed information as possible
from previous connections not in the same connection group
\cite{paulyArchitectureRequirementsTransport2025}.

There are already several working TAPS implementations, but they are lacking
in different ways. Two prominent implementations are \texttt{dynamic-taps}
implemented in C 
and \texttt{python-asyncio-taps} implemented in Python 
\cite{dukeDynamictaps2025}\todo{Is this actually the one Michael was talking about, what is NEAT?}
\cite{frankePythonasynciotaps2025}. This thesis
aims to be an improvement over these implementations. By being a C implementation
it will be suitable for lower level applications. It aims to improve on the
existing C implementation by being more modular and easier to use. Currently,
\texttt{dynamic-taps} is a monolitic implementation \todo{is this actually true}
and has not been updated since 2021. This thesis will provide an alternative
written after the final TAPS RFC has been published. It will also be more
modular and well tested, making it easier to use an to integrate into an
existing library or application.

\textbf{Write about \texttt{Python alternative} \texttt{Old bloated
C alternative} etc.}
    

\sout{Things we need to mention:}
\begin{itemize}
    \item \sout{Sockets are dominant}
    \item Why are they dominant?
    \item Advantages of sockets
    \item Disadvantages of sockets
        \begin{itemize}
            \item \sout{Compile-time binding to transport layer protocols}
            \item Lack of flexibility
        \end{itemize}
    \item What is QUIC?
        \begin{itemize}
            \item Why QUIC?
            \item Advantages of QUIC
            \item Disadvantages of QUIC
        \end{itemize}
    \item What is TAPS?
        \begin{itemize}
            \item Why TAPS?
                \begin{itemize}
                    \item When was TAPS developed?
                    \item Who developed TAPS?
                    \item What is the goal of TAPS?
                \end{itemize}
            \item Advantages of TAPS
                \begin{itemize}
                    \item \sout{Flexibility}
                    \item \sout{Transport protocol agnostic}
                    \item \sout{switching between transport protocols}
                    \item utilizing QUIC
                \end{itemize}
            \item Disadvantages of TAPS
            \item Specific TAPS features
        \end{itemize}
    \item Why does this thesis exist
        \begin{itemize}
            \item Existing implementations
            \item Problems with existing implementations
            \item Advantages of this implementation
        \end{itemize}
\end{itemize}



\section{Past work}

