\chapter{Design} \label{chap:design}
This chapter will cover the design decisions which has affected
CTaps, both internally and for the external API. The overarching
design goal was to conform to the API described in RFC~9622~\cite{rfc9622},
while exploring how to best expose as much QUIC functionality as possible
through the TAPS API. The chapter first describes the overall scope
of CTaps, before expanding on the chosen dependencies, internal architecture
and the external API presented to applications.

\section{Choice of Scope}
One of the key research questions motivating the development of CTaps is exactly
how well the TAPS API as it exists currently can be presented programmatically.
CTaps therefore aims to present as much of the TAPS API as possible, despite
implementing several features as no-ops.

As it stands now, CTaps implements a large subset of the TAPS API and presents
several of the most beneficial features of the TAPS API such as candidate gathering
and candidate racing, as well as support for TCP, UDP and QUIC. In addition
it presents all of the core objects of the TAPS API, such as Connections, Listeners and
Connection Groups. It allows the application to set all Transport Properties listed
in RFC~9622, although not all properties are supported internally and setting them has no effect.
When performing
racing the most desirable protocols are attempted first and staggered racing
is performed for less desirable protocols, as described in RFC~9623
\cite{brunstromImplementingInterfacesTransport2025}.
\todo{Check in on this later, describe how large of a subset}

The parts of the TAPS API which were prioritized for CTaps are common features related
to establishing client and server connections, transferring data and managing
the lifetime of Connections. The highest prioritized internal TAPS features
were the most unique parts of the TAPS API, which only a software library
presenting a generic interface to the transport layer could present. This
is naturally candidate gathering and candidate racing, which allows a single
client to be able to connect to multiple different endpoints using different
protocols.

Certain parts of the TAPS API are not present in CTaps, the
largest of which is the peer to peer connection functionality.
Implementing this was deemed to be out of scope, while presenting
the API as a no-op was deemed to be of little value, since CTaps would not be able to predict
how the internal rendezvous implementation would affect the external API.

Other functionality which is not present in the CTaps API is, among others,
being able to add endpoints to an active Connection.
This is omitted because an actual implementation was deemed
out of scope, and presenting it as a no-op was deemed to be of little value.

Overall, CTaps serves as a reference
to show the architectural and design decisions which are necessary when
implementing the TAPS API, as well as to show which features of QUIC can
be exposed through TAPS.

\section{Choice of dependencies}
\subsection{libuv} \label{sec:dependency-libuv}
Since TAPS is an asynchronous API at its core, CTaps needed to
reflect this. This could have been achieved in several ways,
with the primary alternatives being internal threading or an event loop. For
CTaps, the event loop was found to have several advantages,
primarily avoiding race conditions and not having to guard shared
state, since a single threaded event loop ensures that state
is never accessed concurrently.

When selecting a specific event loop library libuv~\cite{Libuv2025} presented
several advantages such as built-in protocol handles for TCP
and UDP, in addition to being a mature library which is already
used in production systems such as Node.js~\cite{nodejsEventLoop}.
Libuv is also the event loop engine behind
NEAT, which influenced the decision of CTaps~\cite{neatCode}.

Internally in CTaps, libuv provides the asynchronous interface
which CTaps uses to interact with the network. It also provides
a timer, which is used to query Picoquic at given intervals
for new packets to send.

\subsection{Picoquic}
CTaps has been developed on Linux kernel version 6.1.0, which
features no kernel-level QUIC support.
Since QUIC is also not supported natively by libuv an external library
was needed to add QUIC support to CTaps.
For this the largest priority was ease of use, in order
to assist in exploring how to QUIC features map onto TAPS.

Picoquic~\cite{christianhuitemaPicoquic} was therefore chosen because
of its simplicity and was
prioritized over larger and more performance-oriented libraries such as
MsQuic~\cite{MsQuic} from Microsoft.

A large advantage of Picoquic is that it features two modes of operation.
It can either be used independently with its own event loop, or be
integrated with an existing one by querying at given timer intervals.
The first mode allowed the benchmarking to be performed with a Picoquic
client completely independent of CTaps, while the second mode allowed Picoquic
to integrate with libuv without having to manage two event loops.

\subsection{GLib} \label{sec:glib-design-reason}
CTaps uses GLib~\cite{GLibGitLab} for implementations of
hash maps, linked lists, trees and arrays.

GLib was added as a dependency since C has a small standard library
which does not feature any of the data structures mentioned.
If CTaps were to implement all the data structures internally
it would create a large surface area liable for bugs requiring
extensive testing, increasing the scope of CTaps.

GLib itself was chosen because it is a large and mature library
which is featured in several other large well known software libraries,
such as GTK~\cite{gtk_website}.

\subsection{log.c}
In order to create structured logs which can
easily be filtered by level
CTaps uses log.c~\cite{logc}.
log.c was specifically chosen because of its simplicity
and pure C implementation, making it easy to integrate
with CTaps.

\section{Architecture of CTaps} \label{sec:ctaps-arch-design}
As mentioned, CTaps aims to conform to the asynchronous TAPS API described
in RFC~9622~\cite{rfc9622}. Since CTaps is asynchronous it has
to rely on the result of any asynchronous operation being passed via
callback, which affects the look of the external API and
the internal architecture.

CTaps features a shared protocol API internally and is therefore extensible
when it comes to adding support for new protocols.
In order to keep the implementations as small as possible, common logic such as
state tracking
is extracted to the Socket Manager, which is an internal data structure
responsible for network sockets.

An overview of the architecture of CTaps can be seen in \cref{fig:ctaps-arch}.
Here, all the core aspects of CTaps can be seen, with the Socket Manager
acting as an intermediary between the CTaps API and the
protocol interface, which provides a common API to the different
protocol implementations. The unique consideration of QUIC can
also be seen, requiring Picoquic as an external library due to lack
of support in libuv.

This section will now expand on these key concepts which have shaped the
architecture of CTaps and the reasoning behind each of them.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=0.5cm,
    layer/.style={
        draw, 
        line width=0.8pt, 
        minimum width=8cm, 
        minimum height=1.2cm, 
        align=center, 
        fill=gray!5,
        font=\sffamily\bfseries
    },
    sublayer/.style={
        draw, 
        line width=0.6pt, 
        minimum width=3.8cm, 
        minimum height=1cm, 
        fill=white,
        font=\sffamily\small
    }
]

% 1. User Application Layer
\node [layer, fill=blue!10] (app) {Application};

% 2. CTaps API Layer
\node [layer, below=of app] (api) {CTaps API\\ \mdseries\small (TAPS interface, Callback Registration)};

% 3. Socket Manager / Abstraction
\node [layer, below=of api] (manager) {Socket Manager\\ \mdseries\small (Owns network socket)};

% 3. Socket Manager / Abstraction
\node [layer, below=of manager] (interface) {Protocol Interface\\ \mdseries\small (Common interface to network protocols)};

% % 4. Implementation Engines (Split)
\node [sublayer, below=0.5cm of interface.south west, anchor=north west] (uv) {libuv \mdseries\small (TCP/UDP)};
\node [sublayer, below=0.5cm of interface.south east, anchor=north east] (pico) {Picoquic \mdseries (QUIC)};

\node [sublayer, below=0.5cm of pico.south east, anchor=north east] (libuvraw) {libuv \mdseries (raw UDP socket)};

% % 5. OS Layer
\node [layer, below=1cm of libuvraw.south west, fill=gray!20] (os) {Network Stack};

% Connectors
\draw[->, thick] (app) -- (api);
\draw[->, thick] (api) -- (manager);
\draw[->, thick] (manager) -- (interface);
\draw[->, thick] (interface.south) -- ++(0,-0.15) -| (uv);
\draw[->, thick] (interface.south) -- ++(0,-0.15) -| (pico);
\draw[->, thick] (uv) -- (uv |- os.north);
\draw[->, thick] (pico) -- (pico |- libuvraw.north);
\draw[->, thick] (libuvraw) -- (libuvraw |- os.north);
\end{tikzpicture}
\caption[CTaps Architecture Overview]
{
An overview of the architecture of CTaps, featuring the Socket Manager as common as a
mediator between the external functions and the API, with the common protocol interface 
and how it interacts with Picoquic and libuv.
}
\label{fig:ctaps-arch}
\end{figure}

\subsection{Mapping QUIC to TAPS} \label{sec:quic-arch-mapping}
While the TAPS RFCs were designed with QUIC in mind, they were written before the
QUIC design was finalized in RFC~9000~\cite{rfc9000}. This means that RFC~9623 does not
feature any specific transport protocol considerations for mapping QUIC into
a TAPS system~\cite{brunstromImplementingInterfacesTransport2025}.
There exists a draft for mapping Transport Services to QUIC, but this has
not made it into the official RFC~\cite{transportServicesMappingForQuic}.

The most important design decision made in CTaps in regards
to QUIC was to map a QUIC Connection
to a Connection Group instead of to a Connection. Since
Connection Groups are populated by Connection cloning, and RFC
9622 says \enquote{if the underlying protocol supports multi-streaming, it is natural to use this functionality to
implement \mono{Clone}}~\cite[Sec. 7.4]{rfc9622}, the Connection Group
becomes the natural model for the QUIC Connection.

CTaps could
map QUIC connections to a TAPS Connection, but this would make
multistreaming unavailable through the TAPS API, since TAPS
features no equivalent concept of a stream within a Connection.
Multistreaming is a core benefit of QUIC and allows for more
efficient bandwidth usage, as showcased in \cref{sec:smallfiletransfer}.

The consequence of this mapping is that the behavior 
when closing a single Connection depends on the underlying
protocol. For TCP, closing a CTaps Connection will initiate
teardown of the entire connection, and no more data will be
delivered after the \mono{closed()} callback has been invoked.

Since QUIC features no way of gracefully closing
a single stream other than both peers sending a FIN flag,
CTaps has to let a Connection remain open to receiving data
after an application has invoked \mono{closed()} until the
peer has indicated that it has finished writing, which breaks
with the expected behavior from TCP.

This differing behavior can be mitigated by only interacting
with Connection Groups when wanting to shut down a logical connection,
which is compatible with how TCP support is implemented.
This is discussed in detail in \cref{sec:quicUnderTaps}.

\subsection{Socket Manager} \label{sec:socket-manager-design}
The Socket Manager is an internal structure in CTaps responsible
for acting as an asynchronous intermediary between the API and the protocol implementations,
as well as managing the lifetimes of the underlying sockets.

The Socket Manager has to be compatible with both multiple unrelated
Connection Groups using a single Socket, such as when listening
on an UDP socket, as well as Connections in the same Connection
Group using different sockets, such as when cloning TCP connections.
Therefore, the Socket Manager is
referenced on the Connection level and a single shared Socket Manager is created
for each used Socket.
Since UDP Connections in CTaps can share a single socket the 
Socket Manager is also responsible for demultiplexing incoming messages,
as UDP is not connection-oriented.

Since several functions within the various protocol implementations
are asynchronous, communication between the protocol implementation
and the Socket Manager has to be asynchronous as well. The Socket
Manager features a structure with internal callbacks which are invoked by the protocol
implementation at appropriate times, for example to notify
the Socket Manager that a Connection has been closed after teardown
has completed. This callback structure can be seen in \cref{lst:socket-manager-callbacks}.

This pattern of the protocol implementation exclusively speaking
to the Socket Manager only applies to memory-critical paths, in order
to mitigate
potential memory leaks.

An example of such a memory-critical internal flow in CTaps can be seen in
\cref{fig:async-close-flow}, showing the asynchronous flow of
graceful connection teardown for a Connection using a shared
Socket Manager. Here, a single Connection in a Connection Group
is closed, and the Socket Manager invokes the application callback
asynchronously.

The figure also shows that crucially the Socket Manager keeps
the socket open, so that it can be used by the remaining Connections.

\definecolor{ConnectionGroupTextColor}{RGB}{139, 194, 110}
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\sffamily,
    block/.style={rectangle, draw, text centered, rounded corners, fill=white, font=\bfseries},
    client/.style={rectangle, draw, text centered, fill=white, font=\small},
    arrow/.style={-Stealth, thick},
    asyncArrow/.style={arrow, dashed, red!70!black},
    label/.style={font=\scriptsize\itshape}
    ]

% Place connections first
\node[client] (Conn1) at (0,0) {Connection 1};
\node[client, below=0.3cm of Conn1] (Conn2) {Connection 2};
\node[client, below=1.2cm of Conn2] (ConnN) {Connection N};

% SM centred vertically on the connections group
\node[block, align=center] (SM) at
    ($(Conn1.north east)!0.5!(ConnN.south east) + (4,0)$)
    {Socket Manager};

\node[block, right=of SM] (Prot) {Protocol Impl};

\draw[arrow] (Conn1.north east) -- node[above, sloped, label, yshift=3pt] {close()} (SM.north west);
\draw[arrow] (Conn2.east) -- (SM.west);
\draw[arrow] (ConnN.east) -- (SM.south west);

\node[font=\Huge, text=gray] at ($(Conn2.south)!0.4!(ConnN.north)$) {$\vdots$};

\begin{scope}[on background layer]
    \draw[rounded corners, fill=blue!5!white, draw=blue!20]
        ($(Conn1.north west)+(-0.3, 0.4)$) rectangle ($(ConnN.south east)+(0.3,-0.3)$);
    \node[font=\normalsize\bfseries, above=0.4cm of Conn1, blue!70!black, align=center]
        {Connections};
    \node[fit=(Conn1)(Conn2), rounded corners, fill=green!5!white,
          draw=green!20, inner sep=0.13cm] (GreenBox1) {};
    \node[fit=(ConnN), rounded corners, fill=green!5!white,
          draw=green!20, inner sep=0.13cm] (GreenBoxN) {};
\end{scope}
\node[font=\tiny\bfseries, text=ConnectionGroupTextColor, above=-1.2mm of GreenBox1] {Connection Group 1};
\node[font=\tiny\bfseries, text=ConnectionGroupTextColor, above=-1.2mm of GreenBoxN] {Connection Group N};

% Request Flow
\draw[arrow]
    ([yshift=-0.5mm]SM.north east)
    -- node[above, sloped, label, align=center] {Close connection}
    ([yshift=-0.5mm]Prot.north west);
\draw[asyncArrow] 
    ([yshift=0.5mm]Prot.south west)
    -- node[below, label, text=red!70!black, align=center] {Socket Manager cb}
    ([yshift=0.5mm]SM.south east);
\draw[asyncArrow]
    ([yshift=4pt]SM.west)
    -- node[below, sloped, label, text=red!70!black, align=center] {User cb}
    ([yshift=3pt]Conn1.east);
\end{tikzpicture}
\caption[Flow of internal callbacks in CTaps.]{Example model of the asynchronous communication
between the CTaps Socket Manager and the protocol implementation when an application invokes \mono{close()}}
\label{fig:async-close-flow}
\end{figure}

\begin{listing}[htbp]
\begin{minted}{C}
typedef struct ct_socket_manager_callbacks_s {
void (*closed_connection)(ct_connection_t* connection);
void (*aborted_connection)(ct_connection_t* connection);
void (*establishment_error)(ct_connection_t* connection);
void (*connection_ready)(ct_connection_t* connection);
void (*message_sent)(ct_connection_t* connection, ct_message_context_t* message_context);
void (*message_send_error)(ct_connection_t* connection, ct_message_context_t* message_context,
                           int reason_code);
void (*socket_closed)(struct ct_socket_manager_s* socket_manager);
void (*connection_received)(ct_listener_t* listener, ct_connection_t* connection);
void (*listener_ready)(ct_listener_t* listener);
void (*closed_listener)(struct ct_socket_manager_s* socket_manager);
} ct_socket_manager_callbacks_t;
\end{minted}
\caption[Socket Manager Internal Callbacks]{The internal callbacks used for communication between
the Socket Manager and the protocol implementation.}
\label{lst:socket-manager-callbacks}
\end{listing}

\subsection{Protocol Interface}
In order to make the CTaps design as modular as possible,
CTaps uses an internal interface to interact with the underlying protocol.
This interface abstracts away the implementation details of the individual
protocols and allows the Socket Manager to not have to implement protocol
specific logic, making it far easier to extend support to more protocols in
the future.

An excerpt of the protocol interface struct can be seen in
\cref{lst:protocol-impl-struct}. Here the \mono{supports_alpn} flag
and the Selection Properties can be seen. These are used
to encode the capabilities of the protocol when defining an
instance of the interface. The encoded capabilities 
allows CTaps to filter and sort protocols according to the
Transport Services desired by the application, while
the implemented function pointers present a generic
interface to protocol-specific behavior.

\begin{listing}[htbp]
\begin{minted}{C}
typedef struct ct_protocol_impl_s {
    ///< Protocol name (e.g., "TCP", "UDP", "QUIC")
    const char* name;
    ///< Protocol enumeration value
    ct_protocol_enum_t protocol_enum;
    ///< True if protocol supports ALPN negotiation
    bool supports_alpn;
    ///< Properties supported by this protocol
    ct_selection_properties_t selection_properties;
    /** @brief Initialize a new connection using this protocol. */
    int (*init)(ct_connection_t* connection);
    /** @brief Send a message over the protocol. */
    int (*send)(ct_connection_t*, ct_message_t*, ct_message_context_t*);
    // ... Remaining protocol interface function pointers ...
} ct_protocol_impl_t;
\end{minted}
\caption{Excerpt from the protocol implementation interface}
\label{lst:protocol-impl-struct}
\end{listing}


\section{API Design}
The external API of CTaps is designed according to the TAPS API
described in RFC~9622~\cite{rfc9622}.
The TAPS API is object oriented and asynchronous, which are the core
aspects of CTaps as well. In addition, CTaps hides
internal state from the user and only exposes getters and setters
to fields which are meant to be public.

These concerns, in addition to a consistent ownership model which helps
prevent memory leaks, are the largest influences on the external
CTaps API. This section will describe these design decisions,
while a full API overview, which objects are available and how they
map to TAPS can be seen in the attached Doxygen output in the
appendix. \todo{Add to appendix}

\subsection{Object Orientation}
Since TAPS itself is object oriented CTaps aims to emulate this.
This affects how CTaps presents data structures to external application
and how it structures them internally. An example can be 
seen with the Remote Endpoint object in \cref{lst:remote-endpoint}.
This object contains all the context related to remote endpoints, which then is used
in Connections and Listeners. The object has dedicated functions to operate
on it, and therefore encapsulates all responsibilities related to Remote Endpoints.

CTaps aims to present objects as consistently as possible to the application.
An example of this is that 
in CTaps every Connection belongs to a Connection Group,
even if it has not been used for cloning. This
was done in order to both present a unified interface to applications using
CTaps. This also allows CTaps to avoid internal checks for whether a Connection was
part of a Connection Group or not.

In order to hide internal state from applications, the public CTaps API only
features forward declarations of objects. This means that the application
cannot interact directly with the struct members and instead must
interact with the objects through a list of functions exposed by CTaps,
which are used to allocate a struct on the heap and to modify the object using
its pointer, a pattern which can be seen in \cref{lst:using-remote-endpoint}.

This pattern was chosen in order to make it more clear exactly which fields it is
allowed to interact with, and which fields are only used internally in CTaps. This
also allowed CTaps to differentiate between read-only Connection Properties and writable
Connection Properties as described in the TAPS API~\cite[Sec. 8.1.11]{rfc9622}.

\begin{listing}[htbp]
\caption[Remote Endpoint Structure]
{The internal definition of a Remote Endpoint in CTaps, with example functions from the CTaps API.}
\begin{minted}{C}
typedef struct ct_remote_endpoint_s {
    ///< Port number set by user
    uint16_t port;
    ///< Service name (e.g., "https") or NULL
    char* service;
    ///< Hostname for DNS resolution or NULL
    char* hostname;
    ///< Resolved socket address
    struct sockaddr_storage resolved_address;
} ct_remote_endpoint_t;

// Example functions
ct_remote_endpoint_t* ct_remote_endpoint_new(void);
int ct_remote_endpoint_with_ipv4(
        ct_remote_endpoint_t* remote_endpoint,
        in_addr_t ipv4_addr
);
void ct_remote_endpoint_with_port(
    ct_remote_endpoint_t* remote_endpoint,
    uint16_t port
);
void ct_remote_endpoint_free(
    ct_remote_endpoint_t* remote_endpoint
);
\end{minted}
\label{lst:remote-endpoint}
\end{listing}

\begin{listing}[htbp]
\begin{minted}{C}
    // Step one allocate object on the heap
    ct_remote_endpoint_t* remote_endpoint = ct_remote_endpoint_new();
    // Check for allocation error
    if (!remote_endpoint) {
      return -1;
    }

    // Then modify it
    ct_remote_endpoint_with_ipv4(
        remote_endpoint,
        inet_addr("127.0.0.1")
    );
    ct_remote_endpoint_with_port(remote_endpoint, 1234); // example port

    // ... Then use it in a Preconnection for example...


    // And finally free it when done
    ct_remote_endpoint_free(remote_endpoint);
\end{minted} 
\caption{Remote endpoint lifetime example}
\label{lst:using-remote-endpoint}
\end{listing}

\subsection{Asynchrony} \label{asynchrony}
Being asynchronous and built around an event loop has
several implications for the external CTaps API, with the most important being
that most CTaps functions return immediately, with the results given 
through a callback invoked later.

When registering callbacks in CTaps this is done as close
to the relevant event as possible, for example Connection-related callbacks 
are registered when the Preconnection is initiated, an example of flow
and the registration of callbacks can be seen in \cref{lst:preconnection-init-flow}.
Not all callbacks have to be registered, although not registering important
callbacks such as the \mono{ready()} or \mono{closed()} would lead 
to the application missing important events that it should react to.

Since the results of most operations in CTaps is available through callbacks,
the application needs a way to connect these results to state external to the
callback, for example the filename to transfer when the \mono{ready()} callback is invoked.
The application could achieve this through global variables, but CTaps 
allows the user to pass a void pointer when registering callbacks so that
downstream applications can avoid global variables as much as possible.

When initiating a Connection, the void pointer will be stored in the Connection
structure and it can be fetched when the Connection is for example passed
to the \mono{ready()} callback.

CTaps uses callbacks to create Connection and Listener
objects themselves. Both of these objects require asynchronous operations
to instantiate, and therefore require the resulting object to be passed
through the application through callback.
The Connection is asynchronous both
because Candidate Gathering is asynchronous, but
also because it is only created after an actual connection has
been created over the network, in the case of connection-oriented
protocols. The Listener is asynchronous because it uses the
same candidate gathering as normal Connection initiation, which
is asynchronous in CTaps.

CTaps has two ways of notifying an application of errors, either
through a return code or an error callback.
Return codes are used for synchronous error while callbacks
are used for asynchronous errors.
Error codes in CTaps are
negative values matching
an errno corresponding to the given error. This matches the
error convention used by libuv~\cite{libuv_errors}.

If a function 
can fail asynchronously then it also takes error callbacks as
parameters, which can for example be invoked if a remote endpoint
rejects an connection attempt.

In theory, error handling could be unified into using callbacks
for all errors, but CTaps splits it into two categories in order
to fail as early as possible and to differentiate simple errors
such as illegal parameters from more advanced asynchronous problems,
which often may stem from things outside the application's control.

The asynchronous operations described above are all reliant on
the event loop provided by libuv as described in
\cref{sec:dependency-libuv}. CTaps
uses a single global event loop from libuv internally which has to
be initialized before operation can be performed.

This initialization is done through the
setup function \mono{ct_initialize()}, and teardown has to
be performed through \mono{ct_close()}.

After initializing CTaps, the application must perform any setup
required and start a Listener or initiate a Connection in order to
start interacting with the network. After this has been performed the
application must start the event loop with \mono{ctaps_start_event_loop()}.
This event loop will block the invoking thread until there are 
no more open handles performing work.

CTaps is not designed with multi-threading in mind and does not make use
of the threading API available through libuv~\cite{libuv_threading}. This
means that there is no internal synchronization and it expects applications
to only interact with CTaps through a single thread, which eventually will
start executing the event loop. Multi-threading is avoided in order to
minimize complexity and it would not add much value to the nature of CTaps
as a reference implementation of TAPS.

\begin{listing}[htbp]
    \begin{minted}{C}
ct_preconnection_t* preconnection = ct_preconnection_new(
    NULL,
    0,
    &remote_endpoint,
    1,
    transport_properties,
    NULL);

ct_connection_callbacks_t connection_callbacks = {
.establishment_error = on_establishment_error,
.ready = send_message_and_receive,
.per_connection_context = &test_context,
};

int rc = ct_preconnection_initiate(preconnection, &connection_callbacks);
    \end{minted}
    \caption[Example Preconnection initiation flow]{Example flow registering callbacks when initiating a Preconnection.}
    \label{lst:preconnection-init-flow}
\end{listing}

\subsection{Ownership Model}
Because all objects in CTaps are allocated on the heap in CTaps using \mono{ct_*_new()}
functions, it is important that the client frees the allocated object at the appropriate time.
In order to make this as intuitive and consistent as possible, CTaps has a clear ownership model
where the application retains ownership of any allocated objects and is responsible for freeing them.
In addition to the objects allocated synchronously using a \mono{ct_*_new()} function, applications
also gain ownership of Connection and Listener objects which are created asynchronously.

CTaps passes ownership to the created objects, even in the cases where
the application has not registered the
correct callbacks for receiving the object.
This will lead to a memory leak if an application is registered without
a \mono{ready()} callback, since the Connection will never be freed.

When freeing object which require teardown, the application must first
invoke the associated \mono{close()} function, as the object
will only be safe to free after the associated \mono{closed()}
callback has been invoked by CTaps.

Connection data and any waiting messages can still be read from a connection after it has
closed, but the data is invalidated on free.

The only location where the application does not gain ownership over an object is in
the receive callbacks passed to CTaps when an application wants to receive a message. 
Here, the received Message object and Message Context only remain valid for the duration
of the callback, and must be deep copied if they are to be referenced outside of this scope.
CTaps decided to deviate from the standard model here because of the frequency
of receiving Messages and that they would very rarely have to be stored for long term use.

An alternative to deep copying objects would be for CTaps
to implement reference counting for all objects, but this was decided to be unnecessarily complex
when compared to deep copying.

\section{QUIC Under TAPS} \label{sec:quicUnderTaps}
As the first TAPS implementation to feature QUIC as
an underlying protocol option, finding out exactly which features of
QUIC are available through the default TAPS interface was one of the
key research questions for this thesis. Since TAPS is an interface to
a generic transport layer, TAPS concepts, terminology and intended functionality
naturally does not map one to one onto concepts, terminology and functionality
of QUIC.

\subsection{Streams and Closing Them}
The largest differences when it comes to the expected behavior when
for QUIC and TAPS stems from the
the Connection and the Connection Group.
Since multistreaming is handled through Connection
Groups in TAPS, this means that a single QUIC connection maps to a Connection Group
while a TAPS Connection maps to a QUIC stream.

This has an especially large effect in when it comes to closing TAPS Connections
since QUIC has no support for a graceful teardown of a single stream.
A QUIC stream is intended to transfer
a single piece of data, with both the client and the server marking the transfer
from their respective ends as independently completed to close the stream.

While TCP connections can be closed one way, the typical behavior is that
when it closed by one peer, then the other peer closes the connection
in conjunction, which differs from the expected one-way closing of QUIC
streams.

This means that a TAPS system has to make a choice for how to close individual
TAPS connections when using QUIC as the underlying protocol. Based on TCP, the 
expected behavior
is that closing an individual Connection will close both the sending
and receiving side and start teardown. For QUIC, this can in theory be emulated by sending a \mono{FIN}
flag and a \mono{STOP_SENDING} frame. The peer then must respond with a \mono{RESET_STREAM}
frame if their stream is in a ready or sending state~\cite{rfc9000}.
A \mono{RESET_STREAM} frame tears down a QUIC stream in the sending direction,
which in combination with the \mono{FIN} flag sent would completely close the stream.

Even though this handling of underlying QUIC streams in TAPS is superficially
closer to the expected behavior from TCP, \mono{STOP_SENDING} is not a graceful close and is
only
expected to be used on error conditions. For a graceful teardown of a single
stream only, the expected behavior is both sides sending \mono{FIN} flags when
they are done sending.

In order to avoid unnecessary error behavior for QUIC the most natural mapping
for closing a TAPS Connection over QUIC is therefore to simply
send a \mono{FIN} flag in the sending direction and remain open in the receiving direction.
This will appear to the application in the form of it still having to monitor
the connection for incoming messages and it not receiving a closed()-callback until
after the peer has finished sending its data.

If a peer wants to gracefully close
a connection before all data has been received, the only way of handling it gracefully is to
close the entire QUIC connection. This naturally maps to the \mono{ConnectionGroup.close()}
functionality from RFC~9622~\cite{rfc9622}.

If the application takes care of performing TAPS connection cloning in a consistent way,
it can consistently use \mono{ConnectionGroup.close()} and achieve the same
expected behavior irrespective of the underlying protocol in the TAPS system, this
is however an example of underlying functionality leaking through to how the application 
should interact with the API.

This mapping of QUIC connections to TAPS connection groups also affects how
CTaps handles the notification of path changes. In TAPS, the PathChange 
event is per-Connection~\cite[Sec. 8.3.2]{rfc9622},
while QUIC migrations is per-Connection Group.

For CTaps this means that for each path change event every CTaps Connection is notified
separately, even though the event is always common to an
entire Connection Group. This could be mitigated by allowing registration
of callbacks on a per-Connection Group basis, but this not featured
in the current TAPS API.

There are also features related to streams in QUIC which are not available
through a TAPS Connection.
One example of this is the one-way abort mentioned
earlier through \mono{STOP_SENDING}, which requests that the peer abort their
sending, and \mono{RESET_STREAM}
which aborts the local sending side~\cite[Sec. 19.4, 19.5]{rfc9000}.
For aborting Connections, the TAPS API presents only the generic
\mono{Connection.abort()}~\cite[Sec. 10]{rfc9622}.
A TAPS implementation therefore has no way of knowing whether the application
wants to abort sending, abort receiving, or both. The only natural conclusion is
therefore to abort both sending and receiving, leaving the granular QUIC control 
unavailable to the application.

Another stream functionality which is not exposed through CTaps is
initiating unidirectional streams~\cite[Sec. 1]{rfc9000}.
Currently TAPS implementations have to assume that every new stream opened via \mono{Connection.clone()}
is bidirectional, although since the TAPS API is extensible a parameter could
easily be added when Cloning a Connection to support this.

Despite not being able to initiate unidirectional sending, a TAPS implementation
is able to receive an unidirectional stream initiated by the peer as expected by QUIC.
It can infer from the stream ID that it is unidirectional~\cite[Sec. 2.1]{rfc9000}
and mark the resulting Connection as closed for sending before presenting it to the application.

Withing QUIC, there are also several transport parameters~\cite[Sec. 18]{rfc9000} which are not
exposed through the TAPS API. TAPS does not feature a generic timeout value
which can be passed as \mono{max_idle_timeout}~\cite[Sec. 10.1]{rfc9000}.
While TAPS does support timeouts for when sending data, for example through
the Timeout for Aborting Connection Connection Property~\cite[Sec. 8.1.3]{rfc9622},
it does not feature a generic idle timeout for automatically closing a connection.

The TAPS API also does not support setting a maximum data to be received on either
a Connection Group or on a Connection, as supported on QUIC connections and streams
through the \mono{initial_max_data} and various \mono{initial_max_stream_data_*}
transport properties respectively.

In theory, an application
using TAPS could desire to set these, but they can also be assumed to
be handled by the underlying QUIC implementation and therefore not
necessarily be considered a gap in the TAPS API.
In CTaps' case, Picoquic provides
an optional \mono{picoquic_set_max_data_control} for
connection-level max data, but this is not surfaced through CTaps.
Picoquic does not provide a per-stream max data configuration.

While TAPS does allow the application to control the number of Connections
in a Connection Group through the groupConnLimit Connection Property~\cite[Sec. 8.1.9]{rfc9622},
it loses some granularity exposed through QUIC when controlling the number
of streams in a connection. QUIC allows control over both the maximum number
of both unidirectional and bidirectional streams through either the
\mono{initial_max_streams_uni} and \mono{initial_max_streams_bidi} or a
\mono{MAX_STREAMS} frame~\cite[Sec. 18.2, 19.11]{rfc9000}. 

\subsection{Using Multiple Paths}
QUIC path migration is supported in TAPS, and the migration
policy can be controlled through the
\mono{multipath} Selection Property.
For QUIC servers it can be used to set the
\mono{disable_active_migration} transport parameter~\cite[Sec. 18.2]{rfc9000},
while for clients it can be used to simply disable active migration.

TAPS does not expose the \mono{preferred_address} 
transport parameter~\cite[Sec. 18.2]{rfc9000} directly.
This parameter is used by a QUIC server when it wants to accept on a given address but wants
to migrate it to a new address shortly after the handshake~\cite[Sec. 9.6]{rfc9000}.
A workaround could be made by setting \mono{advertisesAltaddr} and
interpreting a second Local Endpoint as a preferred address for a Listener, but this
would be a fairly large break with the expected behavior from other protocols.

The combination of \mono{multipath} and the \mono{multipathPolicy} 
Connection Property~\cite[Sec. 8.1.7]{rfc9622} gives
a more granular control of the expected behavior when using
multipath than is featured in QUIC.
This is expected from TAPS
and it is clearly documented in RFC~9622 that Connection Properties
will only have an effect if they can actually be mapped onto the
underlying connection~\cite[Sec. 5]{rfc9622}. 

An overview of the
expected behavior according to RFC~9622 can be seen in
\cref{tab:multipath-combinations}.
The \mono{interactive} and \mono{aggregate} \mono{multipath} policies require
simultaneous use of multiple paths, which is not supported in RFC~9000 QUIC.
The IETF QUIC multipath extension~\cite{ietf-quic-multipath-21} is currently
under development and would enable these policies to be fully supported in a
future version of CTaps.

\begin{table}[htbp]
  \centering
  \begin{threeparttable}
    \caption{Effective behavior for \mono{multipath}/\mono{multipathPolicy} combinations}
    \label{tab:multipath-combinations}
    \begin{tabular}{llp{7cm}}
      \toprule
      \rowcolor{gray!25}
      \textbf{\mono{multipath}} & \textbf{\mono{multipathPolicy}} & \textbf{Description} \\
      \midrule
      \multicolumn{3}{l}{\textbf{\texttt{disabled}}\tnote{1}} \\
        & \multicolumn{1}{|l||}{---}
        & \mono{multipathPolicy} has no effect if multipath is disabled. \\
      \midrule
      \multicolumn{3}{l}{\textbf{\texttt{active}}} \\
        & \multicolumn{1}{|l||}{\texttt{handover}}    & Will initiate migration on path failure. \\[8pt]
        & \multicolumn{1}{|l||}{\texttt{interactive}} & Will use multiple paths concurrently to minimize latency. \\[8pt]
        & \multicolumn{1}{|l||}{\texttt{aggregate}}   & Will use multiple paths concurrently to maximize throughput. \\[8pt]
      \midrule
      \multicolumn{3}{l}{\textbf{\texttt{passive}}} \\
        & \multicolumn{1}{|l||}{\texttt{handover}}    & Will accept migration and multipath used by the remote. \\[8pt]
        & \multicolumn{1}{|l||}{\texttt{interactive}} & Will accept migration and multipath from the remote, will attempt to minimize latency.\\[8pt]
        & \multicolumn{1}{|l||}{\texttt{aggregate}}   & Will accept migration and multipath from the remote, will attempt to maximize throughput.\\[8pt]
      \bottomrule
    \end{tabular}
    \begin{tablenotes}
      \footnotesize
  \item[1] For QUIC servers, which cannot migrate~\cite[Sec. 1]{rfc9000}, \mono{disabled}
           can be used to set the \mono{disable_active_migration} transport parameter. For
           QUIC clients \mono{passive} and \mono{disabled} are equivalent since servers
           cannot migrate.
    \end{tablenotes}
  \end{threeparttable}
\end{table}

The remaining QUIC transport parameters which have no suitable mapping in TAPS
and cannot be assumed to be handled automatically are \mono{ack_delay_exponent} 
and \mono{max_ack_delay}~\cite[Sec. 18.2]{rfc9000}, which both can be
assumed to be reasonably handled through the underlying QUIC implementation.

There are also other features such as stateless reset which are not exposed through the TAPS
API directly, but are expected to be handled by the underlying QUIC implementation,
so that there is no need to expose them in a TAPS system.

An overview over the most important features in QUIC
and how they may into CTaps and TAPS in general can be seen in
\cref{tab:quic-taps-mapping}.


\begin{table}[htbp]
\centering
\begin{threeparttable}
\caption{Mapping of QUIC features to the TAPS API as implemented in CTaps.}
\label{tab:quic-taps-mapping}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
\rowcolor{gray!25}
\textbf{QUIC Feature} & \textbf{TAPS Mapping} \\
\midrule
\multicolumn{2}{l}{\textit{Supported in CTaps with expected behavior\tnote{1}}} \\
\midrule
\multicolumn{1}{l||}{ALPN negotiation} & Security Parameter \\
\multicolumn{1}{l||}{0-RTT establishment} & Message Property\\
\multicolumn{1}{l||}{TLS certificates} & Security Parameter\\
\multicolumn{1}{l||}{Connection Migration} & Multiple Local Endpoints\\
\multicolumn{1}{l||}{Bidirectional streams} & \mono{Clone()}\\
\multicolumn{1}{l||}{Stream Priority} & Connection Property\\
\midrule
\multicolumn{2}{l}{\textit{Supported in CTaps with deviation from expected behavior\tnote{1}}} \\
\midrule
\multicolumn{1}{l||}{QUIC connection\tnote{2}} & Connection Group\\
\multicolumn{1}{l||}{QUIC stream} & Connection\\
\multicolumn{1}{l||}{Unidirectional stream close (FIN)} & Connection \mono{Close()}\\
\multicolumn{1}{l||}{Connection migration\tnote{7}} & Selection Property\\
\multicolumn{1}{l||}{Path change events} & Per-Connection callback\\
\multicolumn{1}{l||}{\mono{initial_max_streams_bidi}} & Connection Property \\
\midrule
\multicolumn{2}{l}{\textit{Supported through TAPS, but not implemented in CTaps}} \\
\midrule
\multicolumn{1}{l||}{Multipath} & Selection Property\\
\multicolumn{1}{l||}{QUIC Datagrams~\cite{rfc9221}} & Message Property \\
\multicolumn{1}{l||}{PING Frames} & Connection Property \\
\multicolumn{1}{l||}{Key Update~\cite[Sec. 6]{rfc9001}} & Security Parameter Callback \\
\multicolumn{1}{l||}{ECN} & Message Property \\
\multicolumn{1}{l||}{\mono{disable_active_migration}} & Selection Property\\
\multicolumn{1}{l||}{\mono{max_udp_payload_size}\tnote{3}} & Connection Property \\
\multicolumn{1}{l||}{\mono{STOP_SENDING}\tnote{8}} & Connection \mono{Abort()}\\
\multicolumn{1}{l||}{\mono{RESET_STREAM}\tnote{8}} & Connection \mono{Abort()}\\
\midrule
\multicolumn{2}{l}{\textit{Not supported through TAPS}} \\
\midrule
\multicolumn{1}{l||}{Unidirectional stream init} & --- \\
\multicolumn{1}{l||}{\mono{initial_max_data}} & ---\\
\multicolumn{1}{l||}{\mono{initial_max_stream_data_*}} & ---\\
\multicolumn{1}{l||}{Idle Timeout\tnote{6}} & ---\\
\multicolumn{1}{l||}{\mono{ack_delay_exponent}} & ---\\
\multicolumn{1}{l||}{\mono{max_ack_delay}} & ---\\
\multicolumn{1}{l||}{Directional \mono{initial_max_streams_*}} & ---\\
\multicolumn{1}{l||}{\mono{preferred_address}} & ---\\
\multicolumn{1}{l||}{\mono{Stateless reset}    } & --- \\
\multicolumn{1}{l||}{Connection ID handling\tnote{4}} & ---\\
\bottomrule
\end{tabularx}
\begin{tablenotes}
\footnotesize
\item[1] Expected Behavior is defined to be a
mapping which behaves consistently with the protocol-agnostic TAPS API. Generally this means matching 
or exceeding the behavior of TCP as the underlying protocol.
\item[2] A QUIC connection maps to a Connection Group rather than a single Connection, since TAPS has no concept of streams within a Connection.
\item[3] Requires that recvMsgMaxLen is communicated to the peer.
\item[4] When to rotate, maximum limit etc. Would however be reasonable to assume that this is handled by the underlying QUIC implementation
\item[6] CTaps can set Connection timeout, but only when \enquote{trying to reliably deliver data to the Remote Endpoint}~\cite[Sec. 8.1.3]{rfc9622}.
\item[7] advertisesAltAddr is a boolean in TAPS, which cannot reflect that advertisement is voluntary in QUIC. This means that a TAPS implementation filtering
    protocols in candidate gathering cannot know if this property is mandatory in a protocol or not.
\item[8] Only available in unison through TAPS, not individually.
\end{tablenotes}
\end{threeparttable}
\end{table}
\todo{Fix ordering of notes}

Since the TAPS RFCs were written mostly before the QUIC design was finalized
in RFC~9000~\cite{rfc9000}, part of the motivation for developing CTaps
was to find out which QUIC features could not be presented through the
TAPS API.

The largest issues with creating a common abstraction for several different protocols like
TAPS is that there ends up being distinct concepts in each protocol which do
not necessarily map neatly into the shared abstraction layer concepts. This
is apparent in the differing behavior of Connections in CTaps depending on
whether the underlying protocol is QUIC or not.

With QUIC the Connection abstraction
becomes the connection group and treating individual streams as connections
leads to unexpected behavior, such as one-way closing by default except
when aborting the stream. This causes the application to have to interact
with TAPS in a way which does not reflect how they would interact with TCP,
for example always closing the entire connection group when done with a logical
connection to a server, to ensure that if the underlying protocol supports multistreaming
then the entire connection will be closed.

While this breaks slightly with 
with the principle of least surprise as described in RFC~9622~\cite{rfc9622}, it provides
greater flexibility and consistent behavior to the downstream application.

\section{Candidate Racing}
\todo{Think about talking about generic candidate racing too}
When it comes to connection racing, it is a useful feature but can
be hard to map for non-connection oriented protocols such as UDP, leading to
unexpected or even undesirable behavior.

While the default preferences in CTaps prefers TCP or
QUIC over UDP, if UDP is preferred but not required then there is no way for
racing to function properly.

Since UDP has no handshake to complete, CTaps has chosen to assume that a UDP connection
is established as soon as local endpoint has been set up. This means that
if the remote is does not support UDP, the racing will be unable to detect
this and UDP can potentially win the candidate racing,
but the remote will be unreachable.

An alternative
solution would be to send a test message and to wait for a reply, but
this relies on the remote peer having implemented this functionality and
therefore  risks designating legitimate endpoints as unresponsive. Since
the default preferences rank UDP lower than other supported protocols, it was
deemed that for CTaps the risk of UDP being ranked higher than other protocols when
the endpoint does not support it is quite low. The simple solution of
of assuming an established connection on opening the socket is adequate was therefore
chosen, for lack of a better option.

\section{Security}
The security of a TAPS system is reliant on the security of the underlying protocols.
While there are several key security measures which should be taken before
CTaps could be used in a security-sensitive production context, CTaps
also shows that it is possible to achieve equivalent security benefits
in a TAPS system as when using the underlying protocols directly.

% Handling of underlying credentials
Despite TAPS being designed before the official release of QUIC, it supports
the same security settings due to QUIC using TLS under the hood. It supports
both 0-RTT handshakes and the necessary client and server certificates for both
clients and servers using TLS. In addition it supports the required ALPNs and
and makes them accessible through Security Parameters. Security Parameters as
designed in TAPS and implemented by CTaps are by nature extensible, without
introducing breaking API changes.

Since TAPS provides a common API to several different protocols, there may
of course be situations where TAPS places a limitation on the security of a system.
This is however likely to be more related to the specific
TAPS implementation than to TAPS in general. One can for example imagine a scenario
where the key files should specifically be kept in memory and cannot be referred
to via a file name, then that could not be supported by CTaps as of the current
implementation. The TAPS API does however support this since it is generic
and only specifies server and client certificates through the use of
certificate bundles, but does not mandate the form which these bundles should take.
CTaps has chosen the form of referring to files on disk, but there is no reason why
it could not be adopted into taking the memory address of a buffer instead.

%TAPS downgrading security
The user of a TAPS system is also reliant on TAPS not silently downgrading
the required security of the application. Here TAPS needs to follow sane defaults
which default to mandating security, otherwise TAPS would be liable
to downgrade security since the insecure variants of protocols are at
an advantage during candidate racing since they require fewer RTTs.

In addition to mandating that TAPS implementations set the default allowed
security protocols to be the best available, RFC~9622 also mentions that applications
should use common safe defaults for security parameters
\cite{rfc9622}. This means that the
default should be requiring an encrypted connection. RFC~9622 also describe
the difference between disabled Security Parameters and opportunistic
Security Parameters.

Disabled Security Parameters are an option in order
to support compatibility with endpoints that do not support transport security
protocols, such as a pure HTTP website, while opportunistic Security Parameters
allow for the coexistence of secure and insecure transport layer protocols.
It is natural to assume that in cases where both insecure and secure transport
protocols are viable, any TAPS implementation would choose to prioritize the
secure protocol if performing candidate racing. For this case, the TAPS implementation
should also take the increased number of RTTs for the secure protocol into account
when calculating the stagger delay for candidate racing, in order to not give
the unencrypted protocols undue advantages.


Another concern with TAPS is that when applications utilize a large library
like a full TAPS implementation would be, the surface area for potential
attacks and errors becomes much larger.

The most useful mitigation for issues like this would be for a community
effort to develop a true fully featured open source implementation of TAPS
which achieves genuine real world usage. This would create a real incentive
to find and patch as many vulnerabilities as possible in the library.

Although TAPS aims to be as secure as possible by default, there are of course
trade-offs to be made between the default configuration of a TAPS system 
when it comes to feature-richness and security. With the relative complexity
and large number of configurations available within TAPS there is a genuine risk that
security minded applications may not have combed through all
available configurations in TAPS and therefore not be aware of a security risk.

An example of a potential unknown unknown is the potential of isolating sessions
through Security Parameters. By default this is set to false, allowing a TAPS implementation
to use an internal cache to optimize established connections. Using this cache
could lead to a leak of connection linkability, as observers over the wire could
observe different connections exhibiting the same behavior, and thereby induce that
they are made by the same client.

TAPS strikes a balance between being extremely
cautious and being secure by default by allowing certain configurations to be the most
performant by default instead of necessarily the most secure. Since a choice has to be made
for TAPS it can be assumed that extremely security conscious applications where
issues like fingerprinting are absolutely unacceptable will perform a careful comb
through of the security parameters, making the most cautious decision at every step.
The same cannot be assumed of a more casual application, having to comb through
Security Parameter configurations in order to find relatively large performance
gains to be made with relatively small security trade-offs.

%Racing reveal who uses TAPS? Other revealing options?
For applications where fingerprinting are absolutely unacceptable there are
also other issues related to TAPS, which cannot necessarily be traded away.
First of all, using a networking library which standardizes the interaction
with the network to a certain degree would naturally lead to potential fingerprinting
of clients. This is not something which can be configured away and is an
inherent property of software libraries whose behavior are exposed to the internet.
Without a large and diverse enough user base, fairly accurate assumptions can
be made about the users behind a connection, if the underlying library being
used can be ascertained. This becomes a larger risk if the behavior
if the underlying library is fairly unique, such as the extensive
candidate racing in TAPS.
