\chapter{Related work}\label{chap:related-work}
CTaps is not the only implementation of TAPS, as several
related open and closed source projects exist. This chapter
will describe the current state of the existing TAPS implementations,
forming a basis for why a new open source implementation is still
valuable. It will also compare the features currently present in CTaps with
the most complete previous implementations of TAPS, highlighting strengths
and weaknesses of CTaps.

\section{NEAT}
One of the first open source implementations of TAPS is 
NEAT~\cite{khademiNEATPlatformProtocolIndependent2017, neatCode},
whose development was funded by the EU horizon 2020 project
from 2015 to 2018~\cite{neatFunding}.
Developed alongside TAPS, the development of TAPS and NEAT naturally
affected each other. However, while NEAT halted development in 2018,
the TAPS API has continued to evolve, and the final TAPS RFC~was published
in 2025~\cite{rfc9622}.

NEAT is an asynchronous library and
uses an event loop, allowing for users to add their own callbacks.

The core architecture of TAPS and NEAT are similar, with both being
asynchronous, event-driven APIs, where the underlying connections
can be configured through properties.

Within the documented properties, NEAT has a different approach
to central concepts than TAPS. For example with the Transport Property,
the application can specify reliable, but also directly
specify one or more potential transport protocols to use. This
is in contrast to TAPS, where the application only interfaces
with the protocol through selection properties.

\section{NEATPy}
NEATPy~\cite{gundersenNEATPyStandardsConformantProtocolIndependent2020}
is a Python library developed as a master's thesis at UiO in 2020.
It was developed as an attempt to create a
Python layer to bridge the growing gap between NEAT and TAPS. Among
the goals was to move away from specifying the protocol at design time
instead using Selection Properties as specified by TAPS.

Because NEATPy is a Python shim on top of NEAT, certain parts of it are
limited by NEAT itself. For example,
using NEATPy it is not possible to specify the \texttt{Interface Instance or Type}
or the \texttt{Use Temporary Local Address} Selection Properties 
\cite{NEATPyDocumentationNEATPy}. In addition, NEATPy does not
support QUIC, as this is not supported by NEAT. 

Another
limitation of NEATPy is that it was developed before the finalization of
the TAPS standard. Developed in 2020, NEATPy is based mostly on
version 4 and 5 of the TAPS interface draft~\cite[p. 109]{gundersenNEATPyStandardsConformantProtocolIndependent2020}.
The TAPS API has since evolved, and was 
published as a proposed standard in January 2025
\cite{rfc9622}.

Since the finalization of NEATPy which implements
version 5 of the TAPS draft~\cite{ietf-taps-interface-05} several
new features have been added to TAPS.
For example, the support for giving more than one local and remote
endpoint when creating a Preconnection. This is done when the application
wants to indicate that several endpoints are eligible for the connection
or provide the same service.

In addition, there are several new Selection Properties that have been added
to TAPS. These are:
\begin{itemize}
    \item Keep-Alive Packets
    \item Use Temporary Local Address
    \item Multipath Transport
    \item Advertisement of Alternative Addresses
    \item Initiating Side Is Not the First to Write
\end{itemize}

Some selection properties have also been removed from the TAPS API:
\begin{itemize}
    \item Parallel Use of Multiple Paths
    \item Notification of excessive retransmissions
\end{itemize}

In addition to the outward differences to the API, there are also several
changes to the required internal workings of TAPS since the development of NEATPy.
For example, in the final RFC~TAPS has loosened the required
order of operations when working with selection properties.
Previously, there was a strict precedence in the order of
which preference values were to be applied.
This has since been
relaxed.% and the prioritization has been moved from being
% related to the preference level, to being ordered by which
% selection properties will affect the path selection.

\section{Network.framework}
Apple's Network.framework~\cite{networkFramework} is Apple's closed-source
TAPS implementation. It features several important TAPS abstractions
such as the Connection and Listener, but does also differ from the official
API described in the TAPS RFCs. The most important difference between
Network.framework and TAPS is that Network.framework does not currently
suppot generic Selection Properties as applications have to bind to a
specific transport protocol when initiating a Connection.

Network.framework does Happy Eyeballs-style racing of IPv4/IPv6 addresses,
but it does not support racing across multiple remote endpoints or across
different transport protocols, as described in RFC~9623~\cite{rfc9623}.

\section{PyTAPS}
PyTAPS~\cite{frankePythonasynciotaps2025} is a Python library aiming
to be fully compliant with TAPS. Unlike NEATPy, PyTAPS is not built
on top of NEAT, and therefore does not suffer from the same limitations
that NEATPy does. However, PyTAPS is based on the fourth version of the TAPS
interface draft~\cite{frankePythonasynciotaps2025} and has not had active
development since 2020. As such, it naturally suffers from the same
issues as NEATPy when it comes to the changes in the TAPS API.

\section{Dynamic TAPS}
dynamic-taps is a C implementation of TAPS, developed by the
American technology company F5~\cite{dukeDynamictaps2025}. It
is in the very early stages of development and has not had
active development since 2021, several years before the
finalzation of the TAPS RFCs.

\section{Comparison of TAPS implementations}\label{sec:tapsfeatures}
In order to compare the state with known existing work a table
comparing the features of CTap swith the known features of existing
TAPS implementations have been compiled and are displayed in~\cref{tab:comparison}.
Here, the most complete found TAPS implementations are compared in order to give an accurate
comparison point of the current state of CTaps. For the sake of brevity not all
implemnetations are included, instead the implementations which are deemed the
most complete are included. For this sake NEAT has also been excluded since
it would be a subset of the features achieved through the NEATPy shim layer.

\begin{table}[htbp]
\centering
\begin{threeparttable}
\caption{Comparison of features of various TAPS implementations}
\label{tab:comparison}
\small
\begin{tabular}{lccccc}
\toprule
\rowcolor{gray!25}
\textbf{Feature}                                     & \textbf{NEATPy\tnote{1}} & \textbf{PyTAPS} & \textbf{Network.framework}     & \textbf{CTaps}\\
\midrule
\multicolumn{1}{l||}{Post TAPS release\tnote{2}}    & \xmark                   & \xmark          & \cmark                     & \cmark         \\
\multicolumn{1}{l||}{Language}                       & Python                   & Python          & Swift / Obj.-C              & C         \\
\multicolumn{1}{l||}{Candidate Racing}               & \cmark                   & \cmark          & \xmark\tnote{3}            & \cmark         \\
\multicolumn{1}{l||}{TLS support}                    & \cmark                   & \xmark          & \cmark                     & \xmark         \\
\multicolumn{1}{l||}{QUIC support}                   & \xmark                   & \xmark          & \cmark                     & \cmark         \\
\multicolumn{1}{l||}{Multipath\tnote{4}}             & \xmark                   & \xmark          & \cmark\tnote{5}            & \cmark         \\
\multicolumn{1}{l||}{Early data}                     & \xmark                   & \xmark          & \cmark\tnote{6}            & \cmark         \\
\multicolumn{1}{l||}{Multistreaming}                 & \xmark                   & \xmark          & \cmark                     & \cmark         \\
\multicolumn{1}{l||}{Rendezvous}                     & \xmark                   & \xmark          & \cmark                     & \xmark         \\
\multicolumn{1}{l||}{OS support\tnote{7}}            & B, L, M                  & L, M\tnote{8}   & M                          & L         \\
\multicolumn{1}{l||}{Open source}                    & \cmark                   & \cmark          & \xmark                     & \cmark         \\
\bottomrule
\end{tabular}
\begin{tablenotes}
    \footnotesize
    \item[1] NEAT is excluded as its features would be a subset of NEATPy's.
    \item[2] Whether the project has been under active development after the full release of the TAPS RFCs in 2025.
    \item[3] Supports Happy Eyeballs-style IP family racing~\cite{networkframework_happy_eyeballs_wwdc}, but not full candidate racing as described by TAPS.
    \item[4] Taken to mean that a single connection can use more than one path in a single session,
        either concurrently or through connection migration.
    \item[5] Supports multipath TCP~\cite{apple_multipathservicetype}, QUIC support is unclear.
    \item[6] Supported for TCP~\cite{apple_allowfastopen}.
        Not supported in QUIC according to an Apple Engineer's post on the Apple Developer Forums in 2023~\cite{apple_dts_0rtt_2023}.
        Lacks explicit support in the NWProtocolQUIC.Options class~\cite{apple_nwprotocolquic}.
    \item[7] B = BSD, L = Linux, M = macOS, W = Windows.
    \item[8] Based on PyTAPS' README
\end{tablenotes}
\end{threeparttable}
\end{table}

While the most complete and production-ready implementation is Apple's Network.framework, it
also has several limitations. The largest limitation is naturally that it is a closed-source
implementation and is only designed to be compatible with the Apple ecosystem. This means
that the implementation cannot be adapted for other use cases outside of the Apple ecosystem.
The fact that Network.framework is closed source also means that the exact features supported
are harder to verify than for open-source implementations, since the exact
logic cannot be inspected. In the overview it has therefore been assumed that any feature not
documented clearly, such as connection migration for QUIC is not supported. THe external
API of Network.framewok is fairly true to the TAPS API, featuring abstractions such as the
NWConnection, NWListener and the NWConnectionGroup. 

There are however still large differences
in both the external and internal API of Network.framework and a true TAPS system. The largest
difference is that Network.framework does not support candidate racing as described in TAPS
RFC~9623~\cite{rfc9623}. While Network.framework supports Happy Eyeballs-style IP-family racing
it does not support racing multiple protocols and clients have to bind to a specific
protocol at compile time, by passing a NWParameters object which matches a specific
protocol when first creating the NWConnection.

An example of this pattern can be
seen in~\cref{lst:network-framework-example}. Here a new connection is opened through
QUIC using the ALPN \enquote{myproto}, the state handler is recognized and the conneciton
is started. This naturally reduces the flexibility of the system and forces applications
to bind to the underlying protocol, even though it is naturally more flexible than a pure
socket implementation.

\begin{listing}[htbp]
\begin{minted}{swift}
// Create a connection using QUIC
let connection = NWConnection(host: "example.com", port: 443, using: .quic(alpn: ["myproto"]))

// Set the state update handler to be notified when the connection is ready
connection.stateUpdateHandler = { newState in
    switch newState {
    case .ready:
        print("Connected using QUIC!")
    default:
        break
    }
}

// Start the connection with callback queue
connection.start(queue: queue)
\end{minted}
\caption{Network.framework example code of opening a QUIC connection to example.com~\cite{WWDC2021_10094}.}
\label{lst:network-framework-example}
\end{listing}

While the Network.framework example in~\cref{lst:network-framework-example} naturally belongs to
the currently most complete and production-ready implementation of TAPS, it also shows the areas
where CTaps has a chance to innovate and why an open source implementation of TAPS is valuable
despite the existence of Network.framework. Network.framework features a common interface
to interact with transport protocols and only forces the application to set the underlying
protocol once when first opening the Connection, it still forces the application to know
which protocol it wants to use, to implement fallback logic if the Connection over that
given protocl should fail and it is not completely intention-based since the application
is interacting with the name of the protoocl instead of which service it provides to the transport layer.

A full TAPS implementation would take the flexibility to another level as it would include
not only an abstract use of Selection Properties, but would also include candidate racing
allowing the application to only care about which service the transport protocol provides
instead of having to bind to a single protocol at compile time. This would also include automatic
connectin fallbacks if establishment fails, leading to a more robust client which could be
compatible with multiple different remote endpoints. In addition an open-source implementation would
be valuable as it would allow for public insight into special concerns when implementing
a TAPS system, enabling a wider adoptoin of TAPS and intention-based networking at the
transport layer, enabling faster innovation.
